### 2019-02-07 MEM.tex
Added the part on false positives.

\documentclass{article}

\usepackage{algorithm2e}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{authblk}
\usepackage[english]{babel}
\usepackage{blkarray}
\usepackage[font=small]{caption}
\usepackage{cite}
\usepackage{graphicx}

\setlength{\thickmuskip}{2mu plus 3mu minus 1mu}
\setlength{\medmuskip}{1mu plus 2mu minus 1mu}

\SetKwComment{Comment}{$\triangleright$\ }{}

% ---- Author affiliations ---- %

\renewcommand\Affilfont{\itshape\small}

% ---- Propositions, lemmas, defintions... ---- %

%\newtheorem{algorithm}{Algorithm}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
%\newtheorem{example}{Example}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
%\newtheorem{remark}{Remark}

% ---- Special environments (examples and remarks) ---- %

\newcounter{examplecounter}
\newenvironment{example}
{\small\vspace{0.5\baselineskip}
  \refstepcounter{examplecounter}%
  \noindent\textbf{Example \arabic{examplecounter}.}%
}{\vspace{-0.2\baselineskip}\begin{center}%
  $\star$\end{center}\vspace{0.5\baselineskip}}

\newcounter{remarkcounter}
\newenvironment{remark}
{\small\it\vspace{0.5\baselineskip}
  \refstepcounter{remarkcounter}%
  \noindent\textbf{Remark \arabic{remarkcounter}.}%
}{\vspace{0.5\baselineskip}}

\newenvironment{inset}
{\vspace{0.5\baselineskip}\begin{center}}
{\end{center}\vspace{0.5\baselineskip}}

% ---- Macros ---- %

\newcommand{\dn}[1]{\scriptstyle{\downarrow_{/#1}}}
\newcommand{\up}[1]{\scriptstyle{\uparrow_{/#1}}}
\newcommand{\nd}{\scriptstyle{|}}

%---------------------------------------------------------------

\title{Computing MEM seeding probabilities}

\author[1,2]{Guillaume J. Filion}
\author[1,2]{Eduard Zorita}
\affil[1]{Genome Architecture, Gene Regulation, Stem Cells and Cancer
Programme, Center for Genomic Regulation (CRG), The Barcelona Institute of
Science and Technology, Dr. Aiguader 88, Barcelona 08003, Spain.}
\affil[2]{University Pompeu Fabra, Doctor Aiguader, 08003 Barcelona,
Spain.}

\date{\today}

%---------------------------------------------------------------
%---------------------------------------------------------------

\begin{document}

\maketitle

\begin{abstract}
The abstract will come later.
A popular mapping heuristic is called \emph{seeding}, and a recent variant
is the \emph{MEM seeding} approach. Here we tackle the question of how to
evaluate the probability that MEM seeding hits are on-target. This in turn
expresses the confidence one should have in the location of the mapped DNA
fragment and can be used to calibrate the MEM seeding heuristic, or to
filter the hits at the desired level of confidence.
\end{abstract}

%---------------------------------------------------------------
%---------------------------------------------------------------

\section{Introduction}

\subsection{Mapping reads to genomes}

Suppose that a DNA fragment drawn at random from a known
genome\footnote{We will assume throughout that the sequence of the genome
is known exactly and with absolute certainty. This is not the case in
practical application, but this assumption will allow us to develop a
theory centered on the read.} is sequenced with an imperfect instrument;
how can we identify the location of the fragment in the genome?

We will refer to this question as the \emph{true location problem}.
It is central to countless applications, including discovering
disease-causing mutations; diagnosing viral, bacterial or fungal
infections; diagnosing cancer and choosing appropriate therapies;
selecting breeds of interest in agriculture; tracing human migrations from
ancient bones; identifying the donor of a biological sample; diagnosing
genetic diseaes at birth; finding compatible graft donors \textit{etc}.

The focus of this document is not to solve the true location problem
\textit{per se}, but rather to evaluate whether a solution is correct. At
first sight, whether we can map reads to their true location seems to
depend only on their lengths and on the error rate of the instrument.
However, there is another important factor to consider: most genomes
contain repeated sequences, \textit{i.e.} relatively large subsequences
(from $\sim20$~bp to more than 10~kb) that are present at more than one
location. So we cannot map the fragment with certainty if there exists an
exact copy of the sequence at somewhere else in the genome, because we
cannot know which of the two copies was sequenced.

This case is rare though. Repeated sequences are in general not identical,
but merely homologous\footnote{By that we mean that drawing nucleotides
uniformly at random is very unlikely to produce sequences with the
observed degree of identity.}. So when the DNA fragment is duplicated, we
can map it with high confidence provided we can distinguish it from its
duplicates. This in turn depends on their similarity with the target
fragment and on the error rate of the sequencer.

Because sequencing errors can convert a DNA fragment into one of its
duplicates, the \emph{true} location of a DNA fragment is not always the
\emph{best} (as measured by the identity between the sequence and the
candidate location). This leads us to consider the following question: how
can we identify the \emph{best} location of the fragment in the genome?

This is the \emph{best location problem}. It ammounts to finding the
optimal alignment between two sequences, and for this reason has received
substantial attention in bioinformatics. It is relevant for the present
discussion because the candidate solution of the \emph{true} location
problem is always the \emph{best} location. The next section will further
clarify the relationship between the two problems.

\subsection{Seeding heuristics}
\label{sec:seedheur}

There exists exact algorithms to solve the best location
problem\cite{pmid7265238,pmid5420325}, but they are too slow to process
the large amount of data generated by modern sequencers. Instead, one uses
heuristic methods, \textit{i.e.} algorithms that run fast, but may not
return the best location\cite{Waterman1984}. Heuristic mapping algorithms
thus have three possible outcomes: $i.$ the DNA fragment is mapped to the
best location, $ii.$ the DNA fragment is mapped to a suboptimal location,
$iii.$ the DNA fragment is not mapped at all.

The most popular heuristic to address the best location problem is to
first identify short regions of high similiarity between the read and the
genome, and then run an exact alignment algorithm around those regions.
The first step is usually called ``seeding'' and the regions of high
similarity are called ``seeds''. This is for instance the general strategy
of the popular alignment algorithm BLAST\cite{pmid2231712}.

If seeds can be discovered fast, we can quickly zoom in on a small set of
candidate locations, and use slower exact algorithms on a reduced data
set. The disadvantage is that the target location is not always in the
candidate set. In this case, it cannot be discovered and the outcome is
either a false positive or a false negative.

This imposes a trade-off between speed and accuracy between each candidate
location must be tested with a slow algorithm. If the seeding step is set
to produce many candidates, the target is likely to be among them and it
will be discovered. The downside is that this will take long. Conversely,
if the seeding step is set to produce few candidates, the process will
terminate faster but the target is less likely to sit among the candidates
so it is less likely to be discovered.

Importantly, the terms of this trade-off are not set in stone. Some
seeding methods are better than others\cite{pmid16533404,pmid20460430}, so
one can develop faster mapping algorithms at no cost for accuracy.

This framework was developed for the best location problem, but it also
holds for the true location problem. Actually, the process is exactly the
same but the outcomes $i.$ and $ii.$ are different. The first, called a
``true positive'' or an ``on-target hit'', is that the DNA fragment is
mapped to the \emph{true} location; the second, called a ``false
positive'' or an ``off-target hit'' is that the DNA fragment is mapped
somewhere else in the genome. A true positive thus requires that the
target (the true location) is in the candidate set, and that it is the
optimum.

\subsection{MEM seeds}

In general, seeds do not have to be exact matches\cite{pmid20460430}. For
instance, the algorithm PatternHunter\cite{pmid11934743} uses ``spaced
seeds'' that tolerate mismatches\footnote{The method is actually covered
by a patent (reference US20070088510A1).}. However, in this document, we
will always assume that a seed is an exact match with no gap. In other
words, a seed is a perfect match between a part of the read and a genomic
site.

The rest of the document will focus on a seeding heuristic known as ``MEM
seeding''. This approach gives good empirical results when mapping short
reads (it is for instance the strategy\footnote{To be exact, BWA-MEM uses
SMEM seeds (Super Maximal Exact Match seeds).} of the popular mapper
BWA-MEM\cite{li2013aligning}) but there is presently no theory to justify
this performance. A deeper understanding of MEM seeding would thus be
useful to calibrate the speed/accuracy trade-off of this heurisitic.

Before developing a formal theory, let us clarify the meaning of the MEM
acronym.

\begin{definition}
A Maximal Exact Match (MEM) seed is a subsequence of a read that is
present in the genome of interest and that cannot be extended --- either
because the read ends or because the extended subsequence is not in the
genome.
\end{definition}

This definition presents a computational challenge: to know if a given
subsequence is a MEM seed, we need to know if it exists somewhere in the
genome. This is a non trivial problem in itself, but fortunately there
exists practical methods to identify all the MEM seeds from a read, even
for very large genomes\cite{ferragina2000opportunistic,
ferragina2005indexing}. Such algorithms are crucial for the present
discussion, but describing them falls outside the scope of the document.
We refer the interested reader to the relevant
literature\cite{pmid24336412,pmid25399029,pmid23349213,pmid19389736,
li2013aligning}, and here on we will assume that MEM seeds are simply
available at all times. Looking ahead to section~\ref{sec:est}, let
us just say that this can be done with an algorithm called the
\emph{backward search}\cite{ferragina2000opportunistic}.

Before going further, we need to move a practical consideration out of the
way. It stands to reason that very short matches between the read and the
genome are uninformative. For instance, $99.7\%$ of all the possible
12-mers are present in the reference human genome, so an exact match of
size 12 or lower says nothing about the location of the DNA fragment.
In what follows, we will assume that subsequences of the read can only
match the original DNA sequence and/or its duplicates, but not the rest of
the genome. This will simplify the discussion without loss of generality
because non homologous matches are short and they can be ignored.

Fig.~\ref{fig:MEM_example} shows a concrete example. The target (the true
location of the read) is first potential location. The locations below
correspond to duplicate sequences. The mismatches between the genome and
the read are highlighted in bold. The read has three errors indicated by
stars, so \texttt{T}, \texttt{T} and \texttt{C} should actually be
\texttt{A}, \texttt{A} and \texttt{G}.

\begin{figure}[h]
\centering
\includegraphics[scale=1]{MEM_example.pdf}
\caption{\textbf{A concrete example of MEM seeding.}
A DNA fragment is sequenced with an imperfect instrument. It comes from
the top genomic location (Truth), but three errors were introduced
(indicated by stars). The sequence of the DNA fragment has three
duplicates at other genomic locations, shown below the target. Mismatches
between the read and the candidate sequence are highlighted in bold. MEM
seeds are indicated by grey boxes.}
\label{fig:MEM_example}
\end{figure}

The MEM seeds are indicated by grey boxes. Technically, they are part of
the read, but they are shown on the genome to highlight the position of
the match. One MEM seed cannot cover a smaller one (otherwise the smaller
would not be a MEM seed), but two consecutive MEM seeds can overlap, as in
the case of the frst two from the left. Note that overlapping MEM seeds
must match distinct sequences of the genome (otherwise neither of them
would be a MEM seed). Consecutive MEM seeds can also be disjoint, as in
the case of the second and the third from the left. In this case, they
must be separated by a nucleotide that is a mismatch against \emph{all}
the duplicates.

Observe that the rightmost MEM seed matches two genomic locations. This
case motivates the following definition, which will play an important role
for the development of the theory.

\begin{definition}
A \emph{strict} MEM seed has a single genomic match.
A \emph{shared} MEM seed has several genomic matches.
\end{definition}

Without surprise, the principle of MEM seeding is to use the genomic
matches of MEM seeds as candidate locations. Importantly, MEM seeds
shorter than a certain threshold $\gamma$ are ignored (in line with the
observation that short matches are uninformative). The parameter $\gamma$
is the tuning knob on the speed/accuracy trade-off. In
Fig.~\ref{fig:MEM_example} for instance, if $\gamma = 7$, there are three
candidate locations (the first, the second and the fourth) and if $\gamma
= 8$ there are only two (the first and the fourth).

Not every MEM seed makes the target a candidate location, so we introduce
the following terms to simplify the discussion.

\begin{definition}
Assume that the thresohld $\gamma$ is fixed. A \emph{good} seed is an
on-target MEM seed of size $\gamma$ or greater. Every other MEM seed is a
\emph{bad} seed. In other words, the true location is in the candidate set
if and only if the read has a good seed.
\end{definition}

Compared to seeds of fixed size, MEM seeds have at least two
counter-intuitive properties. The first is that there are cases where
there is no good seed, even when $\gamma$ is decreased.
Fig.~\ref{fig:full_masking_example} shows such an example. Even though
there is a single sequencing error, the read does not have any MEM seed
matching the target. Lowering $\gamma$ does not change this, so there is
no way to discover the true location (even though it is also the best
location).

\begin{figure}[h]
\centering
\includegraphics[scale=1]{full_masking_example.pdf}
\caption{\textbf{MEM masking case 1.}
The read, the MEM seeds and the sequences are as in
Fig.~\ref{fig:MEM_example}. The MEM seeds matching the two incorrect
locations at the bottom effectively mask the target, so it cannot
be discovered. MEM masking can occur even when the true location is the
best and when there is a single error on the read.}
\label{fig:full_masking_example}
\end{figure}

The second counter-intuitive property of MEM seeding is that there are
cases where there would be a good seed if the read were shorter.
Fig.~\ref{fig:short_vs_long} shows an example of this case. Again, there
is no on-target MEM seed, but there would be if the read were two
nucleotides shorter on the right side. Indeed, in this case there would be
a shared MEM seed matching the first and the second locations. It would
also be a good seed for $\gamma \leq 12$.

\begin{figure}[h]
\centering
\includegraphics[scale=1]{short_vs_long_example.pdf}
\caption{\textbf{MEM masking case 2.}
The read, the MEM seeds and the sequences are as in
Fig.~\ref{fig:MEM_example}. There is no on-target MEM seed, but there
would be a shared MEM seed if the read were two nucleotides shorter. The
true location has been masked by sequencing the last two nucleotides.}
\label{fig:short_vs_long}
\end{figure}

These examples show that MEM seeds can perform worse than seeds of fixed
size, again raising the question why they seem to give good performance in
practice. The theory developed below will allow us to compute the
probability that a read has no good seed and thus that the true location
is missed. This probability is the key to evaluating and calibrating the
MEM seeding heuristic.

From now we will make the simplifying assumption that a read with a good
seed is always mapped at the correct location. In other words, we assume
that a good seed is necessary and sufficient for a true positive or
outcome $i.$ In practice this is not always the case, but this assumption
shifts the burden of mapping to the alignment algorithm and allows us
focus on improving the seeding heuristic.

We first explain a detailed strategy to compute the probablity of outcome
$i.$ We then explain how to compute the probability of outcome $iii.$,
\textit{i.e.} that the seeding step returns no hit. Combining these two
quantities allows us to compute the probability of interest, which is that
the read is mapped to an incorrect location or outcome $ii.$ Finally, we
provide a computationally efficient method for estimating the parameters
that are necessary to estimate the probabilities of interest.

Before going further, we recommend the readers that are not already
familiar with MEM seeding to take a moment to study the examples above.
The complexity of the next sections will go \textit{crescendo} and they
will draw heavily from the concepts and definitions above.

\section{Combinatorial construction}

In this section we represent reads as combinatorial ojects. The purpose of
this construction is to give a generative model of reads without good
seeds so that we can later compute their probability of occurrence.

\subsection{Hard and soft masking}

For each potential location including the target, we define the
\emph{match streak} at a given nucleotide as the number of nucleotides
since the last mismatch on the left, or since the beginning of the read.
For instance, in Fig.~\ref{fig:MEM_example}, the match streak of the
target at the ninth nucleotide is $1$ while for the second potential
location, it is $7$.

\begin{definition}
At a given position of the read, an off-target location is a \emph{hard
mask} if its success streak is strictly longer than the match streak of
the target. An off-target location is a \emph{soft mask} if its has the
same match streak as the target.
\end{definition}

For instance, the second location in the example above is a hard mask at
the ninth nucleotide. On Fig.~\ref{fig:MEM_example}, one can further see
that at the second nucleotide, all the off-target locations are soft masks
and that at the last nucleotide, the second location is a soft mask.

From the definition, the last nucleotide of every strict on-target MEM
seed is always unmasked. Conversely, an unmasked nucleotide always belongs
to exactly one strict on-target MEM seed. Also, the last nucleotide of
every shared on-target MEM seed is always soft-masked, but a soft-masked
nucleotide does not always belong to a shared on-target MEM seed.

Since hard and soft masks inform us about the positions of on-target
MEM seeds, we construct an alphabet that encodes the masking status of the
nucleotides.

\subsection{The MEM alphabet}

We recode the reads as sequences of letters from the so-called MEM
alphabet $\mathcal{A} = \{\square, |, \uparrow_{/1}, \uparrow_{/2},
\uparrow_{/3} \ldots, \downarrow_{/0}, \downarrow_{/1}, \downarrow_{/2},
\ldots\}$.

The symbols $\downarrow_{/m}$ indicate that the nucleotide is a sequencing
error and $m \geq 0$ is the number of off-target locations that
\emph{match} the nucleotide. Since a sequencing error is always a mismatch
against the target, the symbol $\downarrow_{/0}$ indicates that the
ncleotide is a mismatch against \emph{every} candidate.

The symbols $\uparrow_{/j}$ indicate a change in masking status: the
nucleotide is not masked but its left neighbor is. The index $j \geq 1$ is
the number of nucleotides since the last mismatch or since the beginning
of the read.

All the other nucleotides are represented by the $\square$ symbol.
Finally, the $|$ symbol is appended after the last nucleotide of the read.
It is an obligatory terminator that does not correspond to an actual
nucleotide.

Note that the numbers decorating $\uparrow$ and $\downarrow$ are different
in nature. In the case of the symbol $\downarrow$ it is an information
about the number of sequences matching the read; in the case of the symbol
$\uparrow$ it is the position of the previous sequencing error.
Fig.~\ref{fig:sketch_extended} shows the encoding of the read from
Fig.~\ref{fig:MEM_example} in the MEM alphabet.

\begin{figure}[h]
\centering
\includegraphics[scale=.85]{sketch_extended.pdf}
\caption{\textbf{The extended MEM encoding}.
The read of Fig.~\ref{fig:MEM_example} is represented in the MEM alphabet.
The arrows departing from those numbers help understand their meaning. The
$\downarrow$ symbol is decorated by the number of locations that match the
nucleotide. The $\uparrow$ symbol is decorated by the number of
nucleotides from the last error or from the beginning of the read. Also
note the presence of the terminator $|$ at the end of the read.}
\label{fig:sketch_extended}
\end{figure}

The MEM alphabet captures the masking status of the nucleotide as follows:
If we denote the number of off-targets as $N$, the symbol
$\downarrow_{/m}$ indicates that the nucleotide has $m$ hard masks and
$N-m$ soft masks. In addition, the symbols $\uparrow_{/j}$ indicate that
the nucleotide is unmasked.

In the MEM alphabet, strict on-target MEM seeds are the longest stretches
of symbols containing $\uparrow$ and not containing $\downarrow$. Indeed,
such a stretch matches the target, it matches the target only because it
contains at least one unmasked nucleotide (marked by $\uparrow$), and it
cannot be extended because it is flanked by sequencing errors or by the
ends of the reads. Note that there is exactly one symbol $\uparrow$ per
strict on-target MEM seed, and therefore that two symbols $\uparrow$ must
be separated by at least one symbol $\downarrow$.

Shared on-target MEM seeds are the longest stretches of $\square$ symbols
flanked by $\downarrow_{/0}$, or by the ends of the read. Indeed, such a
stretch is a MEM seed because it matches the target and it cannot be
extended ($\downarrow_{/0}$ is a mismatch against every sequence). Also,
it cannot be a strict on-target MEM seed because it does not contain the
$\uparrow$ symbol, so it must be shared.

\subsection{Reads as extended MEM segments}

Reads can be viewed as sequences of symbols, or as sequences of
\emph{segments}. For the purpose of constructing reads without good seeds,
the second point of view turns out to be essential.

\begin{definition}
\label{def:segment}
A segment is a maximal sequence of $\square$ symbols flanked on the right
side by one of the symbols $\uparrow_{/j}$, $\downarrow_{/m}$, or $|$. The
one and only segment flanked by $|$ is called the tail. It is the only
segment that can have size 0.
\end{definition}

In line with definition~\ref{def:segment}, the symbols $\uparrow_{/j}$,
$\downarrow_{/m}$ and $|$ are referrred to as \emph{terminators}.
Fig.~\ref{fig:sketch_segment} shows the segment representation of the read
from Fig.~\ref{fig:sketch_extended}.

\begin{figure}[h]
\centering
\includegraphics[scale=.85]{sketch_segments.pdf}
\caption{\textbf{The extended MEM segment encoding}.
The read from Fig.~\ref{fig:sketch_extended} is represented as a sequence
of segments. Each segment is terminated by either $\uparrow_{/j}$, or
$\downarrow_{/m}$ or $|$. The rightmost fragment is the tail. The numbers
below indicate the size of each segment in number of nucleotides.}
\label{fig:sketch_segment}
\end{figure}

We now have a representation of reads as sequences of segments. Only some
of these sequences are possible. For instance, a segment terminated by
$\uparrow$ cannot follow another. The task, however, is not to give a
construction of reads, but a construction of reads without good seeds. For
this, we need to introduce a new set of tools.

\section{The computational strategy}
\label{sec:symbolic}

We turn to the question of translating the extended MEM segment
representation to a construction of reads without good seeds. For this, we
use tools borrowed from analytic combinatorics. We give the minimum amount
of background that is necessary to expose the theory. A more detailed
account of the strategy to compute seeding probabilities with analytic
combinatorics has been published
elsewhere\cite{filion2017analytic,filion2018analytic}. We refer the
readers interested in learning more about analytic combinatorics to the
textbooks\cite{flajolet2009analytic,sedgewick2013introduction}.

\subsection{Weighted generating functions}

The central object of analytic combinatorics is the \emph{generating
function}.

\begin{definition}
\label{def:GF}
Let $\mathcal{A}$ be a set of combinatorial objects characterized by a
size in $\mathbb{N}$ and a weight in $\mathbb{R}^+$. The weighted
generating function of $\mathcal{A}$ is defined as

\begin{equation}
\label{eq:GF1}
A(z) = \sum_{a \in \mathcal{A}} w(a) z^{|a|},
\end{equation}
where $|a|$ and $w(a)$ denote the size and weight of the object $a$,
respectively. Expression (\ref{eq:GF1}) also defines a sequence $(a_k)_{k
\geq 0}$ such that

\begin{equation*}
A(z) = \sum_{k=0}^\infty a_k z^k.
\end{equation*}

By definition $a_k = \sum_{a \in A_k}w(a)$, where $A_k$ is the class of
objects of size $k$ in $\mathcal{A}$. The number $a_k$ is the
total weight of objects of size $k$.
\end{definition}

To give a concrete example, if the probability of the symbol
$\downarrow_{/0}$ is $q$, the weighted generating function of a segment
containing only this symbol is $qz$. In our case, we are interested in the
weighted generating function of reads without good seeds, so that the
coefficient $a_k$ is equal to the probability that a read of size $k$ has
no good seed.

The motivation for introducing weighted generating functions is that
operations on combinatorial objects translate into operations on their
weighted generating functions. If $A(z)$ and $B(z)$ are the weighted
generating functions of two mutually exclusive sets $\mathcal{A}$ and
$\mathcal{B}$, the weighted generating function of $\mathcal{A} \cup
\mathcal{B}$ is $A(z) + B(z)$, as evident from expression (\ref{eq:GF1}).
Size and weight can be defined for pairs of objects in $\mathcal{A} \times
\mathcal{B}$ as $|(a,b)| = |a| + |b|$ and $w(a,b) = w(a)w(b)$. In other
words the sizes are added and the weights are multiplied.  With this
convention, the weighted generating function of the Cartesian product
$\mathcal{A} \times \mathcal{B}$ is $A(z)B(z)$. This simply follows from
expression (\ref{eq:GF1}) and from

\begin{equation*}
A(z)B(z) =
\sum_{a\in \mathcal{A}}w(a)z^{|a|} \sum_{b\in \mathcal{B}}w(b)z^{|b|}
= \sum_{(a,b) \in \mathcal{A} \times \mathcal{B}} w(a)w(b)z^{|a|+|b|}.
\end{equation*}

These two operations are all we need in order to compute the weighted
generating functions of the reads of interest. Addition corresponds to
creating a new family by merging reads from families $\mathcal{A}$ and
$\mathcal{B}$. Multiplication corresponds to creating a new family by
concatenating reads from families $\mathcal{A}$ and $\mathcal{B}$.

\subsection{Analytic representation}

The symbolic method is a strategy belonging to the general \emph{analytic
combinatorics} framework. The idea is to combine simple combinatorial
objects into more complex objects. Each combinatorial operation on the
objects corresponds to a mathematical operation on their weighted
generating functions. One can thus obtain the weighted generating function
of complex objects, and from there extract their probability of
occurrence.

The combinatorial construction of reads using the extended MEM alphabet
gives us a method to express their weighted generating function from the
simpler generating function of segments. More specifically, we can
describe complex sequences of segments through a so-called \emph{transfer
matrix} $M(z)$ that specifies which pairs of consecutive segments are
allowed and which are forbidden in the reads of interest.

The segment types are identified by their terminators arranged in a
predefined oder. The entry of $M(z)$ at coordinates $(i,j)$ is the
weighted generating function of segments with the $j$-th terminator that
can be inserted immediately after segments with the $i$-th terminator. The
entries of the matrix $M(z) + M(z)^2 + \ldots = M(z) \cdot (I-M(z))^{-1}$
are the weighted generating functions of the reads of all sizes using all
the allowed combinations of segments.

The entry of interest from $M(z)$ is the one that corresponds to the
terminators $\downarrow_{/0}$ and $|$ because it represents the reads of
any size whose first segment can be inserted after $\downarrow_{/0}$, and
whose last segment is terminated by $|$. Indeed, the $|$ terminator marks
the end of the read, so it must be the final symbol of every valid read.
As for the first segment, observe that all the segments that can follow
the $\downarrow_{/0}$ symbol are exactly those that can be at the frst
postion of the read\footnote{One could have introduced a special start
symbol. It is equivalent to an invisible $\downarrow_{/0}$ prepending
the reads.}.

In summary, the transfer matrix $M(z)$ contains the weighted generating
function of interest. Our next task is thus to specify $M(z)$.

\section{The transfer matrix}
\label{sec:transfer_mat}

\subsection{Error model and divergence of the duplicates}

We assume that the target sequence has $N$ duplicates, so that there are
$N$ off-target sequences. We further assume that duplication was
instantaenous and that all $N+1$ sequences diverge independently of each
other at a constant rate. In other words, we ignore the complications due
to the genealogy of the duplication events and we simply assume that at
each position, any given duplicate is identical to the target with
probability $1-\mu$. If it is not, we assume that the duplicate has any of
the remaining three nucleotides with equal probability (\textit{i.e.} each
is found with probability $\mu/3$).

We also assume that the sequencing instrument has a constant substitution
rate $p$, and that insertions and deletions never occur. In case of a
substitution, we again assume that the remaining three nucleotides are
equally likely.

Read errors are always mismatches against the target (because we assume
that the target is the true sequence), and they match each off-target
sequence with probability $\mu/3$. Correct nucleotides are always a match
for the target, and they match each off-target sequence with probability
$1-\mu$.

We are now ready to describe the transfer matrix $M(z)$. We do this entry
by entry, in order of increasing complexity.

\subsection{Segments following $\uparrow_{/j}$}

A segment terminated by $\uparrow_{/j}$ is the beginning of an on-target
MEM seed of size at least $j$. The seed reaches the next sequencing
error or the end of the read, so the number of $\square$ symbols in the
next segment must be at most $\gamma-j-1$ and it must be terminated by a
$\downarrow$ symbol or by the tail terminator $|$.

On this segment, the matches between the read and the off-target sequences
are irrelevant. The weighted genearting function of each $\square$ symbol
is simply $qz$ ($q$ is the probability that the nucleotide is not an error
and $z$ marks objects of size $1$). The following definition will simplify
the notations.

\begin{definition}
The probability that a symbol is $\downarrow_{/m}$ given that the
nucleotide is a read error is
\begin{equation}
\label{eq:omega}
\omega_m = {N \choose m} \big(1 - \mu/3\big)^{N-m} \big(\mu/3\big)^m.
\end{equation}
\end{definition}

The weighted generating function of the $\downarrow_{/m}$ terminator is
thus $\omega_m pz$ (the probability of the symbol $\downarrow_{/m}$ is
$\omega_m p$ and $z$ marks objects of size $1$).

The entries of the transfer matrix that correspond to transitions from
$\uparrow_{/j}$ to other segments are now easy to find. The weighted
generating function of the segments terminated by $\downarrow_{/m}$
following a segment terminated by $\uparrow_{/j}$ is
\begin{equation}
\label{eq:D}
D_{j,m}(z) = \omega_m pz \sum_{i=0}^{\gamma-j-1} (qz)^i.
\end{equation}

And the weighted generating function of the tail segments following
segments terminated by $\uparrow_{/j}$ is
\begin{equation}
\label{eq:E}
E_j(z) = \sum_{i=0}^{\gamma-j-1} (qz)^i.
\end{equation}

\subsection{Segments following $\downarrow_{/m}$}

The symbol $\downarrow_{/m}$ signifies that the nucleotide has $m$ hard
masks and $N-m$ soft masks. If all the masks disappear before the first
read error, the next terminator will be a $\uparrow$ symbol, otherwise it
will be a $\downarrow$ symbol. We separate the cases based on the
terminator of the segment.

\subsubsection*{Case 1: the terminator is $\uparrow_{/i}$}

\begin{definition}
The probability that a given off-target sequence contains a mismatch in a
sequence of $i$ error-free nucleotides is
\begin{equation}
\label{eq:xi}
\xi_i = 1-(1-\mu)^i.
\end{equation}
\end{definition}

With this notation, the probability that at least one mask survives a
sequence of $i$ error-free nucleotides is thus $1-\xi_i^N$, and the
probability that there remains a mask at the $i-1$-th but not at the 
$i$-th error-free nucleotide is $\xi_i^N - \xi_{i-1}^N$. From this we
conclude that the weighted generating function of the segments terminated
by $\uparrow_{/i}$ following a segment terminated by $\downarrow_{/m}$ is
\begin{equation}
\label{eq:B}
B_i(z) = \Big( \xi_i^N-\xi_{i-1}^N \Big) (qz)^i.
\end{equation}

In the equation above, $q^i$ is the probability that there is no
sequencing error in $i$ nucleotides and $z^i$ marks objects of size $i$.
The fact that the reads have no good seeds imposes $i < \gamma$. Also note
that this expression is the same for all $\downarrow$ symbols; it does not
depend on $m$.

\subsubsection*{Case 2a: the terminator $|$ comes before the $\gamma$-th
nucleotide}

In this case there can be no good seed because the read terminates too
early. We must just enforce the condition that at least one of the $N$
masks survives until the end of the segment in order to exclude the
$\uparrow$ symbol. The weighted generating function is
\begin{equation*}
\sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i.
\end{equation*}

\subsubsection*{Case 2b: the terminator $|$ comes after the $\gamma$-th
nucleotide}

In this case, the soft masks do not hide the target: even if one of them
survives until the end of the read, there will be a good seed (shared in
this case). To make sure that the read has no good seed, we must enforce
the condition that at least one hard mask survives until the end of the
segment (which is impossible if $m = 0$). The weighted generating function
is
\begin{equation*}
\sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \Big) (qz)^i.
\end{equation*}

Finally, the weighted generating function of the tail segments following
segments terminated by $\downarrow_{/m}$ is
\begin{equation}
\label{eq:C}
C_m(z) =
\sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i +
  \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \Big) (qz)^i.
\end{equation}

\subsubsection*{Case 3a: the terminator $\downarrow_{/n}$ comes before the
$\gamma$-th nucleotide}

In this case, there can be no good seed and we must only exclude the
$\uparrow$ terminator. The weighted generating function is
\begin{equation}
\omega_n pz \sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i.
\end{equation}

\subsubsection*{Case 3b: the terminator $\downarrow_{/n}$ comes after the
$\gamma$-th nucleotide}

This case is by far the most convoluted. Since the segment contains at
least $\gamma$ error-free nucleotides, we must ensure that it does not
contain a good seed. This will be the case if any of the two following
conditions is validated: $i)$ at least one hard mask covers all the
error-free nucleotides, or $ii)$ all the hard masks vanish but at least
one soft mask covers the whole segment (including the terminator).

The two conditions are mutually exclusive by construction. They are
graphically represented in the diagram below. The left panel corresponds
to case $i)$ and the panel to case $ii)$. The top row represents the
target, and the bottom rows represent off-target sequences (using the same
symbols as in Fig.~\ref{fig:MEM_example}).
\begin{inset}
\includegraphics{masks.pdf}
\end{inset}

Whenever a hard mask (here the first off-target sequence) covers the
nucleotides as shown in the left panel, there can be no good seed,
irrespective of the value of $\gamma$. The positions marked with a
question mark are irrelevant, they cannot change the fact that there is no
on-target MEM seed. If the hard masks vanish, as in the right panel, then
it all depends on the soft masks. If a soft mask covers the whole segment,
then there can be no good seed, irrespective of the value of $\gamma$.

Condition $i)$ has probability $\big(1 - \xi_i^m \big)$ and the
weighted generating function is thus
\begin{equation*}
\omega_n pz \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \Big) (qz)^i.
\end{equation*}

Condition $ii)$ is more convoluted, so we introduce some further
notations to solve this sub-case.
\begin{definition}
The probability that a given off-target sequence contains a mismatch in a
sequence of $i$ error-free nucleotides followed by an error is
\begin{equation}
\label{eq:eta}
\eta_i = 1-(1-\mu)^i\mu/3.
\end{equation}
\end{definition}

The probability of condition $ii)$ is
\begin{equation*}
\xi_i^m \Big(1 - \eta_i^{N-m} \Big),
\end{equation*}
but we need to break up this term among all the terminators
$\downarrow_{/n}$, ($0 \leq n \leq N$). For this, we decompose the sum on
the number of soft masks that run to the end of the segment, including the
terminator. The probability that there are $r \geq 1$ such soft masks is
\begin{equation*}
{N-m \choose r} (1 - \eta_i)^r \eta_i^{N-m-r}.
\end{equation*}

Each of them matches the nucleotide at the end of the segment, so the
total number of matches is $r$ plus the number of sequences that also
match the last nucleotide, among the remaining $N-m-r$ soft masks and $m$
hard masks.

Let us start with the $N-m-r$ that vanished somewhere in the segment. The
conditional probability that they match the last nucleotide given that
they vanish somewhere in the segment is $\mu/3 \cdot \xi_i / \eta_i$. The
conditional probability that they do not match it is $(1-\mu/3) / \eta_i$.
For the $m$ hard masks, the probability that they match the last
nucleotide given that they vanish somewhere in the segment is simply
$\mu/3$, and the probability that they do not match it is $1-\mu/3$.

The probability that the last nucleotide of the segment matches $n-r$ of
the $N-r$ masks that vanished before the terminator is
\begin{equation*}
\frac{(\mu/3)^{n-r}(1-\mu/3)^{N-n}}{\eta_i^{N-m-r}}
\psi_{i,m,n,r}\text{, where \;}
\psi_{i,m,n,r} = \sum_{q \geq 0}{m \choose q}{N-m-r \choose n-r-q}
\xi_i^{n-r-q}.
\end{equation*}

Finally, the probability that a total of $n$ off-target sequences match
the terminator and thus that the terminator is the symbol
$\downarrow_{/n}$ is
\begin{eqnarray*}
&\;& \sum_{r\geq1} {N-m \choose r}
(1 - \eta_i)^r (\mu/3)^{n-r} (1-\mu/3)^{N-n} \psi_{i,m,n,r} \\
&=& (\mu/3)^n(1-\mu/3)^{N-n} \sum_{r\geq1} {N-m \choose r}
  (1 - \mu)^{ri} \psi_{i,m,n,r} \\
&=& \omega_n \cdot \zeta_{i,m,n},
\end{eqnarray*}
where
\begin{equation}
\label{eq:zeta}
\zeta_{i,m,n} = \sum_{r\geq1} {N-m \choose r}
(1-\mu)^{ri} \psi_{i,m,n,r} \bigg/ {N \choose n}.
\end{equation}

The weighted generating function of the segments terminated by
$\downarrow_{/n}$ following segments terminated by $\downarrow_{/m}$ is
\begin{equation}
\label{eq:A}
A_{m,n} =
\omega_n pz \sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i + \omega_n
pz \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \cdot
(1- \zeta_{i,m,n}) \Big) (qz)^i.
\end{equation}

\subsection{The expression of the transfer matrix}
\label{sec:expression_of_M}

Collecting and arranging the results above, the final expression of the
transfer matrix $M(z)$ appears as
\begin{equation*}
\begin{blockarray}{cccccccc}
   & \dn{0} & \ldots & \dn{N} & \up{1} & \ldots & \up{\gamma-1} & \nd \\
\begin{block}{c[ccccccc]}
\dn{0} & A_{0,0}(z) & \ldots & A_{0,N}(z) & B_1(z) & \ldots &
    B_{\gamma-1}(z) & C_0(z) \\
\vdots & \vdots & \ddots & \vdots & \vdots & \ddots &
    \vdots & \vdots \\
\dn{N} & A_{N,0}(z) & \ldots & A_{N,N}(z) & B_1(z) & \ldots &
    B_{\gamma-1}(z) & C_N(z) \\
\up{1} & D_{1,0}(z) & \ldots & D_{1,N}(z) & 0 & \ldots & 0 & E_1(z) \\
\vdots & \vdots & \ddots & \vdots & \vdots & \ddots &
    \vdots & \vdots \\
\up{\gamma-1} & D_{\gamma-1,0}(z) & \ldots & D_{\gamma-1,N}(z) & 0 &
  \ldots & 0 & E_{\gamma-1}(z) \\
\nd & 0 & \ldots & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray}
\end{equation*}
where
\begin{gather}
\tag{\ref{eq:A}}
A_{m,n} =
\omega_n pz \sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i + \omega_n
pz \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \cdot
(1- \zeta_{i,m,n}) \Big) (qz)^i \\
\tag{\ref{eq:B}}
B_i(z) = \Big( \xi_i^N-\xi_{i-1}^N \Big) (qz)^i \\
\tag{\ref{eq:C}}
C_m(z) =
\sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i +
  \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \Big) (qz)^i \\
\tag{\ref{eq:D}}
D_{j,m}(z) = \omega_m pz \sum_{i=0}^{\gamma-j-1} (qz)^i \\
\tag{\ref{eq:E}}
E_j(z) = \sum_{i=0}^{\gamma-j-1} (qz)^i
\end{gather}
and where
\begin{gather}
\tag{\ref{eq:omega}}
\omega_m = {N \choose m} \big(1 - \mu/3\big)^{N-m} \big(\mu/3\big)^m \\
\tag{\ref{eq:xi}}
\xi_i = 1-(1-\mu)^i \\
\tag{\ref{eq:eta}}
\eta_i = 1-(1-\mu)^i\mu/3 \\
\tag{\ref{eq:zeta}}
\zeta_{i,m,n} = \sum_{r\geq1} {N-m \choose r}
(1-\mu)^{ri} \psi_{i,m,n,r} \bigg/ {N \choose n} \\
\notag
\psi_{i,m,n,r} = \sum_{q \geq 0}{m \choose q}{N-m-r \choose n-r-q}
\xi_i^{n-r-q}.
\end{gather}

\section{Computing the probabilities}
\label{sec:compute}

\subsection{Coefficient extraction}
\label{sec:extract}

The purpose of constructing the weighted generating function $F(z) = a_0 +
a_1z + a_2z^2 + \ldots$ is to extract the coefficient $a_k$, which
represents the probability that a read of size $k$ does not contain a
good seed.

$F(z)$ is the entry of the matrix $M(z) \cdot (I-M(z))^{-1}$ at the first
row and the last column, associated with terminators $\downarrow_{/0}$ and
$|$. The expression of $M(z)$ is too complex to compute this function
directly. Instead, we return to the definition $M(z) \cdot (I-M(z))^{-1} =
M(z) + M(z)^2 + \ldots$ and observe that the terms $M(z)^{k+2},
M(z)^{k+3}, \ldots$ have no influence on the coefficients $a_0, a_1,
\ldots, a_k$.

Indeed, a read of size $k$ has at most $k+1$ segments\footnote{Each
segment contains a terminator, which has size 1, except the unique
terminator $|$, which has size 0.}. Since the entries of $M(z)^n$ are the
weighted generating functions of reads with exactly $n$ segments, $a_k$
cannot depend on $M(z)^{k+2}, M(z)^{k+3}$. More formally, we can prove by
induction that all the entries of $M(z)^k$ are divisible by $z^{k-1}$,
showing that the contribution of $M(z)^{k+2} + M(z)^{k+3} + \ldots$ to
$a_0 + a_1z + \ldots +a_kz^k$ is $0$.

So to extract the coefficients of $F(z)$ up to order $k$ we can compute
the matrix $M(z) + M(z)^2 + \ldots + M(z)^{k+1}$, isolate the entry at the
first row and the last column and then compute the terms of the Taylor
expansion up to order $k$.

But we can do better than that. Since we are only interested in the
coefficients up to order $k$, we can replace the infinite power series
of the terms $A_{m,n}(z)$ and $C_m(z)$ by their truncated versions where
we keep only the terms up to order $k$. Likewise, we perform all
algebraic operation on truncated polynomials of order $k$, \textit{i.e.},
we discard at every stage the coefficients of order $k+1$ or greater.

More formally, truncated polynomials are defined as polynomials whose
coeffients of degree greater than $k$ are 0. Multiplying two truncated
polynomials $P(z)$ and $Q(z)$ gives a truncated polynomial $R(z)$ whose
coefficients up to order $k$ are those of $P(z)Q(z)$. Using truncated
polynomials of order $k$, the entry of the matrix $M(z) + M(z)^2 + \ldots
+ M(z)^{k+1}$ at the first row and the last column directly provides the
coefficients of interest.

But we can do even better than that. A read with $n$ segments contains at
least $m=\lfloor{(n-1)/2}\rfloor$ errors. Indeed, two segments terminated
by $\uparrow$ cannot follow each other, so at least (approximately) half
of the segments must be terminated by $\downarrow$, which means that the
segment contains a read error. Since read errors typically have a small
probability of occurrence, all the coefficients of $M(z)^n$ rapidly
converge to $0$ as $n$ increases. So instead of computing the matrix $M(z)
+ M(z)^2 + \ldots + M(z)^{k+1}$, we can interrupt the summation after a
certain power of $M(z)$ because the terms become negligible.

The number of errors $X$ in a read of size $k$ has a Binomial
distribution $X \sim B(k,p)$. From\cite{arratia1989tutorial} we can bound
the probabilities of the tail with the expression

\begin{equation}
\label{eq:bound}
Pr(X \geq m) \leq \exp \left( (m-k)\log \frac{k-m}{k(1-p)} -m\log
\frac{m}{kp} \right).
\end{equation}

Using the formula above, we can thus bound the probability that a read has
$n$ segments or more. We compute the matrices $M(z) + M(z)^2 + \ldots
+M(z)^n$ where the weighted generating functions have been replaced by
truncated polynomials and we extract $F_n(z)$, the entry at the first
column and the last row. When the bound is lower than a set fraction
$\varepsilon$ of the coefficient of $z^k$, we stop the computations.
Typically $\varepsilon = 0.01$ so this method ensures that the
probabilities that a read of size $k$ has no good seeds are accurate to
within 1\%.

This approach is feasible when the matrix $M(z)$ is small, but recall that
$M(z)$ has $N+\gamma+1$ rows/columns. Some sequences can have millions of
duplicates in plant and animal genomes (\textit{i.e.} $N > 10^6$), which
makes the computations with this method unfeasible because even a few
matrix multiplications are prohibitive. For large $N$, another method is
clearly needed.

\subsection{Monte Carlo sampling}

The symbolic representation as MEM segments can be used for an efficient
method to sample reads. Instead of generating the nucleotides of the $N+1$
sequences one by one, one can generate a single sequence of MEM segments.
Since the number of segment does not depend on $N$, we can obtain a fast
Monte Carlo method to sample millions of reads and count the proportion
that contain a good seed.

The principle is to proceed in cycles of two steps. We first sample the
position of the next read error, which gives the position of the next
symbol $\downarrow$. The second step is to determine whether there is a
symbol $\uparrow$ before that. For this we sample the number of masks that
vanish before the symbol $\downarrow$. If they all vanish, the read
contains a good seed, provided the next read error is at a distance
greater than $\gamma$. Otherwise, the process is repeated until we
generate a good seed, or until the read has size $k$ or greater (in which
case it has no good seed).

The method is summarized by the algorithm below. It requires efficient
methods to generate random samples with a geometric distribution and a
binomial distribution. Sampling from a geometric distribution can be done
by computing the logarithm of a uniform $(0,1)$ random variable. Sampling
from a binomial distribution can be done by the method of
Kachitvichyanukul and Schmeiser\cite{kachitvichyanukul1988binomial}.

\begin{algorithm}[H]
\label{alg:mcmc}
\SetAlgoLined
\KwResult {Sample a read at random. Return 1 if the read contains a
good seed, otherwise return 0.}
  $\lambda \leftarrow k$ \Comment*[r]{Size remaining to generate.}
  $m \leftarrow 0$ \Comment*[r]{Current number of hard masks.}
  \While {$\lambda > 0$}{
    $i \leftarrow geom(p)-1$ \Comment*[r]{Error-free nucleotides.}
    \eIf {$i \geq \lambda$}{
      \eIf{$\lambda < \gamma$} {
        return 0\; }{
        $h \leftarrow binom(m,(1-\mu)^\lambda)$
            \Comment*[r]{Surviving hard masks.}
        return 1 if $h = 0$, otherwise return 0\; }
    }
    {
      $h \leftarrow binom(m,(1-\mu)^i)$
          \Comment*[r]{Surviving hard masks.}
      $s \leftarrow binom(N-m,(1-\mu)^i \mu/3)$
          \Comment*[r]{Surviving soft masks.}
      \eIf {$ i \geq \gamma$ and $h = 0$ and $s = 0$}{
        return 1\;}{
        $m \leftarrow s + binom(N-s, \mu/3)$\;
        $\lambda \leftarrow \lambda - (i+1)$\;}
    }
 }
\end{algorithm}

In practice, we use a slightly modified version of this algorithm, that
returns the position of the good seeds if they are generated. This allows
us to estimate the probabilities for \emph{all} read sizes up to $k$. We
typically run this Monte Carlo sampling scheme 10,000,000 times, so that
we can estimate probabilities as low as $\sim1$ in 1 million.

The running speed depends on the error rate of the instrument $p$, but not
on the number of off-target sequences $N$. For values of $p$ around 0.01,
the running time for 10,000,000 iterations is $1$-$10$ seconds on current
hardware.

\section{False positives}

So far we have designed a method to estimate the probability that a
sequencing read contains a good seed. We referred to this as outcome $i.$
in section~\ref{sec:seedheur}. However, we are interested in the
probability that a read is mapped at an incorrect location, which is
outcome $ii.$

Computing the probability of outcome $ii.$ is hard, but we can compute the
probability that the read is not mapped at all, which is outcome $iii.$
The probability of outcome $ii.$ is then found by subtraction.

\subsection{The probability of outcome $iii.$}

Outcome $iii.$ implies that the read contains no seed at all, neither
on-target, nor off-target. This requires that there is no exact match of
size $\geq \gamma$ for \emph{any} sequence. Indeed, if it were the case,
then there would a MEM seed of size $\geq \gamma$ for at least one of the
sequences.

Let $E_0$ denote the event that there is no on-target match of size $\geq
\gamma$, and $E_j$ denote the event that there is no match of size $\geq
\gamma$ for off-target sequence $j$ ($1 \leq j \leq N$). We are thus
interested in computing $P(E_0 \cap \ldots \cap E_N)$. Since we assume
that the off-target sequences have evolved through the same mutagenesis
process and independently of each other, we can write

\begin{eqnarray}
\notag
P(E_0 \cap \ldots \cap E_N) &=&
  P(E_0) \cdot P(E_1 \cap \ldots \cap E_N | E_0) \\
\label{eq:PE0E1}
  &=& P(E_0) \cdot P(E_1 | E_0)^N =
  P(E_0) \cdot \left( \frac{P(E_0 \cap E_1)}{P(E_0)} \right)^N.
\end{eqnarray}

Expression (\ref{eq:PE0E1}) shows that we can compute the probability of
outcome $iii.$ from two quantities, namely $P(E_0)$ and $P(E_0 \cap E_1)$.

\subsection{Computing $P(E_0)$}
\label{sec:cPE0}

$P(E_0)$ is straightforward to evaluate because it involves a single
sequence. Methods to do this efficiently have been explained in detail
elsewhere\cite{filion2018analytic}. Here we only give the recurrence
equation to compute $P(E_0)$. Assume that the probability of a sequencing
error $p$ is known, that $q = 1-p$, and that $a_k$ is the probability that
a read of size $k$ does not contain a match of size $\geq \gamma$ for the
target. Then the following equation can be used to compute $a_k$ by
recurrence:

\begin{equation}
\label{eq:PE0}
a_k = 
\begin{cases}
1            &\quad\text{if } k < \gamma, \\
1 -pq^\gamma &\quad\text{if } k = \gamma, \\
a_{k-1} -pq^\gamma \cdot a_{k-\gamma-1} &\quad\text{otherwise.}
\end{cases}
\end{equation}

Knowing the size of the read and the error rate of the sequencing
instrument, we can thus compute $P(E_0)$ using expression (\ref{eq:PE0}).

\subsection{Computing $P(E_0 \cap E_1)$}
\label{sec:cPE0E1}

The event $E_0 \cap E_1$ is that the read contains no match of size $\geq
\gamma$ for either the target or the first off-target sequence (here
``first'' refers to any fixed off-target sequence). We compute the
probability of this event with the strategy described in
section~\ref{sec:symbolic}: we first recode the reads using a different
alphabet and then we specify the transfer matrix of the reads that have no
match for any sequence. The powers of this matrix hold the weighted
generating function of interest, from which we extract the coefficients
that correspond to the terms $P(E_0 \cap E_1)$ for reads of different
sizes.

The reads are recoded as sequences of symbols from the following alphabet
$\mathcal{A}_* = \{\square, |, \downarrow_{/1}^-, \downarrow_{/2}^-,
\ldots, \downarrow_{/1}^+, \downarrow_{/2}^+, \ldots, \Downarrow\}$. The
symbols $\downarrow_{/j}^-$ signify that the nucleotide is a mismatch
against the off-target sequence, the symbols $\downarrow_{/j}^+$ signify
that it is a mismatch against the target, and the symbol $\Downarrow$
signifies that it is a mismatch against both. As before, every other
nucleotide is replaced by the symbol $\square$, and the terminator $|$
is appended at the end of the read. We again define reads as sequences of
segments, except that now the terminators are the symbols
$\downarrow_{/j}^-$, $\downarrow_{/j}^+$, $\Downarrow$ and $|$.

The number $j$ in the symbol $\downarrow_{/j}^-$ indicates the relative
position of the last mismatch against the target. Likewise, the number $j$
in the symbol $\downarrow_{/j}^+$ indicates the relative position of the
last mismatch against the off-target sequence. In other words, the index
$j$ is the size of the match for the sequence that does not have a
mismatch. The terminators thus capture the local state of the read. For
instance the symbol $\downarrow_{/7}^-$ indicates that the nucleotide is a
mismatch against the off-target sequence, that it is a match for the
target, and that the six previous nucleotides were also a match for the
target.

\begin{figure}[h]
\centering
\includegraphics[scale=0.88]{sketch_dual.pdf}
\caption{\textbf{Title.}
Caption.}
\label{fig:dual}
\end{figure}

We use the same error model as before: for every nucleotide, the read
differs from the target with a fixed probability $p$ and the sequences
differ from each other with a fixed probability $\mu$. With these
definitions, the probability of occurrence of the symbol $\square$ is $a =
(1-p)(1-\mu)$. Likewise, the collective probability of occurrence of the
symbols $\downarrow_{/j}^-$ is $b = p\mu/3$, that of the symbols 
$\downarrow_{/j}^+$ is $c = (1-p)\mu$, and the probability of occurrence
of the symbol $\Downarrow$ is $d = p(1-\mu/3)$. It is easy to verify that
$a+b+c+d=1$.

These definitions allow us to specify the weighted generating function of
the segments. For instance, the weighted generating function of the
first segment of Fig.~\ref{fig:dual} is $(az)bz$, and that of the second is
$(az)^5cz$.

To specify the transfer matrix, we must ensure that we eliminate
the concatenations of two segments that would create a match of size $\geq
\gamma$ for any of the two sequences. After defining
\begin{equation*}
\begin{gathered}
r_i^+(z) = (az)^icz, \\
r_i^-(z) = (az)^ibz, \\
R_i(z) = \big( 1 + az + \ldots + (az)^i \big)dz, \\
F_i(z) = 1 + az + \ldots + (az)^i,
\end{gathered}
\end{equation*}
we can verify that the expression of the transfer matrix $L(z)$ is

\begin{equation*}
\begin{blockarray}{ccccccccc}
   & \scriptstyle{\Downarrow} & \scriptstyle{\downarrow_{/1}^+} & 
    \ldots & \scriptstyle{\downarrow_{\gamma-1}^+} &
    \scriptstyle{\downarrow_{/1}^-} & \ldots &
    \scriptstyle{\downarrow_{/\gamma-1}^-} & \scriptstyle{|} \\
\begin{block}{c[cccccccc]}
\scriptstyle{\Downarrow} & R_{\gamma-1}(z)  & r_0^+(z) & \ldots &
    r_{\gamma-2}^+(z) & r_0^-(z) & \ldots & r_{\gamma-2}^-(z) &
    F_{\gamma-1}(z) \\
\scriptstyle{\downarrow_{/1}^+} & R_{\gamma-2}(z) & & & & & & &
    F_{\gamma-2}(z) \\
\scriptstyle{\downarrow_{/2}^+} & R_{\gamma-3}(z) & & & & & & &
    F_{\gamma-3}(z) \\
\vdots & \vdots & & A(z) & & & B(z) & & \vdots \\
\scriptstyle{\downarrow_{/\gamma-1}^+} & R_0(z) & & & & & & & F_0(z) \\
\scriptstyle{\downarrow_{/1}^-} & R_{\gamma-2}(z) & & & & & & &
    F_{\gamma-1}(z) \\
\scriptstyle{\downarrow_{/2}^+} & R_{\gamma-3}(z) & & & & & & &
    F_{\gamma-2}(z) \\
\vdots & \vdots & & C(z) & & & D(z) & & \vdots \\
\scriptstyle{\downarrow_{/\gamma-1}^+} & R_0(z) & & & & & & & F_0(z) \\
\scriptstyle{|} & 0 & 0 & \ldots & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}
where $A(z)$, $B(z)$, $C(z)$ and $D(z)$ are square matrices with
$\gamma-1$ rows/columns, that are defined as

\begin{equation*}
A(z) = 
\begin{blockarray}{cccccc}
   & \scriptstyle{\downarrow_{/1}^+} & \scriptstyle{\downarrow_{/2}^+} &
    \ldots & \scriptstyle{\downarrow_{/\gamma-2}^+} &
    \scriptstyle{\downarrow_{/\gamma-1}^+} \\
\begin{block}{c[ccccc]}
\scriptstyle{\downarrow_{/1}^+} & 0 & r_0^+(z) & \ldots &
    r_{\gamma-4}^+(z) & r_{\gamma-3}^+(z) \\
\scriptstyle{\downarrow_{/2}^+} & 0 & 0 & \ldots &
    r_{\gamma-5}^+(z) & r_{\gamma-4}^+(z) \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\scriptstyle{\downarrow_{/\gamma-2}^+} & 0 & 0 & \ldots & 0 & r_0^+(z) \\
\scriptstyle{\downarrow_{/\gamma-1}^+} & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}

\begin{equation*}
B(z) = 
\begin{blockarray}{cccccc}
   & \scriptstyle{\downarrow_{/1}^-} & \scriptstyle{\downarrow_{/2}^-} &
   \ldots & \scriptstyle{\downarrow_{/\gamma-2}^-} &
   \scriptstyle{\downarrow_{/\gamma-1}^-} \\
\begin{block}{c[ccccc]}
\scriptstyle{\downarrow_{/1}^+} & r_0^-(z) & r_1^-(z) & \ldots &
    r_{\gamma-3}^-(z) & r_{\gamma-2}^-(z) \\
\scriptstyle{\downarrow_{/2}^+} & r_0^-(z) & r_1^-(z) & \ldots &
    r_{\gamma-3}^-(z) & 0 \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\scriptstyle{\downarrow_{/\gamma-2}^+} & r_0^-(z) & r_1^-(z) &
    \ldots & 0 & 0 \\
\scriptstyle{\downarrow_{/\gamma-1}^+} & r_0^-(z) & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}

\begin{equation*}
C(z) = 
\begin{blockarray}{cccccc}
   & \scriptstyle{\downarrow_{/1}^+} & \scriptstyle{\downarrow_{/2}^+} &
    \ldots & \scriptstyle{\downarrow_{/\gamma-2}^+} &
    \scriptstyle{\downarrow_{/\gamma-1}^+} \\
\begin{block}{c[ccccc]}
\scriptstyle{\downarrow_{/1}^-} & r_0^+(z) & r_1^+(z) & \ldots &
    r_{\gamma-3}^+(z) & r_{\gamma-2}^+(z) \\
\scriptstyle{\downarrow_{/2}^-} & r_0^+(z) & r_1^+(z) & \ldots &
    r_{\gamma-3}^+(z) & 0 \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\scriptstyle{\downarrow_{/\gamma-2}^-} & r_0^+(z) & r_1^+(z) & \ldots &
    0 & 0 \\
\scriptstyle{\downarrow_{/\gamma-1}^-} & r_0^+(z) & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}

\begin{equation*}
D(z) = 
\begin{blockarray}{cccccc}
   & \scriptstyle{\downarrow_{/1}^-} & \scriptstyle{\downarrow_{/2}^-} &
    \ldots & \scriptstyle{\downarrow_{/\gamma-2}^-} &
    \scriptstyle{\downarrow_{/\gamma-1}^-} \\
\begin{block}{c[ccccc]}
\scriptstyle{\downarrow_{/1}^-} & 0 & r_0^-(z) & \ldots &
    r_{\gamma-4}^-(z) & r_{\gamma-3}^-(z) \\
\scriptstyle{\downarrow_{/2}^-} & 0 & 0 & \ldots &
    r_{\gamma-5}^-(z) & r_{\gamma-4}^-(z) \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\scriptstyle{\downarrow_{/\gamma-2}^-} & 0 & 0 & \ldots & 0 & r_0^-(z) \\
\scriptstyle{\downarrow_{/\gamma-1}^-} & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray}.
\end{equation*}

The matrix $L(z)$ is relatively sparse and most non-zero entries are
monomials. In addition, the number of rows/columns is approximately
$2\gamma$, so the dimension of the matrix rarely exceeds 50. This means
that computing the powers of $L(z)$ is not prohibitive and we do not need
to design a Monte Carlo sampling scheme as we did before.

That being said, we do not need to compute all the powers up to $k$
because the terms vanish and become negligible after a certain iteration.
Indeed, the terms of $L(z)^n$ are the weighted generating function of the
reads without a match for either sequences and that contain exactly $n$
segments. Of those $n$ segments, $n-1$ are terminated by a mismatch
against at least one of the sequences. Following the rationale of
section~\ref{sec:extract}, we define $\tilde{p}$ as the upper bound on the
probability of a mismatch $\tilde{p} = \max\{b,c,d\}$, and updating
formula (\ref{eq:bound}) we bound the probability that a read of size $k$
contains $m = n-1$ or more mismatches

\begin{equation*}
Pr(X \geq m) \leq \exp \left( (m-k)\log \frac{k-m}{k(1-\tilde{p})} -m\log
\frac{m}{k\tilde{p}} \right).
\end{equation*}

The strategy is the same as in section~\ref{sec:extract}: we compute the
sum of matrices $L(z) + L(z)^2 + \ldots + L(z)^n$ where the weighted
generating functions have been replaced by truncated polynomials and we
extract the entry at the first column and the last row. When the bound is
lower than a set fraction $\varepsilon$ of the coefficient of $z^k$, we
stop the computations.

\subsection{The probability of outcome $ii.$}

For given values of the parameters $p$, $\mu$ and $N$, we can now compute
the probability of a false positive, which is outcome $ii$.

For this, we compute the probability of outcome $i.$, either by computing
the powers of $M(z)$ or by Monte Carlo sampling as explained in
section~\ref{sec:compute}. Next we compute the probability of outcome
$iii.$ using expression (\ref{eq:PE0E1}), where the terms are computed as
explained in sections~\ref{sec:cPE0E1} and \ref{sec:cPE0}.

\section{Estimating the parameters}
\label{sec:est}

So far we have assumed that we know all the specifications of the problem,
including the error rate of the instrument $p$, the number of off-target
sequences $N$, and their rate of sequence divergence $\mu$ (or more
accurately the probability that they differ from the target at every
nucleotide).

In general, $p$ is known from the vendor of the sequencing instrument, but
$N$ and $\mu$ can change for every read and they are usually unknown. We
thus need an estimation algorithm that can run fast enough that it does
not slow the mapping process. The main idea here is to use the backward
search, which is the method used to find the MEM seeds in the first place.

The backward search is a method that was first proposed by Ferragina and
Mazini\cite{ferragina2000opportunistic} to identify exact matches between
a text query and an index. The query is extended one letter/nucleotide at
a time, and at each iteration the search returns the number of matches in
the index/genome. This is thus a very efficient method to identify MEM
seeds, because one can keep extending the query until there is no match in
the genome or until the end of the read.

What matters for us is that we can run the backward search at some
arbitrary position of the read. The $i$-th iteration returns the number
$N_i$ of matches in the genome for a query sequence of length $i$.
Ignoring sequencing errors and focusing only on the $N$ off-target
sequences, the probability that the count goes from $N_i$ to $N_{i+1}$ is

\begin{equation*}
P(N_{i+1}|N_i) = {N_i \choose N_{i+1}}
\mu^{N_i-N_{i+1}}(1-\mu)^{N_{i+1}}.
\end{equation*}

Assuming that there are $N = N_0$ duplicates and that we run $m$
iterations of the backward search, the probability of the sequence of
observations $(N_1, N_2, \ldots, N_m)$ is thus

\begin{equation*}
P(N_1, \ldots, N_m|N_0) = 
{N_0 \choose N_1 } \ldots {N_{m-1} \choose N_m }
\mu^{N_0-N_m}(1-\mu)^{N_1+\ldots+N_m}.
\end{equation*}

This is a familiar distribution in disguise: by defining $\Delta_i =
N_{i-1} - N_i$, the series of observations $(\Delta_1, \ldots, \Delta_m,
N_m)$ given $N_0$ has a multinomial distribution with parameters
$\big(\mu, \mu(1-\mu), \ldots, \mu(1-\mu)^{m-1}, (1-\mu)^m\big)$. More
specifically, the probability of observing the sequence $(\Delta_1,
\ldots, \Delta_m, N_m)$ given $N_0$ is

\begin{align*}
{N_0 \choose \Delta_1, \ldots, \Delta_m, N_m}
\mu^{\Delta_1}\big(\mu(1-\mu)\big)^{\Delta_2}
\ldots \big(\mu(1-\mu)^{m-1}\big)^{\Delta_m}(1-\mu)^{mN_m}.
\end{align*}

At that point, we need to bring back into the light the practical
consideration that short matches are uninformative. This implies that the
first iterations of the backward search say nothing about the number of
duplicates of the sequence. In practice, we are better off ignoring the
output of the first $n-1$ iterations by just discarding them. We observed
that running 30 iterations of the backward search and discarding the
output of the first 20 gives good results (\textit{i.e.} we set $n = 21$
and $m=30$).

Now, if $N_0, N_1, \ldots, N_{n-1}$ are missing, we obtain the probability
of observing $(\Delta_{n+1}, \ldots, \Delta_m, N_m)$ from the marginal
distribution as

\begin{align*}
\begin{split}
{N_0 \choose N_0-N_n, \Delta_{n+1}, \ldots, \Delta_m, N_m}
\big(1-(1-\mu)^n\big)^{N_0-N_n}
\big(\mu(1-\mu)^n\big)^{\Delta_{n+1}} \times
\ldots \\
\ldots \times
\big(\mu(1-\mu)^{m-1}\big)^{\Delta_m}(1-\mu)^{mN_m}.
\end{split}
\end{align*}

Now replacing again $\Delta_i$ by $N_{i-1}-N_i$, the log-likelihood can be
expressed as

\begin{equation}
\begin{split}
\ell = cst + \log \Gamma(N_0+1) - \log \Gamma(N_0-N_n-1)
+ (N_0-N_n) \log \big(1-(1-\mu)^n\big) + \\
(N_n-N_m) \log(\mu) + (nN_n + N_{n+1} + \ldots + N_m) \log(1-\mu).
\end{split}
\end{equation}

To estimate the parameters by maximum likelihood, we differentiate $\ell$
with respect to $N_0$ and with respect to $\mu$, then we equate both terms
to 0. We thus have to solve

\begin{gather}
\label{eq:psi}
\Psi(N_0+1)-\Psi(N_0-N_n+1) +
  \log\big( 1-(1-\mu)^n \big) = 0, \text{ and} \\
\label{eq:N0}
\frac{N_n-N_m}{\mu}
-\frac{nN_n + N_{n+1}+\ldots+N_m}{1-\mu} +
(N_0-N_n)\frac{n(1-\mu)^{n-1}}{1-(1-\mu)^n} = 0.
\end{gather}

In the expressions above, $\Psi(\cdot)$ is the digamma function, defined
as the derivative of $\log \Gamma(\cdot)$. We observe that equation
(\ref{eq:N0}) gives an expression of $N_0$ as a function of $\mu$ and of
the observations, namely

\begin{equation*}
N_0 = N_n +
\left(\frac{nN_n+N_{n+1}+\ldots+N_m}{1-\mu}-
\frac{N_n-N_m}{\mu} \right)
\frac{1-(1-\mu)^n}{n(1-\mu)^{n-1}}.
\end{equation*}

To find the maximum likelihood estimate of $\mu$, we can substitute this
expression in equation (\ref{eq:psi}), obtaining an equation depending on
$\mu$ and on the observations, which can be solved numerically. One can
then substitute the solution into equation (\ref{eq:N0}) to obtain the
estimate of the number of duplicate sequences $N_0$.

However, the estimations of $\mu$ and $N_0$ show large deviations from the
target values on simulated data. In practice, one does not need to know
the parameters with exquisite precision, so to avoid unrealistic
estimates, we pick the best $\mu$ from a short list of candidate values
(\textit{e.g.}, 0.02, 0.04 and 0.06). To do this, we choose a value
$\mu^*$ from the candidate set, we compute $N_0(\mu^*)$ from equation
(\ref{eq:N0}) and we evaluate $\ell$ with these values. Finally, we select
the $\mu$ that gives the higest value of $\ell$, and the estimate for
$N$ is set to $N_0(\mu)$.

Finding the estimates of $N$ and $\mu$ can be performed in negligible time
compared to running the backward search. We can repeat this method at
different areas of the read and finally average the estimates to get
reasonable values of $N$ and $\mu$ for the whole read. With this
information, we can compute the probability that the read has no good seed
using the methods presented above.

\section{Conclusion}

This is awesome.

\section*{Acknowledgements}

We acknowledge the financial support of the Spanish Ministry of Economy
and Competitiveness (‘Centro de Excelencia Severo Ochoa 2013-2017’, Plan
Nacional BFU2012-37168), of the CERCA Programme~/~Generalitat de
Catalunya, and of the European Research Council (Synergy Grant 609989).

%---------------------------------------------------------------
%---------------------------------------------------------------

\bibliography{references,pubmed}
\bibliographystyle{plain}

%----------------------------------------------------------------

\end{document}

%gs -dNoOutputFonts -sDEVICE=pdfwrite -o out.pdf latex.pdf