### 2019-02-15 MEM.tex
Optimizing the figures and the text.

\documentclass{article}

\usepackage{algorithm2e}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{authblk}
\usepackage[english]{babel}
\usepackage{blkarray}
\usepackage[font=small]{caption}
\usepackage{cite}
\usepackage{graphicx}

\setlength{\thickmuskip}{2mu plus 3mu minus 1mu}
\setlength{\medmuskip}{1mu plus 2mu minus 1mu}

\SetKwComment{Comment}{$\triangleright$\ }{}

% ---- Author affiliations ---- %

\renewcommand\Affilfont{\itshape\small}

% ---- Propositions, lemmas, defintions... ---- %

%\newtheorem{algorithm}{Algorithm}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
%\newtheorem{example}{Example}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
%\newtheorem{remark}{Remark}

% ---- Special environments (examples and remarks) ---- %

\newcounter{examplecounter}
\newenvironment{example}
{\small\vspace{0.5\baselineskip}
  \refstepcounter{examplecounter}%
  \noindent\textbf{Example \arabic{examplecounter}.}%
}{\vspace{-0.2\baselineskip}\begin{center}%
  $\star$\end{center}\vspace{0.5\baselineskip}}

\newcounter{remarkcounter}
\newenvironment{remark}
{\small\it\vspace{0.5\baselineskip}
  \refstepcounter{remarkcounter}%
  \noindent\textbf{Remark \arabic{remarkcounter}.}%
}{\vspace{0.5\baselineskip}}

\newenvironment{inset}
{\vspace{0.5\baselineskip}\begin{center}}
{\end{center}\vspace{0.5\baselineskip}}

% ---- Macros ---- %

\newcommand{\dn}[1]{\scriptstyle{\downarrow_{/#1}}}
\newcommand{\up}[1]{\scriptstyle{\uparrow_{/#1}}}
\newcommand{\nd}{\scriptstyle{|}}

%---------------------------------------------------------------

\title{Computing MEM seeding probabilities}

\author[1,2]{Guillaume J. Filion}
\author[1,2]{Eduard Zorita}
\affil[1]{Genome Architecture, Gene Regulation, Stem Cells and Cancer
Programme, Center for Genomic Regulation (CRG), The Barcelona Institute of
Science and Technology, Dr. Aiguader 88, Barcelona 08003, Spain.}
\affil[2]{University Pompeu Fabra, Doctor Aiguader, 08003 Barcelona,
Spain.}

\date{\today}

%---------------------------------------------------------------
%---------------------------------------------------------------

\begin{document}

\maketitle

\begin{abstract}
High throughput sequencing produces massive amounts of data that must be
processed efficiently. Most reads are identified by mapping them to a
reference, typically a genome. Modern mappers rely on heuristics to gain
speed at a reasonable cost of accuracy. MEM seeding (where MEM stands for
Maximal Exact Match) is a popular heuristic that shows good empirical
performance but for which no theory is available. It is thus difficult to
calibrate or optimize mapping strategies based on MEM seeding. Here we
develop a theory to predict the performance of MEM seeding. More
specifically, we design a method to compute the probability a read is
mapped to the right location when using MEM seeds. Our computational
strategy is based on the analytic combinatorics programme, but uses only
part of it. By recoding the reads in a specialized combinatorial alphabet,
we can implicitly specify their generating function and extract the
probabilities of interest by direct methods or by Monte Carlo sampling.
This yields an efficient algorithm to calibrate the MEM seeding heuristic
for short read technologies.
\end{abstract}

%---------------------------------------------------------------
%---------------------------------------------------------------

\section{Introduction}

\subsection{Mapping reads to genomes}

Suppose that a DNA fragment drawn at random from a known
genome\footnote{We will assume that the sequence of the genome is known
exactly and with absolute certainty. This is not the case in practical
application, but this assumption will allow us to develop a theory
centered on the read.} is sequenced with an imperfect instrument; how can
we identify the location of the fragment in the genome?

We will refer to this question as the \emph{true location problem}.
It has applications in countless domains, like discovering
disease-causing mutations; diagnosing viral, bacterial or fungal
infections; diagnosing cancer and choosing appropriate therapies;
selecting breeds of interest in agriculture; tracing human migrations from
ancient bones; identifying the donor of a biological sample; diagnosing
genetic diseaes at birth; finding compatible graft donors \textit{etc}.

The focus of this document is not to solve the true location problem
\textit{per se}, but rather to evaluate whether a given solution is
correct. At first sight, whether we can map a read to its true location
seems to depend only on its length and on the error rate of the
sequencer. However, there is another important factor to consider: most
genomes contain repeated sequences, \textit{i.e.} relatively large
subsequences that are present at more than one location. For instance, we
cannot map the fragment with certainty if there exists an exact copy of
the sequence somewhere else in the genome, because we cannot know which of
the two copies was sequenced.

This case is rare though. Repeated sequences are in general not identical,
but merely homologous\footnote{By that we mean that drawing nucleotides
uniformly at random is very unlikely to produce sequences with the
observed degree of identity, meaning that the similarity is due some
biological duplication process.}. So when the DNA fragment is duplicated,
we can still map it, as long as we can distinguish it from its duplicates.
This in turn depends on their similarity with the target fragment and on
the error rate of the sequencer.

Because sequencing errors can convert a DNA fragment into one of its
duplicates, the \emph{true} location of a DNA fragment is not always the
\emph{best} (as measured by the identity between the sequence and the
candidate location). This leads to the following question: how can we
identify the \emph{best} location of the fragment in the genome?

We will refer to this task as the \emph{best location problem}. It amounts
to finding the optimal alignment between two sequences, and for this
reason has received substantial attention in bioinformatics. Finding the
best location of the read is the only way to get to its true location, so
we need to focus our attention on the best location problem.

\subsection{Seeding heuristics}
\label{sec:seedheur}

There exists exact algorithms to solve the best location
problem\cite{pmid7265238,pmid5420325}, but they are too slow to process
the large amount of data generated by modern sequencers. Instead, one uses
heuristic methods, \textit{i.e.} algorithms that run fast, but may not
return the best location\cite{Waterman1984}. Heuristic mapping algorithms
thus have three possible outcomes: $i.$ the DNA fragment is mapped to the
best location, $ii.$ the DNA fragment is mapped to some other location,
$iii.$ the DNA fragment is not mapped at all.

The most popular heuristic to address the best location problem is to
first identify short regions of high similiarity between the read and the
genome, and then run an exact alignment algorithm around those regions.
The first step is usually called ``seeding'' and the regions of high
similarity are called ``seeds''. This is for instance the general strategy
of the popular alignment algorithm BLAST\cite{pmid2231712}.

If seeds can be discovered fast, we can quickly zoom in on a small set of
candidate locations and reduce the input size of the (slow) alignment
algorithm. The disadvantage is that the target location is not always in
the candidate set. In this case, it cannot be discovered and the outcome
is either $ii.$ or $iii.$

This imposes a trade-off between speed and accuracy because each candidate
location must be tested with a slow algorithm. If the seeding step is set
to produce many candidates, the target is likely to be among them and it
will be discovered. The downside is that this will take long. Conversely,
if the seeding step is set to produce few candidates, the process will
run faster but the target is less likely to be in the candidate list so it
is less likely to be discovered.

Importantly, the terms of this trade-off depend on the algorithm. Some
seeding methods are better than others\cite{pmid16533404,pmid20460430}, so
one can develop faster mapping algorithms at no cost for accuracy.

The seeding framework was developed for the best location problem, but it
also applies to the true location problem. Actually, the process is
exactly the same except that outcomes $i.$ and $ii.$ now refer to the true
location. Outcome $i.$, called a ``true positive'', is that the DNA
fragment is mapped to the \emph{true} location. Outcome $ii.$, called a
``false positive'' is that the DNA fragment is mapped to an incorrect
location. Outcome $iii.$, called a ``false negative'', does not change and
still means that the DNA fragment is not mapped at all.

We will make two assumptions that will simplify the discussion. The first
is that the true location is always the best. We will further comment on
this and explain how to relax this assumption in section
section~\ref{sec:not_the_best}. Until then, we consider that the true
location problem is the same as the best location problem. The second
assumption is that the best location is always discovered if it is in the
candidate set. This implies that outcomes $i.$, $ii.$ and $iii.$ are
determined by the seeding process. For this reason, we will sometimes
refer to them as the outcomes of the seeding step, even though they are
the outcomes of the whole mapping process.

\subsection{MEM seeds}

In general, seeds do not have to be exact matches\cite{pmid20460430}. For
instance, the algorithm PatternHunter\cite{pmid11934743} uses ``spaced
seeds'' that tolerate mismatches\footnote{For the anecdote, the method is
covered by a patent (reference US20070088510A1).}. However, in this
document, we will always assume that a seed is an exact match with no gap.
In other words, a seed is a perfect match between a part of the read and a
genomic site.

The rest of the document will focus on a seeding heuristic known as ``MEM
seeding'' (where MEM stands for Maximal Exact Match). This approach gives
good empirical results when mapping short reads (it is for instance the
strategy\footnote{To be exact, BWA-MEM uses SMEM seeds (Super Maximal
Exact Match seeds).} of the popular mapper BWA-MEM\cite{li2013aligning})
but there is presently no theory to justify its performance. A deeper
understanding of MEM seeding would be useful to calibrate the
speed/accuracy trade-off of this heurisitic.

Before developing a formal theory, let us explain in more detail what
Maximal Exact Matches are and how MEM seeding works.

\begin{definition}
A Maximal Exact Match (MEM) seed is a subsequence of a read that is
present in the genome of interest and that cannot be extended --- either
because the read ends or because the extended subsequence is not in the
genome.
\end{definition}

This definition presents a computational challenge: to know if a given
subsequence is a MEM seed we need to know if it exists somewhere in the
genome. This is a non trivial problem in itself, but fortunately there
exists practical methods to identify MEM seeds, even when the reference
genome is very large\cite{ferragina2000opportunistic,
ferragina2005indexing}. Such algorithms are crucial for the present
discussion, but describing them falls outside the scope of the document.
We refer the interested reader to the relevant
literature\cite{pmid24336412,pmid25399029,pmid23349213,pmid19389736,
li2013aligning}, and here on we will assume that MEM seeds are available
at all times. Looking ahead to section~\ref{sec:est}, let us just mention
that this can be done with an algorithm called the \emph{backward
search}\cite{ferragina2000opportunistic}.

Before going further, we need to move a practical consideration out of the
way. It stands to reason that very short matches between the read and the
genome are uninformative. For instance, $99.7\%$ of all the possible
12-mers are present in the reference human genome, so an exact match of
size 12 or lower says nothing about the location of the DNA fragment. In
what follows, we will assume that subsequences of the read can only match
the original DNA sequence and/or its duplicates, but not the rest of the
genome. This will greatly simplify the exposition of the theory. We will
discuss how to relax this assumption in section~\ref{sec:short_MEMs}.

Fig.~\ref{fig:MEM_example} shows a concrete example where a read is
aligned to candidate locations. The target (the true location of the
read) is first potential location. The locations below correspond to
duplicate sequences. The mismatches between the genome and the read are
highlighted in bold. The read has three errors indicated by stars, so
\texttt{T}, \texttt{T} and \texttt{C} should actually be \texttt{A},
\texttt{A} and \texttt{G} (\textit{i.e.} as in the target sequence).

\begin{figure}[h]
\centering
\includegraphics[scale=1]{MEM_example.pdf}
\caption{\textbf{Example of MEM seeding.}
A DNA fragment is sequenced with an imperfect instrument. It comes from
the top genomic location (Truth), but three errors were introduced
(indicated by stars). The sequence of the DNA fragment has three
duplicates at other genomic locations, shown below the target. Mismatches
between the read and the candidate sequence are highlighted in bold. MEM
seeds are indicated by grey boxes.}
\label{fig:MEM_example}
\end{figure}

The MEM seeds are indicated by grey boxes. Technically, they are part of
the read, but they are shown on the genome to highlight the position of
the match. One MEM seed cannot cover a smaller one (otherwise the smaller
would not be a MEM seed), but two consecutive MEM seeds can overlap, as in
the case of the frst two from the left. Note that overlapping MEM seeds
must match distinct sequences of the genome (otherwise neither of them
would be a MEM seed). Consecutive MEM seeds can also be disjoint, as in
the case of the second and the third from the left. In this case, they
must be separated by a nucleotide that is a mismatch against \emph{all}
the duplicates.

Observe that the rightmost MEM seed matches two genomic locations. This
case motivates the following definition, which will play an important role
for the development of the theory.

\begin{definition}
A \emph{strict} MEM seed has a single genomic match.
A \emph{shared} MEM seed has several genomic matches.
\end{definition}

Without surprise, the principle of MEM seeding is to use the genomic
matches of MEM seeds as candidate locations. Importantly, MEM seeds
shorter than a certain threshold $\gamma$ are ignored (in line with the
observation that short matches are uninformative). The parameter $\gamma$
is the tuning knob on the speed/accuracy trade-off. In
Fig.~\ref{fig:MEM_example} for instance, if $\gamma = 7$, there are three
candidate locations (the first, the second and the fourth) and if $\gamma
= 8$ there are only two (the first and the fourth).

Not every MEM seed makes the target a candidate location, so we introduce
the following terms to simplify the discussion.

\begin{definition}
Assume that the thresohld $\gamma$ is fixed. A \emph{good} seed is an
on-target MEM seed of size $\gamma$ or greater. In other words, the true
location is in the candidate set if and only if the read contains a
good seed.
\end{definition}

Compared to seeds of fixed size, MEM seeds have at least two
counter-intuitive (and undesirable) properties. The first is that there
are cases where there cannot be any good seed, even when changing
$\gamma$. Fig.~\ref{fig:full_masking_example} shows such an example. Even
though there is a single sequencing error, the read has no good seed.
Lowering $\gamma$ does not change this, so there is no way to discover the
true location (even though it is the best location).

\begin{figure}[h]
\centering
\includegraphics[scale=1]{full_masking_example.pdf}
\caption{\textbf{Issues with MEM seeds: inaccessible targets.}
The read, the MEM seeds and the sequences are represented as in
Fig.~\ref{fig:MEM_example}. The MEM seeds matching the two incorrect
locations at the bottom effectively hide the target, so it cannot
be discovered. This can occur even when the true location is the
best and when there is a single error on the read.}
\label{fig:full_masking_example}
\end{figure}

The second counter-intuitive property of MEM seeding is that shortening
a read can sometimes create a good seed. Fig.~\ref{fig:short_vs_long}
shows an example of this case. Again, there is no good seed, but there
would be one if the read were two nucleotides shorter on the right side.
Indeed, in this case there would be a shared MEM seed matching the first
and the second locations, provided $\gamma \leq 12$.

\begin{figure}[h]
\centering
\includegraphics[scale=1]{short_vs_long_example.pdf}
\caption{\textbf{Issues with MEM seeds: too long reads.}
The read, the MEM seeds and the sequences are represented as in
Fig.~\ref{fig:MEM_example}. There would be a good seed (shared) if the
read were two nucleotides shorter. The true location is hidden by
the last two nucleotides.}
\label{fig:short_vs_long}
\end{figure}

These examples show that MEM seeds can perform worse than seeds of fixed
size. MEM seeds yield fewer candidates and therefore speed up the mapping
process, but the question is at what cost for accuracy? The theory
developed below will allow us to compute the probability that a read has
no good seed and thus that the true location is missed. This probability
is the key to evaluating and calibrating the MEM seeding heuristic.

We first expose a strategy to compute the probablity of true positives
(outcome $i.$), then show how to compute the probability of false
negatives (outcome $iii.$), and combine these two quantities to compute
the probability of false positives (outcome $ii.$). Finally, we provide a
computationally efficient method for estimating the parameters that are
necessary to estimate the probabilities of interest.

\section{Combinatorial construction}

In this section we represent reads as combinatorial ojects. The purpose of
this construction is to give a generative model of reads without good
seeds so that we can later compute their probability of occurrence.

\subsection{Hard and soft masking}

For each potential location including the target, we define the
\emph{match streak} at a given nucleotide as the number of nucleotides
since the last mismatch on the left, or since the beginning of the read.

\begin{definition}
At a given position of the read, an off-target sequence is a \emph{hard
mask} if its success streak is strictly longer than the match streak of
the target. An off-target sequence is a \emph{soft mask} if its has the
same match streak as the target.
\end{definition}

Fig.~\ref{fig:hard_vs_soft_masks} illustrates hard and soft masks with an
example. It is important to bear in mind that hard and soft masks depend
on the position of interest: a sequence can be a mask at the left end of
the read but not at the right end, or \textit{vice versa}. In the example
of Fig.~\ref{fig:full_masking_example}, every nucleotide of the read has a
hard mask, which is why the target cannot be discovered. In the example of
Fig.~\ref{fig:short_vs_long}, the middle sequence is a soft mask on the
right half of the read, but it turns into a hard mask two nucleotides
before the right end, which is why the target could be discovered if the
read were shorter.

\begin{figure}[h]
\centering
\includegraphics[scale=0.85]{hard_vs_soft_masks.pdf}
\caption{\textbf{Example of hard and soft masks.}
The nucleotides that do not match the read are reprsented as black boxes.
The nucleotides that contribute to the match streak are represented as
grey boxes. The remaining nucleotides are represented as open boxes. At
the given position, the match streak of the target is 7. The first
off-target is a hard mask because its match streak is $13>7$. The second
off-target sequence is a soft mask because its match streak is 7 like the
target. The third off-target sequence is not a mask because its match
streak is $2<7$.}
\label{fig:hard_vs_soft_masks}
\end{figure}

From the definition, we see that the last nucleotide of every strict
on-target MEM seed is always unmasked. Conversely, an unmasked nucleotide
always belongs to exactly one strict on-target MEM seed. Also, the last
nucleotide of every shared on-target MEM seed is always soft-masked, but a
soft-masked nucleotide does not always belong to a shared on-target MEM
seed.

Since hard and soft masks inform us about the positions of on-target
MEM seeds, we construct an alphabet that encodes the masking status of the
nucleotides.

\subsection{The MEM alphabet}

We recode the reads as sequences of letters from the so-called MEM
alphabet $\mathcal{A} = \{\square, |, \uparrow_{/1}, \uparrow_{/2},
\uparrow_{/3} \ldots, \downarrow_{/0}, \downarrow_{/1}, \downarrow_{/2},
\ldots\}$.

The symbols $\downarrow_{/m}$ indicate that the nucleotide is a sequencing
error and $m \geq 0$ is the number of off-target locations that
\emph{match} the nucleotide. Since a sequencing error is always a mismatch
against the target, the symbol $\downarrow_{/0}$ indicates that the
ncleotide is a mismatch against \emph{every} candidate.

The symbols $\uparrow_{/j}$ indicate a change in masking status: the
nucleotide is not masked but its left neighbor is. The index $j \geq 1$ is
the number of nucleotides since the last mismatch or since the beginning
of the read.

All the other nucleotides are represented by the $\square$ symbol.
Finally, the $|$ symbol is appended after the last nucleotide of the read.
It is an obligatory terminator that does not correspond to an actual
nucleotide.

Note that the numbers decorating $\uparrow$ and $\downarrow$ are different
in nature. In the case of the symbol $\downarrow$ it is an information
about the number of sequences matching the read; in the case of the symbol
$\uparrow$ it is the position of the previous sequencing error.
Fig.~\ref{fig:sketch_extended} shows the encoding of the read from
Fig.~\ref{fig:MEM_example} in the MEM alphabet.

\begin{figure}[h]
\centering
\includegraphics[scale=.85]{sketch_extended.pdf}
\caption{\textbf{The MEM encoding}.
The read of Fig.~\ref{fig:MEM_example} is represented in the MEM alphabet.
The arrows departing from those numbers help understand their meaning. The
$\downarrow$ symbol is decorated by the number of locations that match the
nucleotide. The $\uparrow$ symbol is decorated by the number of
nucleotides from the last error or from the beginning of the read. Also
note the presence of the terminator $|$ at the end of the read.}
\label{fig:sketch_extended}
\end{figure}

The MEM alphabet captures the masking status of the nucleotide as follows:
If we denote the number of off-targets as $N$, the symbol
$\downarrow_{/m}$ indicates that the nucleotide has $m$ hard masks and
$N-m$ soft masks. In addition, the symbols $\uparrow_{/j}$ indicate that
the nucleotide is unmasked.

In the MEM alphabet, strict on-target MEM seeds are the longest stretches
of symbols containing $\uparrow$ and not containing $\downarrow$. Indeed,
such a stretch matches the target, it matches the target only because it
contains at least one unmasked nucleotide (marked by $\uparrow$), and it
cannot be extended because it is flanked by sequencing errors or by the
ends of the reads. Note that there is exactly one symbol $\uparrow$ per
strict on-target MEM seed, and therefore that two symbols $\uparrow$ must
be separated by at least one symbol $\downarrow$.

Shared on-target MEM seeds are the longest stretches of $\square$ symbols
flanked by $\downarrow_{/0}$, or by the ends of the read. Indeed, such a
stretch is a MEM seed because it matches the target and it cannot be
extended ($\downarrow_{/0}$ is a mismatch against every sequence). Also,
it cannot be a strict on-target MEM seed because it does not contain the
$\uparrow$ symbol, so it must be shared.

\subsection{Reads as extended MEM segments}

Reads can be viewed as sequences of symbols, or as sequences of
\emph{segments}. For the purpose of constructing reads without good seeds,
the second point of view turns out to be essential.

\begin{definition}
\label{def:segment}
A segment is a maximal sequence of $\square$ symbols flanked on the right
side by one of the symbols $\uparrow_{/j}$, $\downarrow_{/m}$, or $|$. The
one and only segment flanked by $|$ is called the tail. It is the only
segment that can have size 0.
\end{definition}

In line with definition~\ref{def:segment}, the symbols $\uparrow_{/j}$,
$\downarrow_{/m}$ and $|$ are referrred to as \emph{terminators}.
Fig.~\ref{fig:sketch_segment} shows the segment representation of the read
from Fig.~\ref{fig:sketch_extended}.

\begin{figure}[h]
\centering
\includegraphics[scale=.85]{sketch_segments.pdf}
\caption{\textbf{The extended MEM segment encoding}.
The read from Fig.~\ref{fig:sketch_extended} is represented as a sequence
of segments. Each segment is terminated by either $\uparrow_{/j}$, or
$\downarrow_{/m}$ or $|$. The rightmost fragment is the tail. The numbers
below indicate the size of each segment in number of nucleotides.}
\label{fig:sketch_segment}
\end{figure}

We now have a representation of reads as sequences of segments. Only some
of these sequences are possible. For instance, a segment terminated by
$\uparrow$ cannot follow another. The task, however, is not to give a
construction of reads, but a construction of reads without good seeds. For
this, we need to introduce a new set of tools.

\section{Computational strategy}
\label{sec:symbolic}

We turn to the question of translating the extended MEM segment
representation to a construction of reads without good seeds. For this, we
use tools borrowed from analytic combinatorics. We give the minimum amount
of background that is necessary to expose the theory. A more detailed
account of the strategy to compute seeding probabilities with analytic
combinatorics has been published
elsewhere\cite{filion2017analytic,filion2018analytic}. We refer the
readers interested in learning more about analytic combinatorics to the
textbooks\cite{flajolet2009analytic,sedgewick2013introduction}.

\subsection{Weighted generating functions}

The central object of analytic combinatorics is the \emph{generating
function}.

\begin{definition}
\label{def:GF}
Let $\mathcal{A}$ be a set of combinatorial objects characterized by a
size in $\mathbb{N}$ and a weight in $\mathbb{R}^+$. The weighted
generating function of $\mathcal{A}$ is defined as

\begin{equation}
\label{eq:GF1}
A(z) = \sum_{a \in \mathcal{A}} w(a) z^{|a|},
\end{equation}
where $|a|$ and $w(a)$ denote the size and weight of the object $a$,
respectively. Expression (\ref{eq:GF1}) also defines a sequence $(a_k)_{k
\geq 0}$ such that

\begin{equation*}
A(z) = \sum_{k=0}^\infty a_k z^k.
\end{equation*}

By definition $a_k = \sum_{a \in A_k}w(a)$, where $A_k$ is the class of
objects of size $k$ in $\mathcal{A}$. The number $a_k$ is the
total weight of objects of size $k$.
\end{definition}

To give a concrete example, if the probability of the symbol
$\downarrow_{/0}$ is $q$, the weighted generating function of a segment
containing only this symbol is $qz$. In our case, we are interested in the
weighted generating function of reads without good seeds, so that the
coefficient $a_k$ is equal to the probability that a read of size $k$ has
no good seed.

The motivation for introducing weighted generating functions is that
operations on combinatorial objects translate into operations on their
weighted generating functions. If $A(z)$ and $B(z)$ are the weighted
generating functions of two mutually exclusive sets $\mathcal{A}$ and
$\mathcal{B}$, the weighted generating function of $\mathcal{A} \cup
\mathcal{B}$ is $A(z) + B(z)$, as evident from expression (\ref{eq:GF1}).
Size and weight can be defined for pairs of objects in $\mathcal{A} \times
\mathcal{B}$ as $|(a,b)| = |a| + |b|$ and $w(a,b) = w(a)w(b)$. In other
words the sizes are added and the weights are multiplied.  With this
convention, the weighted generating function of the Cartesian product
$\mathcal{A} \times \mathcal{B}$ is $A(z)B(z)$. This simply follows from
expression (\ref{eq:GF1}) and from

\begin{equation*}
A(z)B(z) =
\sum_{a\in \mathcal{A}}w(a)z^{|a|} \sum_{b\in \mathcal{B}}w(b)z^{|b|}
= \sum_{(a,b) \in \mathcal{A} \times \mathcal{B}} w(a)w(b)z^{|a|+|b|}.
\end{equation*}

These two operations are all we need in order to compute the weighted
generating functions of the reads of interest. Addition corresponds to
creating a new family by merging reads from families $\mathcal{A}$ and
$\mathcal{B}$. Multiplication corresponds to creating a new family by
concatenating reads from families $\mathcal{A}$ and $\mathcal{B}$.

\subsection{Analytic representation}

The symbolic method is a strategy belonging to the general \emph{analytic
combinatorics} framework. The idea is to combine simple combinatorial
objects into more complex objects. Each combinatorial operation on the
objects corresponds to a mathematical operation on their weighted
generating functions. One can thus obtain the weighted generating function
of complex objects, and from there extract their probability of
occurrence.

The combinatorial construction of reads using the extended MEM alphabet
gives us a method to express their weighted generating function from the
simpler generating function of segments. More specifically, we can
describe complex sequences of segments through a so-called \emph{transfer
matrix} $M(z)$ that specifies which pairs of consecutive segments are
allowed and which are forbidden in the reads of interest.

The segment types are identified by their terminators arranged in a
predefined oder. The entry of $M(z)$ at coordinates $(i,j)$ is the
weighted generating function of segments with the $j$-th terminator that
can be inserted immediately after segments with the $i$-th terminator. The
entries of the matrix $M(z) + M(z)^2 + \ldots = M(z) \cdot (I-M(z))^{-1}$
are the weighted generating functions of the reads of all sizes using all
the allowed combinations of segments.

The entry of interest from $M(z)$ is the one that corresponds to the
terminators $\downarrow_{/0}$ and $|$ because it represents the reads of
any size whose first segment can be inserted after $\downarrow_{/0}$, and
whose last segment is terminated by $|$. Indeed, the $|$ terminator marks
the end of the read, so it must be the final symbol of every valid read.
As for the first segment, observe that all the segments that can follow
the $\downarrow_{/0}$ symbol are exactly those that can be at the frst
postion of the read\footnote{One could have introduced a special start
symbol. It is equivalent to an invisible $\downarrow_{/0}$ prepending
the reads.}.

In summary, the transfer matrix $M(z)$ contains the weighted generating
function of interest. Our next task is thus to specify $M(z)$.

\section{The transfer matrix}
\label{sec:transfer_mat}

\subsection{Error model and divergence of the duplicates}

We assume that the target sequence has $N$ duplicates, so that there are
$N$ off-target sequences. We further assume that duplication was
instantaenous and that all $N+1$ sequences diverge independently of each
other at a constant rate. In other words, we ignore the complications due
to the genealogy of the duplication events and we simply assume that at
each position, any given duplicate is identical to the target with
probability $1-\mu$. If it is not, we assume that the duplicate has any of
the remaining three nucleotides with equal probability (\textit{i.e.} each
is found with probability $\mu/3$).

We also assume that the sequencing instrument has a constant substitution
rate $p$, and that insertions and deletions never occur. In case of a
substitution, we again assume that the remaining three nucleotides are
equally likely.

Read errors are always mismatches against the target (because we assume
that the target is the true sequence), and they match each off-target
sequence with probability $\mu/3$. Correct nucleotides are always a match
for the target, and they match each off-target sequence with probability
$1-\mu$.

We are now ready to describe the transfer matrix $M(z)$. We do this entry
by entry, in order of increasing complexity.

\subsection{Segments following $\uparrow_{/j}$}

A segment terminated by $\uparrow_{/j}$ is the beginning of an on-target
MEM seed of size at least $j$. The seed reaches the next sequencing
error or the end of the read, so the number of $\square$ symbols in the
next segment must be at most $\gamma-j-1$ and it must be terminated by a
$\downarrow$ symbol or by the tail terminator $|$.

On this segment, the matches between the read and the off-target sequences
are irrelevant. The weighted genearting function of each $\square$ symbol
is simply $qz$ ($q$ is the probability that the nucleotide is not an error
and $z$ marks objects of size $1$). The following definition will simplify
the notations.

\begin{definition}
The probability that a symbol is $\downarrow_{/m}$ given that the
nucleotide is a read error is
\begin{equation}
\label{eq:omega}
\omega_m = {N \choose m} \big(1 - \mu/3\big)^{N-m} \big(\mu/3\big)^m.
\end{equation}
\end{definition}

The weighted generating function of the $\downarrow_{/m}$ terminator is
thus $\omega_m pz$ (the probability of the symbol $\downarrow_{/m}$ is
$\omega_m p$ and $z$ marks objects of size $1$).

The entries of the transfer matrix that correspond to transitions from
$\uparrow_{/j}$ to other segments are now easy to find. The weighted
generating function of the segments terminated by $\downarrow_{/m}$
following a segment terminated by $\uparrow_{/j}$ is
\begin{equation}
\label{eq:D}
D_{j,m}(z) = \omega_m pz \sum_{i=0}^{\gamma-j-1} (qz)^i.
\end{equation}

And the weighted generating function of the tail segments following
segments terminated by $\uparrow_{/j}$ is
\begin{equation}
\label{eq:E}
E_j(z) = \sum_{i=0}^{\gamma-j-1} (qz)^i.
\end{equation}

\subsection{Segments following $\downarrow_{/m}$}

The symbol $\downarrow_{/m}$ signifies that the nucleotide has $m$ hard
masks and $N-m$ soft masks. If all the masks disappear before the first
read error, the next terminator will be a $\uparrow$ symbol, otherwise it
will be a $\downarrow$ symbol. We separate the cases based on the
terminator of the segment.

\subsubsection*{Case 1: the terminator is $\uparrow_{/i}$}

\begin{definition}
The probability that a given off-target sequence contains a mismatch in a
sequence of $i$ error-free nucleotides is
\begin{equation}
\label{eq:xi}
\xi_i = 1-(1-\mu)^i.
\end{equation}
\end{definition}

With this notation, the probability that at least one mask survives a
sequence of $i$ error-free nucleotides is thus $1-\xi_i^N$, and the
probability that there remains a mask at the $i-1$-th but not at the 
$i$-th error-free nucleotide is $\xi_i^N - \xi_{i-1}^N$. From this we
conclude that the weighted generating function of the segments terminated
by $\uparrow_{/i}$ following a segment terminated by $\downarrow_{/m}$ is
\begin{equation}
\label{eq:B}
B_i(z) = \Big( \xi_i^N-\xi_{i-1}^N \Big) (qz)^i.
\end{equation}

In the equation above, $q^i$ is the probability that there is no
sequencing error in $i$ nucleotides and $z^i$ marks objects of size $i$.
The fact that the reads have no good seeds imposes $i < \gamma$. Also note
that this expression is the same for all $\downarrow$ symbols; it does not
depend on $m$.

\subsubsection*{Case 2a: the terminator $|$ comes before the $\gamma$-th
nucleotide}

In this case there can be no good seed because the read terminates too
early. We must just enforce the condition that at least one of the $N$
masks survives until the end of the segment in order to exclude the
$\uparrow$ symbol. The weighted generating function is
\begin{equation*}
\sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i.
\end{equation*}

\subsubsection*{Case 2b: the terminator $|$ comes after the $\gamma$-th
nucleotide}

In this case, the soft masks do not hide the target: even if one of them
survives until the end of the read, there will be a good seed (shared in
this case). To make sure that the read has no good seed, we must enforce
the condition that at least one hard mask survives until the end of the
segment (which is impossible if $m = 0$). The weighted generating function
is
\begin{equation*}
\sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \Big) (qz)^i.
\end{equation*}

Finally, the weighted generating function of the tail segments following
segments terminated by $\downarrow_{/m}$ is
\begin{equation}
\label{eq:C}
C_m(z) =
\sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i +
  \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \Big) (qz)^i.
\end{equation}

\subsubsection*{Case 3a: the terminator $\downarrow_{/n}$ comes before the
$\gamma$-th nucleotide}

In this case, there can be no good seed and we must only exclude the
$\uparrow$ terminator. The weighted generating function is
\begin{equation}
\omega_n pz \sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i.
\end{equation}

\subsubsection*{Case 3b: the terminator $\downarrow_{/n}$ comes after the
$\gamma$-th nucleotide}

This case is by far the most convoluted. Since the segment contains at
least $\gamma$ error-free nucleotides, we must ensure that it does not
contain a good seed. This will be the case if any of the two following
conditions is validated: $i)$ at least one hard mask covers all the
error-free nucleotides, or $ii)$ all the hard masks vanish but at least
one soft mask covers the whole segment (including the terminator).

The two conditions are mutually exclusive by construction. They are
graphically represented in the diagram below. The left panel corresponds
to case $i)$ and the panel to case $ii)$. The top row represents the
target, and the bottom rows represent off-target sequences (using the same
symbols as in Fig.~\ref{fig:MEM_example}).
\begin{inset}
\includegraphics{masks.pdf}
\end{inset}

Whenever a hard mask (here the first off-target sequence) covers the
nucleotides as shown in the left panel, there can be no good seed,
irrespective of the value of $\gamma$. The positions marked with a
question mark are irrelevant, they cannot change the fact that there is no
on-target MEM seed. If the hard masks vanish, as in the right panel, then
it all depends on the soft masks. If a soft mask covers the whole segment,
then there can be no good seed, irrespective of the value of $\gamma$.

Condition $i)$ has probability $\big(1 - \xi_i^m \big)$ and the
weighted generating function is thus
\begin{equation*}
\omega_n pz \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \Big) (qz)^i.
\end{equation*}

Condition $ii)$ is more convoluted, so we introduce some further
notations to solve this sub-case.
\begin{definition}
The probability that a given off-target sequence contains a mismatch in a
sequence of $i$ error-free nucleotides followed by an error is
\begin{equation}
\label{eq:eta}
\eta_i = 1-(1-\mu)^i\mu/3.
\end{equation}
\end{definition}

The probability of condition $ii)$ is
\begin{equation*}
\xi_i^m \Big(1 - \eta_i^{N-m} \Big),
\end{equation*}
but we need to break up this term among all the terminators
$\downarrow_{/n}$, ($0 \leq n \leq N$). For this, we decompose the sum on
the number of soft masks that run to the end of the segment, including the
terminator. The probability that there are $r \geq 1$ such soft masks is
\begin{equation*}
{N-m \choose r} (1 - \eta_i)^r \eta_i^{N-m-r}.
\end{equation*}

Each of them matches the nucleotide at the end of the segment, so the
total number of matches is $r$ plus the number of sequences that also
match the last nucleotide, among the remaining $N-m-r$ soft masks and $m$
hard masks.

Let us start with the $N-m-r$ that vanished somewhere in the segment. The
conditional probability that they match the last nucleotide given that
they vanish somewhere in the segment is $\mu/3 \cdot \xi_i / \eta_i$. The
conditional probability that they do not match it is $(1-\mu/3) / \eta_i$.
For the $m$ hard masks, the probability that they match the last
nucleotide given that they vanish somewhere in the segment is simply
$\mu/3$, and the probability that they do not match it is $1-\mu/3$.

The probability that the last nucleotide of the segment matches $n-r$ of
the $N-r$ masks that vanished before the terminator is
\begin{equation*}
\frac{(\mu/3)^{n-r}(1-\mu/3)^{N-n}}{\eta_i^{N-m-r}}
\psi_{i,m,n,r}\text{, where \;}
\psi_{i,m,n,r} = \sum_{q \geq 0}{m \choose q}{N-m-r \choose n-r-q}
\xi_i^{n-r-q}.
\end{equation*}

Finally, the probability that a total of $n$ off-target sequences match
the terminator and thus that the terminator is the symbol
$\downarrow_{/n}$ is
\begin{eqnarray*}
&\;& \sum_{r\geq1} {N-m \choose r}
(1 - \eta_i)^r (\mu/3)^{n-r} (1-\mu/3)^{N-n} \psi_{i,m,n,r} \\
&=& (\mu/3)^n(1-\mu/3)^{N-n} \sum_{r\geq1} {N-m \choose r}
  (1 - \mu)^{ri} \psi_{i,m,n,r} \\
&=& \omega_n \cdot \zeta_{i,m,n},
\end{eqnarray*}
where
\begin{equation}
\label{eq:zeta}
\zeta_{i,m,n} = \sum_{r\geq1} {N-m \choose r}
(1-\mu)^{ri} \psi_{i,m,n,r} \bigg/ {N \choose n}.
\end{equation}

The weighted generating function of the segments terminated by
$\downarrow_{/n}$ following segments terminated by $\downarrow_{/m}$ is
\begin{equation}
\label{eq:A}
A_{m,n} =
\omega_n pz \sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i + \omega_n
pz \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \cdot
(1- \zeta_{i,m,n}) \Big) (qz)^i.
\end{equation}

\subsection{The expression of the transfer matrix}
\label{sec:expression_of_M}

Collecting and arranging the results above, the final expression of the
transfer matrix $M(z)$ appears as
\begin{equation*}
\begin{blockarray}{cccccccc}
   & \dn{0} & \ldots & \dn{N} & \up{1} & \ldots & \up{\gamma-1} & \nd \\
\begin{block}{c[ccccccc]}
\dn{0} & A_{0,0}(z) & \ldots & A_{0,N}(z) & B_1(z) & \ldots &
    B_{\gamma-1}(z) & C_0(z) \\
\vdots & \vdots & \ddots & \vdots & \vdots & \ddots &
    \vdots & \vdots \\
\dn{N} & A_{N,0}(z) & \ldots & A_{N,N}(z) & B_1(z) & \ldots &
    B_{\gamma-1}(z) & C_N(z) \\
\up{1} & D_{1,0}(z) & \ldots & D_{1,N}(z) & 0 & \ldots & 0 & E_1(z) \\
\vdots & \vdots & \ddots & \vdots & \vdots & \ddots &
    \vdots & \vdots \\
\up{\gamma-1} & D_{\gamma-1,0}(z) & \ldots & D_{\gamma-1,N}(z) & 0 &
  \ldots & 0 & E_{\gamma-1}(z) \\
\nd & 0 & \ldots & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray}
\end{equation*}
where
\begin{gather}
\tag{\ref{eq:A}}
A_{m,n} =
\omega_n pz \sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i + \omega_n
pz \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \cdot
(1- \zeta_{i,m,n}) \Big) (qz)^i \\
\tag{\ref{eq:B}}
B_i(z) = \Big( \xi_i^N-\xi_{i-1}^N \Big) (qz)^i \\
\tag{\ref{eq:C}}
C_m(z) =
\sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i +
  \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \Big) (qz)^i \\
\tag{\ref{eq:D}}
D_{j,m}(z) = \omega_m pz \sum_{i=0}^{\gamma-j-1} (qz)^i \\
\tag{\ref{eq:E}}
E_j(z) = \sum_{i=0}^{\gamma-j-1} (qz)^i
\end{gather}
and where
\begin{gather}
\tag{\ref{eq:omega}}
\omega_m = {N \choose m} \big(1 - \mu/3\big)^{N-m} \big(\mu/3\big)^m \\
\tag{\ref{eq:xi}}
\xi_i = 1-(1-\mu)^i \\
\tag{\ref{eq:eta}}
\eta_i = 1-(1-\mu)^i\mu/3 \\
\tag{\ref{eq:zeta}}
\zeta_{i,m,n} = \sum_{r\geq1} {N-m \choose r}
(1-\mu)^{ri} \psi_{i,m,n,r} \bigg/ {N \choose n} \\
\notag
\psi_{i,m,n,r} = \sum_{q \geq 0}{m \choose q}{N-m-r \choose n-r-q}
\xi_i^{n-r-q}.
\end{gather}

\section{Computing the probabilities}
\label{sec:compute}

\subsection{Coefficient extraction}
\label{sec:extract}

The purpose of constructing the weighted generating function $F(z) = a_0 +
a_1z + a_2z^2 + \ldots$ is to extract the coefficient $a_k$, which
represents the probability that a read of size $k$ does not contain a
good seed.

$F(z)$ is the entry of the matrix $M(z) \cdot (I-M(z))^{-1}$ at the first
row and the last column, associated with terminators $\downarrow_{/0}$ and
$|$. The expression of $M(z)$ is too complex to compute this function
directly. Instead, we return to the definition $M(z) \cdot (I-M(z))^{-1} =
M(z) + M(z)^2 + \ldots$ and observe that the terms $M(z)^{k+2},
M(z)^{k+3}, \ldots$ have no influence on the coefficients $a_0, a_1,
\ldots, a_k$.

Indeed, a read of size $k$ has at most $k+1$ segments\footnote{Each
segment contains a terminator, which has size 1, except the unique
terminator $|$, which has size 0.}. Since the entries of $M(z)^n$ are the
weighted generating functions of reads with exactly $n$ segments, $a_k$
cannot depend on $M(z)^{k+2}, M(z)^{k+3}$. More formally, we can prove by
induction that all the entries of $M(z)^k$ are divisible by $z^{k-1}$,
showing that the contribution of $M(z)^{k+2} + M(z)^{k+3} + \ldots$ to
$a_0 + a_1z + \ldots +a_kz^k$ is $0$.

So to extract the coefficients of $F(z)$ up to order $k$ we can compute
the matrix $M(z) + M(z)^2 + \ldots + M(z)^{k+1}$, isolate the entry at the
first row and the last column and then compute the terms of the Taylor
expansion up to order $k$.

But we can do better than that. Since we are only interested in the
coefficients up to order $k$, we can replace the infinite power series
of the terms $A_{m,n}(z)$ and $C_m(z)$ by their truncated versions where
we keep only the terms up to order $k$. Likewise, we perform all
algebraic operation on truncated polynomials of order $k$, \textit{i.e.},
we discard at every stage the coefficients of order $k+1$ or greater.

More formally, truncated polynomials are defined as polynomials whose
coeffients of degree greater than $k$ are 0. Multiplying two truncated
polynomials $P(z)$ and $Q(z)$ gives a truncated polynomial $R(z)$ whose
coefficients up to order $k$ are those of $P(z)Q(z)$. Using truncated
polynomials of order $k$, the entry of the matrix $M(z) + M(z)^2 + \ldots
+ M(z)^{k+1}$ at the first row and the last column directly provides the
coefficients of interest.

But we can do even better than that. A read with $n$ segments contains at
least $m=\lfloor{(n-1)/2}\rfloor$ errors. Indeed, two segments terminated
by $\uparrow$ cannot follow each other, so at least (approximately) half
of the segments must be terminated by $\downarrow$, which means that the
segment contains a read error. Since read errors typically have a small
probability of occurrence, all the coefficients of $M(z)^n$ rapidly
converge to $0$ as $n$ increases. So instead of computing the matrix $M(z)
+ M(z)^2 + \ldots + M(z)^{k+1}$, we can interrupt the summation after a
certain power of $M(z)$ because the terms become negligible.

The number of errors $X$ in a read of size $k$ has a Binomial
distribution $X \sim B(k,p)$. From\cite{arratia1989tutorial} we can bound
the probabilities of the tail with the expression

\begin{equation}
\label{eq:bound}
Pr(X \geq m) \leq \exp \left( (m-k)\log \frac{k-m}{k(1-p)} -m\log
\frac{m}{kp} \right).
\end{equation}

Using the formula above, we can thus bound the probability that a read has
$n$ segments or more. We compute the matrices $M(z) + M(z)^2 + \ldots
+M(z)^n$ where the weighted generating functions have been replaced by
truncated polynomials and we extract $F_n(z)$, the entry at the first
column and the last row. When the bound is lower than a set fraction
$\varepsilon$ of the coefficient of $z^k$, we stop the computations.
Typically $\varepsilon = 0.01$ so this method ensures that the
probabilities that a read of size $k$ has no good seeds are accurate to
within 1\%.

This approach is feasible when the matrix $M(z)$ is small, but recall that
$M(z)$ has $N+\gamma+1$ rows/columns. Some sequences can have millions of
duplicates in plant and animal genomes (\textit{i.e.} $N > 10^6$), which
makes the computations with this method unfeasible because even a few
matrix multiplications are prohibitive. For large $N$, another method is
clearly needed.

\subsection{Monte Carlo sampling}
\label{sec:montecarlo}

The symbolic representation as MEM segments can be used for an efficient
method to sample reads. Instead of generating the nucleotides of the $N+1$
sequences one by one, one can generate a single sequence of MEM segments.
Since the number of segment does not depend on $N$, we can obtain a fast
Monte Carlo method to sample millions of reads and count the proportion
that contain a good seed.

The principle is to proceed in cycles of two steps. We first sample the
position of the next read error, which gives the position of the next
symbol $\downarrow$. The second step is to determine whether there is a
symbol $\uparrow$ before that. For this we sample the number of masks that
vanish before the symbol $\downarrow$. If they all vanish, the read
contains a good seed, provided the next read error is at a distance
greater than $\gamma$. Otherwise, the process is repeated until we
generate a good seed, or until the read has size $k$ or greater (in which
case it has no good seed).

The method is summarized by the algorithm below. It requires efficient
methods to generate random samples with a geometric distribution and a
binomial distribution. Sampling from a geometric distribution can be done
by computing the logarithm of a uniform $(0,1)$ random variable. Sampling
from a binomial distribution can be done by the method of
Kachitvichyanukul and Schmeiser\cite{kachitvichyanukul1988binomial}.

\begin{algorithm}[H]
\label{alg:mcmc}
\SetAlgoLined
\KwResult {Sample a read at random. Return 1 if the read contains a
good seed, otherwise return 0.}
  $\lambda \leftarrow k$ \Comment*[r]{Size remaining to generate.}
  $m \leftarrow 0$ \Comment*[r]{Current number of hard masks.}
  \While {$\lambda > 0$}{
    $i \leftarrow geom(p)-1$ \Comment*[r]{Error-free nucleotides.}
    \eIf {$i \geq \lambda$}{
      \eIf{$\lambda < \gamma$} {
        return 0\; }{
        $h \leftarrow binom(m,(1-\mu)^\lambda)$
            \Comment*[r]{Surviving hard masks.}
        return 1 if $h = 0$, otherwise return 0\; }
    }
    {
      $h \leftarrow binom(m,(1-\mu)^i)$
          \Comment*[r]{Surviving hard masks.}
      $s \leftarrow binom(N-m,(1-\mu)^i \mu/3)$
          \Comment*[r]{Surviving soft masks.}
      \eIf {$ i \geq \gamma$ and $h = 0$ and $s = 0$}{
        return 1\;}{
        $m \leftarrow s + binom(N-s, \mu/3)$\;
        $\lambda \leftarrow \lambda - (i+1)$\;}
    }
 }
\end{algorithm}

In practice, we use a slightly modified version of this algorithm, that
returns the position of the good seeds if they are generated. This allows
us to estimate the probabilities for \emph{all} read sizes up to $k$. We
typically run this Monte Carlo sampling scheme 10,000,000 times, so that
we can estimate probabilities as low as $\sim1$ in 1 million.

The running speed depends on the error rate of the instrument $p$, but not
on the number of off-target sequences $N$. For values of $p$ around 0.01,
the running time for 10,000,000 iterations is $1$-$10$ seconds on current
hardware.

\section{False positives}

So far we have designed a method to estimate the probability that a
sequencing read contains a good seed. We referred to this as outcome $i.$
in section~\ref{sec:seedheur}. However, we are interested in the
probability that a read is mapped at an incorrect location, which is
outcome $ii.$

Computing the probability of outcome $ii.$ is hard, but we can compute the
probability that the read is not mapped at all, which is outcome $iii.$
The probability of outcome $ii.$ is then found by subtraction.

\subsection{The probability of outcome $iii.$}

Outcome $iii.$ implies that the read contains no seed at all, neither
on-target, nor off-target. This requires that there is no exact match of
size $\geq \gamma$ for \emph{any} sequence. Indeed, if it were the case,
then there would a MEM seed of size $\geq \gamma$ for at least one of the
sequences.

Let $E_0$ denote the event that there is no on-target match of size $\geq
\gamma$, and $E_j$ denote the event that there is no match of size $\geq
\gamma$ for off-target sequence $j$ ($1 \leq j \leq N$). We are thus
interested in computing $P(E_0 \cap \ldots \cap E_N)$. Since we assume
that the off-target sequences have evolved through the same mutagenesis
process and independently of each other, we can write

\begin{eqnarray}
\notag
P(E_0 \cap \ldots \cap E_N) &=&
  P(E_0) \cdot P(E_1 \cap \ldots \cap E_N | E_0) \\
\label{eq:PE0E1}
  &=& P(E_0) \cdot P(E_1 | E_0)^N =
  P(E_0) \cdot \left( \frac{P(E_0 \cap E_1)}{P(E_0)} \right)^N.
\end{eqnarray}

Expression (\ref{eq:PE0E1}) shows that we can compute the probability of
outcome $iii.$ from two quantities, namely $P(E_0)$ and $P(E_0 \cap E_1)$.

\subsection{Computing $P(E_0)$}
\label{sec:cPE0}

$P(E_0)$ is straightforward to evaluate because it involves a single
sequence. Methods to do this efficiently have been explained in detail
elsewhere\cite{filion2018analytic}. Here we only give the recurrence
equation to compute $P(E_0)$. Assume that the probability of a sequencing
error $p$ is known, that $q = 1-p$, and that $a_k$ is the probability that
a read of size $k$ does not contain a match of size $\geq \gamma$ for the
target. Then the following equation can be used to compute $a_k$ by
recurrence:

\begin{equation}
\label{eq:PE0}
a_k = 
\begin{cases}
1            &\quad\text{if } k < \gamma, \\
1 -pq^\gamma &\quad\text{if } k = \gamma, \\
a_{k-1} -pq^\gamma \cdot a_{k-\gamma-1} &\quad\text{otherwise.}
\end{cases}
\end{equation}

Knowing the size of the read and the error rate of the sequencing
instrument, we can thus compute $P(E_0)$ using expression (\ref{eq:PE0}).

\subsection{Computing $P(E_0 \cap E_1)$}
\label{sec:cPE0E1}

The event $E_0 \cap E_1$ is that the read contains no match of size $\geq
\gamma$ for either the target or the first off-target sequence (here
``first'' refers to any fixed off-target sequence). We compute the
probability of this event with the strategy described in
section~\ref{sec:symbolic}: we first recode the reads using a different
alphabet and then we specify the transfer matrix of the reads that have no
match for any sequence. The powers of this matrix hold the weighted
generating function of interest, from which we extract the coefficients
that correspond to the terms $P(E_0 \cap E_1)$ for reads of different
sizes.

The reads are recoded as sequences of symbols from the so-called ``+/-
alphabet'' $\mathcal{A}_{+/-} = \{\square, |, \downarrow_{/1}^-,
\downarrow_{/2}^-, \ldots, \downarrow_{/1}^+, \downarrow_{/2}^+, \ldots,
\Downarrow\}$. The symbols $\downarrow_{/j}^-$ signify that the nucleotide
is a mismatch against the off-target sequence, the symbols
$\downarrow_{/j}^+$ signify that it is a mismatch against the target, and
the symbol $\Downarrow$ signifies that it is a mismatch against both. As
before, every other nucleotide is replaced by the symbol $\square$, and
the terminator $|$ is appended at the end of the read. We again define
reads as sequences of segments, except that now the terminators are the
symbols $\downarrow_{/j}^-$, $\downarrow_{/j}^+$, $\Downarrow$ and $|$.

The number $j$ in the symbol $\downarrow_{/j}^-$ indicates the relative
position of the last mismatch against the target. Likewise, the number $j$
in the symbol $\downarrow_{/j}^+$ indicates the relative position of the
last mismatch against the off-target sequence. In other words, the index
$j$ is the size of the match for the sequence that does not have a
mismatch. The terminators thus capture the local state of the read. For
instance the symbol $\downarrow_{/7}^-$ indicates that the nucleotide is a
mismatch against the off-target sequence, that it is a match for the
target, and that the six previous nucleotides were also a match for the
target.

\begin{figure}[h]
\centering
\includegraphics[scale=0.85]{sketch_dual.pdf}
\caption{\textbf{The +/- encoding.}
A example read is represented in the +/- alphabet. The symbols
$\downarrow_{/j}^-$ represent a mismatch against the off-target sequence,
the symbols $\downarrow_{/j}^+$ represent a mismatch against the target,
and the symbol $\Downarrow$ represent a mismatch against both. The index
$j$ is the number of matches since the last mismatch for the other
sequence. Note the presence of the terminator $|$ at the end of the read.
}
\label{fig:dual}
\end{figure}

We use the same error model as before: for every nucleotide, the read
differs from the target with a fixed probability $p$ and the sequences
differ from each other with a fixed probability $\mu$. With these
definitions, the probability of occurrence of the symbol $\square$ is $a =
(1-p)(1-\mu)$. Likewise, the collective probability of occurrence of the
symbols $\downarrow_{/j}^-$ is $b = (1-p)\mu$, that of the symbols
$\downarrow_{/j}^+$ is $c = p\mu/3$, and the probability of occurrence of
the symbol $\Downarrow$ is $d = p(1-\mu/3)$. It is easy to verify that
$a+b+c+d=1$.

These definitions allow us to specify the weighted generating function of
the segments. For instance, the weighted generating function of the
first segment of Fig.~\ref{fig:dual} is $(az)bz$, and that of the second is
$(az)^5cz$.

To specify the transfer matrix, we must ensure that we eliminate
the concatenations of two segments that would create a match of size $\geq
\gamma$ for any of the two sequences. After defining
\begin{equation*}
\begin{gathered}
r_i^+(z) = (az)^icz, \\
r_i^-(z) = (az)^ibz, \\
R_i(z) = \big( 1 + az + \ldots + (az)^i \big)dz, \\
F_i(z) = 1 + az + \ldots + (az)^i,
\end{gathered}
\end{equation*}
we can verify that the expression of the transfer matrix $L(z)$ is

\begin{equation*}
\begin{blockarray}{ccccccccc}
   & \scriptstyle{\Downarrow} & \scriptstyle{\downarrow_{/1}^+} & 
    \ldots & \scriptstyle{\downarrow_{\gamma-1}^+} &
    \scriptstyle{\downarrow_{/1}^-} & \ldots &
    \scriptstyle{\downarrow_{/\gamma-1}^-} & \scriptstyle{|} \\
\begin{block}{c[cccccccc]}
\scriptstyle{\Downarrow} & R_{\gamma-1}(z)  & r_0^+(z) & \ldots &
    r_{\gamma-2}^+(z) & r_0^-(z) & \ldots & r_{\gamma-2}^-(z) &
    F_{\gamma-1}(z) \\
\scriptstyle{\downarrow_{/1}^+} & R_{\gamma-2}(z) & & & & & & &
    F_{\gamma-2}(z) \\
\scriptstyle{\downarrow_{/2}^+} & R_{\gamma-3}(z) & & & & & & &
    F_{\gamma-3}(z) \\
\vdots & \vdots & & A(z) & & & B(z) & & \vdots \\
\scriptstyle{\downarrow_{/\gamma-1}^+} & R_0(z) & & & & & & & F_0(z) \\
\scriptstyle{\downarrow_{/1}^-} & R_{\gamma-2}(z) & & & & & & &
    F_{\gamma-1}(z) \\
\scriptstyle{\downarrow_{/2}^+} & R_{\gamma-3}(z) & & & & & & &
    F_{\gamma-2}(z) \\
\vdots & \vdots & & C(z) & & & D(z) & & \vdots \\
\scriptstyle{\downarrow_{/\gamma-1}^+} & R_0(z) & & & & & & & F_0(z) \\
\scriptstyle{|} & 0 & 0 & \ldots & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}
where $A(z)$, $B(z)$, $C(z)$ and $D(z)$ are square matrices with
$\gamma-1$ rows/columns, that are defined as

\begin{equation*}
A(z) = 
\begin{blockarray}{cccccc}
   & \scriptstyle{\downarrow_{/1}^+} & \scriptstyle{\downarrow_{/2}^+} &
    \ldots & \scriptstyle{\downarrow_{/\gamma-2}^+} &
    \scriptstyle{\downarrow_{/\gamma-1}^+} \\
\begin{block}{c[ccccc]}
\scriptstyle{\downarrow_{/1}^+} & 0 & r_0^+(z) & \ldots &
    r_{\gamma-4}^+(z) & r_{\gamma-3}^+(z) \\
\scriptstyle{\downarrow_{/2}^+} & 0 & 0 & \ldots &
    r_{\gamma-5}^+(z) & r_{\gamma-4}^+(z) \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\scriptstyle{\downarrow_{/\gamma-2}^+} & 0 & 0 & \ldots & 0 & r_0^+(z) \\
\scriptstyle{\downarrow_{/\gamma-1}^+} & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}

\begin{equation*}
B(z) = 
\begin{blockarray}{cccccc}
   & \scriptstyle{\downarrow_{/1}^-} & \scriptstyle{\downarrow_{/2}^-} &
   \ldots & \scriptstyle{\downarrow_{/\gamma-2}^-} &
   \scriptstyle{\downarrow_{/\gamma-1}^-} \\
\begin{block}{c[ccccc]}
\scriptstyle{\downarrow_{/1}^+} & r_0^-(z) & r_1^-(z) & \ldots &
    r_{\gamma-3}^-(z) & r_{\gamma-2}^-(z) \\
\scriptstyle{\downarrow_{/2}^+} & r_0^-(z) & r_1^-(z) & \ldots &
    r_{\gamma-3}^-(z) & 0 \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\scriptstyle{\downarrow_{/\gamma-2}^+} & r_0^-(z) & r_1^-(z) &
    \ldots & 0 & 0 \\
\scriptstyle{\downarrow_{/\gamma-1}^+} & r_0^-(z) & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}

\begin{equation*}
C(z) = 
\begin{blockarray}{cccccc}
   & \scriptstyle{\downarrow_{/1}^+} & \scriptstyle{\downarrow_{/2}^+} &
    \ldots & \scriptstyle{\downarrow_{/\gamma-2}^+} &
    \scriptstyle{\downarrow_{/\gamma-1}^+} \\
\begin{block}{c[ccccc]}
\scriptstyle{\downarrow_{/1}^-} & r_0^+(z) & r_1^+(z) & \ldots &
    r_{\gamma-3}^+(z) & r_{\gamma-2}^+(z) \\
\scriptstyle{\downarrow_{/2}^-} & r_0^+(z) & r_1^+(z) & \ldots &
    r_{\gamma-3}^+(z) & 0 \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\scriptstyle{\downarrow_{/\gamma-2}^-} & r_0^+(z) & r_1^+(z) & \ldots &
    0 & 0 \\
\scriptstyle{\downarrow_{/\gamma-1}^-} & r_0^+(z) & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}

\begin{equation*}
D(z) = 
\begin{blockarray}{cccccc}
   & \scriptstyle{\downarrow_{/1}^-} & \scriptstyle{\downarrow_{/2}^-} &
    \ldots & \scriptstyle{\downarrow_{/\gamma-2}^-} &
    \scriptstyle{\downarrow_{/\gamma-1}^-} \\
\begin{block}{c[ccccc]}
\scriptstyle{\downarrow_{/1}^-} & 0 & r_0^-(z) & \ldots &
    r_{\gamma-4}^-(z) & r_{\gamma-3}^-(z) \\
\scriptstyle{\downarrow_{/2}^-} & 0 & 0 & \ldots &
    r_{\gamma-5}^-(z) & r_{\gamma-4}^-(z) \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\scriptstyle{\downarrow_{/\gamma-2}^-} & 0 & 0 & \ldots & 0 & r_0^-(z) \\
\scriptstyle{\downarrow_{/\gamma-1}^-} & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray}.
\end{equation*}

The matrix $L(z)$ is relatively sparse and most non-zero entries are
monomials. In addition, the number of rows/columns is approximately
$2\gamma$, so the dimension of the matrix rarely exceeds 50. This means
that computing the powers of $L(z)$ is not prohibitive and we do not need
to design a Monte Carlo sampling scheme as we did before.

That being said, we do not need to compute all the powers up to $k$
because the terms vanish and become negligible after a certain iteration.
Indeed, the terms of $L(z)^n$ are the weighted generating function of the
reads without a match for either sequences and that contain exactly $n$
segments. Of those $n$ segments, $n-1$ are terminated by a mismatch
against at least one of the sequences. Following the rationale of
section~\ref{sec:extract}, we define $\tilde{p}$ as the upper bound on the
probability of a mismatch $\tilde{p} = \max\{b,c,d\}$, and updating
formula (\ref{eq:bound}) we bound the probability that a read of size $k$
contains $m = n-1$ or more mismatches

\begin{equation*}
Pr(X \geq m) \leq \exp \left( (m-k)\log \frac{k-m}{k(1-\tilde{p})} -m\log
\frac{m}{k\tilde{p}} \right).
\end{equation*}

The strategy is the same as in section~\ref{sec:extract}: we compute the
sum of matrices $L(z) + L(z)^2 + \ldots + L(z)^n$ where the weighted
generating functions have been replaced by truncated polynomials and we
extract the entry at the first column and the last row. When the bound is
lower than a set fraction $\varepsilon$ of the coefficient of $z^k$, we
stop the computations.

\subsection{The probability of false positives}

For given values of the parameters $p$, $\mu$ and $N$, we can now compute
the probability of a false positive, which is outcome $ii$.

For this, we compute the probability of outcome $i.$, either by computing
the powers of $M(z)$ or by Monte Carlo sampling as explained in
section~\ref{sec:compute}. Next we compute the probability of outcome
$iii.$ using expression (\ref{eq:PE0E1}), where the terms are computed as
explained in sections~\ref{sec:cPE0E1} and \ref{sec:cPE0}.

\section{Estimating the parameters}
\label{sec:est}

So far we have assumed that we know all the specifications of the problem,
including the error rate of the instrument $p$, the number of off-target
sequences $N$, and their rate of sequence divergence $\mu$ (or more
accurately the probability that they differ from the target at every
nucleotide).

In general, $p$ is known from the vendor of the sequencing instrument, but
$N$ and $\mu$ can change for every read and they are usually unknown. We
thus need an estimation algorithm that can run fast enough that it does
not slow the mapping process. The main idea here is to use the backward
search, which is the method used to find the MEM seeds in the first place.

The backward search is a method that was first proposed by Ferragina and
Mazini\cite{ferragina2000opportunistic} to identify exact matches between
a text query and an index. The query is extended one letter/nucleotide at
a time, and at each iteration the search returns the number of matches in
the index/genome. This is thus a very efficient method to identify MEM
seeds, because one can keep extending the query until there is no match in
the genome or until the end of the read.

What matters for us is that we can run the backward search at some
arbitrary position of the read. The $i$-th iteration returns the number
$N_i$ of matches in the genome for a query sequence of length $i$.
Ignoring sequencing errors and focusing only on the $N$ off-target
sequences, the probability that the count goes from $N_i$ to $N_{i+1}$ is

\begin{equation*}
P(N_{i+1}|N_i) = {N_i \choose N_{i+1}}
\mu^{N_i-N_{i+1}}(1-\mu)^{N_{i+1}}.
\end{equation*}

Assuming that there are $N = N_0$ duplicates and that we run $m$
iterations of the backward search, the probability of the observations
$(N_1, N_2, \ldots, N_m)$ is thus

\begin{equation*}
P(N_1, \ldots, N_m|N_0) = 
{N_0 \choose N_1 } \ldots {N_{m-1} \choose N_m }
\mu^{N_0-N_m}(1-\mu)^{N_1+\ldots+N_m}.
\end{equation*}

This is a familiar distribution in disguise: by defining $\Delta_i =
N_{i-1} - N_i$, the series of observations $(\Delta_1, \ldots, \Delta_m,
N_m)$ given $N_0$ has a multinomial distribution with parameters
$\big(\mu, \mu(1-\mu), \ldots, \mu(1-\mu)^{m-1}, (1-\mu)^m\big)$. More
specifically, the probability of observing the sequence $(\Delta_1,
\ldots, \Delta_m, N_m)$ given $N_0$ is

\begin{align*}
{N_0 \choose \Delta_1, \ldots, \Delta_m, N_m}
\mu^{\Delta_1}\big(\mu(1-\mu)\big)^{\Delta_2}
\ldots \big(\mu(1-\mu)^{m-1}\big)^{\Delta_m}(1-\mu)^{mN_m}.
\end{align*}

At that point, we need to bring back into the light the practical
consideration that short matches are uninformative. This implies that the
first iterations of the backward search say nothing about the number of
duplicates of the sequence. In practice, we are better off ignoring the
output of the first $n-1$ iterations by just discarding them. We observed
that running 30 iterations of the backward search and discarding the
output of the first 20 gives good results (\textit{i.e.} we set $n = 21$
and $m=30$).

Now, if $N_0, N_1, \ldots, N_{n-1}$ are missing, we obtain the probability
of observing $(\Delta_{n+1}, \ldots, \Delta_m, N_m)$ from the marginal
distribution as

\begin{align*}
\begin{split}
{N_0 \choose N_0-N_n, \Delta_{n+1}, \ldots, \Delta_m, N_m}
\big(1-(1-\mu)^n\big)^{N_0-N_n}
\big(\mu(1-\mu)^n\big)^{\Delta_{n+1}} \times
\ldots \\
\ldots \times
\big(\mu(1-\mu)^{m-1}\big)^{\Delta_m}(1-\mu)^{mN_m}.
\end{split}
\end{align*}

Now replacing again $\Delta_i$ by $N_{i-1}-N_i$, the log-likelihood can be
expressed as

\begin{equation}
\begin{split}
\ell = cst + \log \Gamma(N_0+1) - \log \Gamma(N_0-N_n-1)
+ (N_0-N_n) \log \big(1-(1-\mu)^n\big) + \\
(N_n-N_m) \log(\mu) + (nN_n + N_{n+1} + \ldots + N_m) \log(1-\mu).
\end{split}
\end{equation}

To estimate the parameters by maximum likelihood, we differentiate $\ell$
with respect to $N_0$ and with respect to $\mu$, then we equate both terms
to 0. We thus have to solve

\begin{gather}
\label{eq:psi}
\Psi(N_0+1)-\Psi(N_0-N_n+1) +
  \log\big( 1-(1-\mu)^n \big) = 0, \text{ and} \\
\label{eq:N0}
\frac{N_n-N_m}{\mu}
-\frac{nN_n + N_{n+1}+\ldots+N_m}{1-\mu} +
(N_0-N_n)\frac{n(1-\mu)^{n-1}}{1-(1-\mu)^n} = 0.
\end{gather}

In the expressions above, $\Psi(\cdot)$ is the digamma function, defined
as the derivative of $\log \Gamma(\cdot)$. We observe that equation
(\ref{eq:N0}) gives an expression of $N_0$ as a function of $\mu$ and of
the observations, namely

\begin{equation*}
N_0 = N_n +
\left(\frac{nN_n+N_{n+1}+\ldots+N_m}{1-\mu}-
\frac{N_n-N_m}{\mu} \right)
\frac{1-(1-\mu)^n}{n(1-\mu)^{n-1}}.
\end{equation*}

To find the maximum likelihood estimate of $\mu$, we can substitute this
expression in equation (\ref{eq:psi}), obtaining an equation depending on
$\mu$ and on the observations, which can be solved numerically. One can
then substitute the solution into equation (\ref{eq:N0}) to obtain the
estimate of the number of duplicate sequences $N_0$.

However, the estimations of $\mu$ and $N_0$ show large deviations from the
target values on simulated data. In practice, one does not need to know
the parameters with exquisite precision, so to avoid unrealistic
estimates, we pick the best $\mu$ from a short list of candidate values
(\textit{e.g.}, 0.02, 0.04 and 0.06). To do this, we choose a value
$\mu^*$ from the candidate set, we compute $N_0(\mu^*)$ from equation
(\ref{eq:N0}) and we evaluate $\ell$ with these values. Finally, we select
the $\mu$ that gives the higest value of $\ell$, and the estimate for
$N$ is set to $N_0(\mu)$.

Finding the estimates of $N$ and $\mu$ can be performed in negligible time
compared to running the backward search. We can repeat this method at
different areas of the read and finally average the estimates to get
reasonable values of $N$ and $\mu$ for the whole read. With this
information, we can compute the probability that the read has no good seed
using the methods presented above.

\section{Neglected practical considerations}

To keep the focus on the MEM seeding process, we have ignored some
practicalities that we now have to revisit in order to evaluate the
feasibility of our approach.

\subsection{Is the true location the best one?}
\label{sec:not_the_best}

So far we have assumed that the true location is also the best one, so
that it is always discovered when it is in the candidate set. But how
likely is it that the true location is not the best one, and thus that it
is virtually impossible to discover?

Obviously, it is possible that the target has an exact duplicate. When
that happens, the read cannot be mapped unambiguously, but this case is
relatively easy to detect, so the mapper can return a low confidence score
and thus avoid making an error.

From now we assume that the read has no strict duplicate. The probability
that the read of size $k$ has $m$ errors is
\begin{equation*}
{k \choose m}p^m(1-p)^{k-m}.
\end{equation*} 
A given duplicate has a better score if its Hamming distance from the read
is less than $k$. In a first order approximation, this happens if the
duplicate matches one of the $k$ errors, and also the $L-k$ correct
nucleotides. The probability of this even is
\begin{equation*}
m(1-\mu)^{k-m}\mu/3.
\end{equation*}

There are $N$ duplicates that evolve independently, so the probability
that one of them has a better score than the target is
\begin{equation*}
1-\left(1-m(1-\mu)^{k-m}\mu/3\right)^N \approx Nm(1-\mu)^{k-m}\mu/3.
\end{equation*}

Combining the first and the last expression, we find that the probability
that the target does not have the best score is approximately
\begin{equation}
\label{eq:not_the_best}
\sum_{m=0}^k {k \choose m}p^m(1-p)^{k-m} Nm\mu/3(1-\mu)^{k-m} = 
Nk p \left(1-\mu(1-p)\right)^{k-1}\mu/3.
\end{equation}

For instance, for reads of size $k=100$ with error rate $p=0.01$, where
the target has $N=5$ duplicates with divergence $\mu=0.05$ we obtain an
estimate close to $10^{-4}$. In comparison, the probability of a false
positive when using MEM seeds with a minimum size $\gamma=17$ is
approximately $4.4\cdot10^{-4}$.

The probability that the target does not have the best score is usually
small, but it is of the same order of magnitude as the probabilities of
false positives with MEM seeds. The reason is that a common source of
false positives are the cases when the target does not have the best
score. Indeed, in this case the target is much less likely to have a MEM
seed.

In short, the event that the true location is not the best one is already
incorporated in the false positive rate and we do not recommend further
updating the estimates. In the case that one is interested in this event
\textit{per se}, formula (\ref{eq:not_the_best}) gives an approximate
probability of its occurrence. When $\mu$ is unknown, we can replace it by
its worst case value $\mu = 1/qL$, where $q=1-p$. But $N$ must be known at
least approximately because the proability increases with $N$.

In summary, the case that the true location is not the best one has a
small but not negligible probability. It is partly redundant with the
false positives generated during the MEM seeding process, and therefore it
is not straightforward to adjust the estimates accordingly. We do not
adjust them at all and consider that this case is already implicitly
accounted for.

\subsection{MEM seeds at random locations}
\label{sec:short_MEMs}

To simplify the problem, we have assumed that MEM seeds can match only the
target or one of its duplicates. In practice, MEM seeds that overlap a
sequencing error often match random locations of the genome instead of the
duplicates. When this occurs, the MEM seeds are typically small because 
the match is random.

We select MEM seeds of size at least $\gamma$, and it is important for the
validity of the theory that all the MEM seeds above the threshold are
considered candidate locations. If not, the estimates of the false
positive rate are off. The hope is that spurious MEM seeds are shorter
than $\gamma$ so that they are invisible, but this is not always the case.
And if they qualify, we need to verify them.

Such random hits can be problematic for two reasons: First, the time spent
verifying them is wasted. Second, they can cause false positives.
Fortunately, both issues can be addressed so that they do not constitute a
major problem.

To save time during the alignment phase, we can prioritize the seeds so
that the best hit is likely to be discovered first, allowing us to bail
out from other alignments as early as possible. We can even bound
the maximum quality of some candidates so that the alignment can be
skipped altogether. For instance, if we already have a candidate at
distance 1 from the read, there is no need to verify candidates from MEM
seeds in the middle of the read. Indeed, the hit must be at distance at
least 2 because there is a mismatch before the MEM seed and another one
after it.

These considerations depend on the implementation of the mapper so we will
not further develop them. What matters is that random hits do not have to
impose a significant burden on the time needed to verify the candidates.

The second issue is that random hits can generate false positives. Such
false positives happen when the read does not belong to the reference
genome, or when it contains so many errors that no proper MEM seed was
found. Since the hit is not homologous to the read, the alignment score is
essentially random. The upshot is that it is easy to detect.

Indeed, if the candidate location is random, mismatches occur with
probability 3/4. If it is the right location, mismatches occur with
probability $p$. We discard the MEM seed, which is an automatic match of
size at least $\gamma$. Say that there remains $L$ nucleotides and that
$m$ of them are mismatches for the candidate location. Using Bayes
formula, we find that the probability that the location is random given
the observed alignment score is
\begin{equation}
\label{eq:bayes}
\frac{1}{1 + 4^L(p/3)^mq^{L-m}(1-\beta)/\beta},
\end{equation}
where $q=1-p$ and $\beta$ is the prior probability that the hit is
random.

One may have information about the level of contamination, allowing to
choose a meaningful value for $\beta$. For instance, sequencing runs are
often spiked by foreign DNA such as the phiX genome or other. If these
reads are not filtered automatically, $\beta$ can be set to their expected
frequency. For sequences that belong to the reference genome, random hits
are rather rare because the read must contain so many error that neither
the target nor its duplicates were discovered. The value of $\beta$ is
much smaller in this case, but nevertheless difficult to estimate.

That being said, the value of $\beta$ has little importance in practice.
Say that a read generates a hit in the genome such that 33 nucleotides
need to be aligned (the alignment status of the others was already known
from the seeding step). If the hit is random there is a 99.99\% chance
that at least 15 are mismatches. If $p = 0.01$, the denominator of
expression (\ref{eq:bayes}) is approximately $1 +
4.3\cdot10^{-18}(1-\beta)/\beta$. So unless $\beta < 10^{-17}$, the
support for the hypothesis that the hit is random is overwhelming.

Conversely, if the hit is not random, there is a 99.99\% chance that 4 or
less are mismathes. In this case, the denominator is approximately
$1+6.8\cdot10^9(1-\beta)/\beta$, so unless $\beta > 1-10^9$, the support
for the hypothesis that the hit is not random is overwhelming.

In general, one does not need to worry about the value of $\beta$ because
random hits are extremely visible when $p$ is small. For instance, one can
choose $\beta=1/2$ so that the term $(1-/\beta)/\beta$ disappears in
expression (\ref{eq:bayes}).

Because the probability that the best candidate is a random sequence is
either very small or very large, it either has no influence on the
probability of false positive, or completely dominates it. For every read,
we thus take the maximum of this probability or the estimated probability
that MEM seeding returned a false positive.

In summary, MEM seeds in random sequences happen relatively frequently,
but it is possible to minimize their computational burden. Also, they
are no cause for concern regarding false positives because they can easily
be detected using Bayes formula.

\section{Discussion}

We have devised a method to compute the probability that the MEM seeding
heuristic misguides the mapping algorithm so that it cannot find the true
location of the read. The pillars of our strategy are borrowed from
analytic combinatorics~\cite{flajolet2009analytic,
sedgewick2013introduction}, even though we do not follow the standard
programme. Constructing generating functions usually serves the purpose of
finding their singularities in order to approximate the solution. In our
case, however, the weighted generating functions cannot be computed so
this strategy is not applicable.

To find the probabilities of interest in the absence of a fully specified
weighted generating function, we compute only the first terms using
iterative methods as in sections~\ref{sec:extract}, \ref{sec:cPE0} and
\ref{sec:cPE0E1}, or using Monte Carlo sampling as in
section~\ref{sec:montecarlo}. In this regard, the breakthrough is the
encoding of reads as MEM segments, which is a form of implicit Markov
imbedding~\cite{fu1994distribution}.

The method is general, but it is important to clearly state the
assumptions it depends on. First and most importantly, we have ignored
insertions and deletions. We assume that the sequencing errors are
substitutions only, which makes the method adapted to the Illumina
technology, but not to deletion-prone instruments such as the Oxford
Nanopore technology. We also assume that insertions and deletions never
occur among duplicated sequences. This is obviously incorrect, but our
initial tests with real data suggest that this is a minor impediment. On
real genomes, the estimations are only approximately correct and this is
but one of several simplifying assumptions. The theory is substantially
intrincate; insertions and deletions would make it intractable.
It is presently unclear how gracefully incorporate them in the theory.

The second assumption is that all the candidate locations with a MEM
seed of size $\geq \gamma$ are tested with the exact sequence alignment
algorithm, so that the best location is always discovered if it is in the
candidate set. This is not completely unrealistic, but it is important to
realize that for plant and animal genomes, a shared MEM seed may have tens
of thousands of hits. Therefore, most mappers impose a limit on the number
of alignments so the best hit may not be found even if it is in the
candidate set. This is again a minor impediment, since we can correct the
probability that the location is a false positive when only some
candidates have been tested.

The third assumption is that all the duplicate sequences evolve
independently of each other and at the same rate. This is again incorrect,
because duplication events can happen continuously, creating complex
ancestry relationships. It is possible to infer the ancestry using tree
reconstruction techniques, but it would be challenging to incorporate this
information in the present theory. The symbols of the MEM alphabet
implicitly assume that the sequences are exchangeable and the complexity
of the calculations explodes if it is not the case.

The last assumption is that MEM seeds can match only the target or one of
its duplicates. This is not the case but we have shown how to gracefully
deal with this in section~\ref{sec:short_MEMs}.

The Monte Carlo approach of algorithm~\ref{alg:mcmc} is relatively
straightforward, so one may wonder why the approach with weighted
generating functions is necessary at all. The only reason is precision. To
estimate the frequency of an event by Monte Carlo sampling, it must occur
at least a few times in the simulation. For instance, with 10 million
rounds of sampling, frequencies around $1/1,000,000$ or lower cannot be
measured accurately. When not interested in rare events, it is thus a
perfectly reasonable strategy to use only Monte Carlo sampling.

For $N < 20$, the probability that a read does not contain a good seed is
the smallest, so we need a method that is accurate in this range.
Fortunately, the transfer matrix method is fast in this regime because the
matrix $M(z)$ has few rows/columns.

The proposed methods are compliant with the demand for speed of the
mapping algorithms. One needs to compute the probabilities only once for a
given value of $N$ and $\mu$ (the error rate $p$ is known and constant).
For $N > 20$, the iterative method is usually prohibitive and we need to
use Monte Carlo sampling instead. The running time depends on the desired
number of iterations, on $p$ and on the size of the reads $k$. Since those
are constant throughout the sequencing run, the method always takes the
same time (around 1 second for reads of size $100$-$200$ nucleotides on
modern hardware). The values of $N$ and $\mu$ can be binned in intervals
of interest so that there are only around $100$ pairs for a total cost of
a few minutes per run. Considering that mappers can process at most
$10,000$ reads per second per core, the time of mapping a sequencing run
of 250 million reads is over 7 hours, two orders of magnitude larger than
the time required to estimate the probabilities of error.

In conclusion, we presented a practical solution to the problem of
estimating the probability of false positives when using MEM seeds. This
solution is adapted for mapping short reads sequenced with the Illumina
technology. Being able to calibrate the MEM seeding heuristic not only
allows the user to choose how to balance speed versus accuracy, but also
opens new applications. For instance, one can map reads from contaminated
samples in pools of closely related genomes (\textit{e.g.}, modern human
and Neanderthal) in order to assign the reads to the organism they belong
to. The probabilities of false positives give the right level of
confidence in the assignment.

More generally, the analytic combinatorics programme is a very powerful
tool to address problems in bioinformatics. Here we have seen how this
strategy can be useful even when the generating functions cannot be
computed. Using the same ideas, one could calibrate heuristics used in
other alignment methods, especially in expanding fields such as long-read
technologies.

\section*{Acknowledgements}

We thank Ruggero Cortini for suggesting Monte Carlo methods as a way to
tackle intractable cases. We acknowledge the financial support of the
Spanish Ministry of Economy and Competitiveness (‘Centro de Excelencia
Severo Ochoa 2013-2017’, Plan Nacional BFU2012-37168), of the CERCA
Programme~/~Generalitat de Catalunya, and of the European Research Council
(Synergy Grant 609989).

%---------------------------------------------------------------
%---------------------------------------------------------------

\bibliography{references,pubmed}
\bibliographystyle{plain}

%----------------------------------------------------------------

\end{document}

%gs -dNoOutputFonts -sDEVICE=pdfwrite -o out.pdf latex.pdf