### 2019-03-31 MEM.tex
First complete sketch with exact, skip and MEM seeds.

\documentclass{article}

\usepackage{algorithm2e}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{authblk}
\usepackage[english]{babel}
\usepackage{blkarray}
\usepackage[font=small]{caption}
\usepackage{cite}
\usepackage{enumerate}
\usepackage{graphicx}

\setlength{\thickmuskip}{2mu plus 3mu minus 1mu}
\setlength{\medmuskip}{1mu plus 2mu minus 1mu}

\SetKwComment{Comment}{$\triangleright$\ }{}

% ---- Author affiliations ---- %

\renewcommand\Affilfont{\itshape\small}

% ---- Propositions, lemmas, defintions... ---- %

%\newtheorem{algorithm}{Algorithm}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
%\newtheorem{example}{Example}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
%\newtheorem{remark}{Remark}

% ---- Special environments (examples and remarks) ---- %

\newcounter{examplecounter}
\newenvironment{example}
{\small\vspace{0.5\baselineskip}
  \refstepcounter{examplecounter}%
  \noindent\textbf{Example \arabic{examplecounter}.}%
}{\vspace{-0.2\baselineskip}\begin{center}%
  $\star$\end{center}\vspace{0.5\baselineskip}}

\newcounter{remarkcounter}
\newenvironment{remark}
{\small\it\vspace{0.5\baselineskip}
  \refstepcounter{remarkcounter}%
  \noindent\textbf{Remark \arabic{remarkcounter}.}%
}{\vspace{0.5\baselineskip}}

\newenvironment{inset}
{\vspace{0.5\baselineskip}\begin{center}}
{\end{center}\vspace{0.5\baselineskip}}

% ---- Macros ---- %

\newcommand{\dn}[1]{\scriptstyle{\downarrow_{/#1}}}
\newcommand{\Dn}[1]{\scriptstyle{\Downarrow_{#1}}}
\newcommand{\up}[1]{\scriptstyle{\uparrow_{/#1}}}
\newcommand{\nd}{\scriptstyle{|}}
\newcommand{\modulo}[1]{\ (\mathrm{mod}\ #1)}

%---------------------------------------------------------------

\title{Estimating mapping quality when using seeding heuristics for
DNA alignment}

\author[1,2]{Guillaume J. Filion}
\author[1,2]{Eduard Zorita}
\affil[1]{Genome Architecture, Gene Regulation, Stem Cells and Cancer
Programme, Center for Genomic Regulation (CRG), The Barcelona Institute of
Science and Technology, Dr. Aiguader 88, Barcelona 08003, Spain.}
\affil[2]{University Pompeu Fabra, Doctor Aiguader, 08003 Barcelona,
Spain.}

\date{\today}

%---------------------------------------------------------------
%---------------------------------------------------------------

\begin{document}

\maketitle

\begin{abstract}
% Statement of the problem.
The increasing throughput of DNA sequencing technologies demands faster
algorithms to analyze sequencing data. The fate of most reads is to be
mapped to a reference sequence, typically a genome. Modern mappers rely on
heuristics to gain speed at a reasonable cost for accuracy. In the seeding
approach, short matches between the reads and the genome are used to
narrow the search to a set of candidate locations. Several seeding
variants used in modern mappers show good empirical performance but
theoretical results are still lacking, making them difficult to calibrate
or top optimize.
% Summary of the results.
Here we develop a theory to estimate the probability that reads are
mismapped due to a seeding failure. We describe the properties of simple
exact seeds, skip-seeds used in Bowtie2 and MEM (Maximal Exact Match)
seeds used in BWA-MEM. 
% Innovation.
The major innovation is to use concepts from analytic combinatorics to
represent reads as abstract sequences, from which we estimate the
probabilities of interest.
% Durability.
We provide several algorithms, which combined together give a practicial
solution to the problem of calibrating seeding heuristic for short reads.
% Scope.
These results significantly improve current mapping algorithms and lay the
foundation of a general strategy to tackle sequence alignment problems.
\end{abstract}

%---------------------------------------------------------------
%---------------------------------------------------------------

\section{Introduction}

\subsection{Mapping reads to genomes}

Suppose that a DNA fragment drawn at random from a known
genome\footnote{We will assume that the sequence of the genome is known
exactly and with absolute certainty. This is not the case in practical
application, but this assumption will allow us to develop a theory
centered on the read.} is sequenced with an imperfect instrument; how can
we identify the location of the fragment in the genome?

We will refer to this question as the \emph{true location problem}.
It has applications in countless domains, like discovering
disease-causing mutations; diagnosing viral, bacterial or fungal
infections; diagnosing cancer and choosing appropriate therapies;
selecting breeds of interest in agriculture; tracing human migrations from
ancient bones; identifying the donor of a biological sample; diagnosing
genetic diseases at birth; finding compatible graft donors \textit{etc}.

At first sight, whether a read read can be mapepd to its correct location
seems to depend only on its length and on the error rate of the sequencer.
However, there is another important factor to consider: most genomes
contain repeated sequences, \textit{i.e.} relatively large subsequences
that are present at more than one location. For instance, we cannot map
the fragment with certainty if there exists an exact copy of the sequence
somewhere else in the genome, because we cannot know which of the two
copies was sequenced.

This case is rare though. Repeated sequences are in general not identical,
but merely homologous\footnote{By that we mean that drawing nucleotides
uniformly at random is very unlikely to produce sequences with the
observed degree of identity, meaning that the similarity is due some
biological duplication process.}. So when the DNA fragment is duplicated,
we can still map it, as long as we can distinguish it from its duplicates.
This in turn depends on their similarity with the target fragment and on
the error rate of the sequencer.

Since repeats play a central role in the problem at hand, we give the
terms ``targets'' and ``duplicates'' a meaning that will facilitate the
exposition of the theory.

\begin{definition}
The target is the DNA fragment that was sequenced. Duplicates are
sequences of the genome that are homologous to the target.
\end{definition}

Because sequencing errors can convert the sequence of the target into one
of its duplicates, the \emph{true} location of a DNA fragment is not
always the \emph{best} (as measured by the identity between the sequence
and the candidate location). This leads to the following question: how can
we identify the \emph{best} location of the fragment in the genome?

We will refer to this task as the \emph{best location problem}. It amounts
to finding the optimal alignment between two sequences, and for this
reason has received substantial attention in bioinformatics. For
simplicity we will assume that the true location is always the best. We
will relax this assumption in section section~\ref{sec:not_the_best}, but
until then we will consider that the solution of the true location problem
is the best location.

\subsection{Seeding heuristics}
\label{sec:seedheur}

There exists algorithms to solve the best location
problem~\cite{pmid7265238,pmid5420325}, but they are too slow to process
the large amount of data generated by modern sequencers. Instead, one uses
heuristic methods, \textit{i.e.} algorithms that run fast, but that may
return an incorrect result~\cite{Waterman1984}.

The most popular heuristic for the best location problem is a filtration
method called ``seeding''. The principle is to first identify short
regions of high similiarity between the read and the genome, and then run
an exact alignment algorithm at the candidate locations. This is for
instance the general strategy of the popular alignment algorithm
BLAST~\cite{pmid2231712}.

If seeding is fast, we can quickly zoom in on a small set of candidate
locations and reduce the input size of the (slow) alignment algorithm. The
disadvantage is that the target location is not always in the candidate
set. When that happens, the read cannot be mapped to the correct location
because it has been filtered out at the seeding step.

The slowness of the alignment algorithm imposes a trade-off between speed
and accuracy: If the seeding step is set to produce many candidates, it is
likely to retrieve the candidate. The downside is that this will take
long. Conversely, if the seeding step is set to produce few candidates,
the process will run faster but the target is less likely to be in the
candidate list.

Importantly, the trade-off depends on the seeding method. This means that
we can develop faster mapping algorithms at no cost for accuracy, as long
as we can find better seeding strategies. Progress on this line of
research has largely benefited from the improvement of computer hardware
and from the development of indexing algorithms. Literature on this topic
is already abundant. Some examples of seeding algorithms are given in
references~\cite{sun2005designing, pmid11934743, xu2006optimizing,
kucherov2005multiseed, brejova2003vector, pmid18684737, pmid15359419}.
References~\cite{pmid16533404} and~\cite{pmid20460430} present high-level
comparisons of different designs, and reference~\cite{navarro2001guided}
gives a global overview of filtration methods in pattern matching.

\subsection{The two types of seeding failure}
\label{sec:twotypes}

Filtering methods are considered to fail whenever the target is not in the
candidate set. For seeding heuristics, we need to distinguish two
different kinds of failure. In the first kind, the candidate set contains 
a duplicate of the target but does not contain the target itself; in the
second kind, the candidate set contains neither.

The distinction is important because duplicates of the target are similar
to the read (due to their similarity to the target), so a failure of the
first kind is difficult to distinguish from a success. A failure of the
second kind is easier to recognize because in this case the best hit is
not similar to the read.

Before going further, we introduce some vocabulary that will be useful to
simplify the exposition of the theory.

\begin{definition}
The seeding process is said to be
\begin{enumerate}[i)]
\item ``on target'' if the candidate set contains the target,
\item ``off target'' if the candidate set contains a duplicate but
not the target, 
\item ``null'' if the candidate set contains neither.
\end{enumerate}
\end{definition}

If a seed-based mapping algorithm returns an incorrect location, then it
must be the case that the best location was never tested in the second
phase, otherwise it would have been singled out by the exact alignment
algorithm.

This observation is fundamental: in order to estimate the mismapping rate
of an algorithm, we must know how often the associated seeding heuristic
is off target. For lack of theoretical framework, this information is
missing for nearly all existing seeding methods, meaning that current
mappers cannot be properly calibrated.

The focus of this article is to develop a theory to estimate the
probability that seeding is off target under the most popular seeding
heuristics for short reads. Our previous work pioneered a method to
compute seeding probabilities but it did not distinguish off-target from
null seeding~\cite{filion2017analytic,filion2018analytic}, and therefore
did not provide a way to control the mismapping rate. Other authors
investigated the reliability of mapping algorithms~\cite{pmid23872968},
but they focused on the probability that random sequences have a single
hit, recognizing that solving the problem requires taking into
consideration the repeat structure of the genome. Here we address this
problem in general terms and we provide a practical way to estimate the
mismapping rate in complex genomes.

The rest of this document is organized to present the content in a
didactic manner. The order of the examples is meant to facilitate the
aquisition of the main concepts, meaning that we sometimes go back and
forth between different types of seeding problems. We first give some
background on the different types of seeding methods considered here, we
then introduce the computational strategy based on transfer matrices, and
then delve into each specific seeding method. Finally, we provide a
statistical method to estimate the parameters of the problem from the
data.

\section{Seeds}

The term ``seed'' does not have exactly the same meaning according to
different authors. It can designate a part of the read, a part of the
genome, a particular sequence motif or a structured pattern of matches.
Also, the term does not always imply an exact match. For instance, the
algorithm PatternHunter~\cite{pmid11934743} uses ``spaced seeds''
that tolerate mismatches\footnote{Spaced seeds are covered
by patent US20070088510A1.}. To avoid potential confusion,
we will adopt the following convention.

\begin{definition}
A seed is a subsequence of the read that is above a certain size $\gamma$
(defined by the context of the problem) and that has at least one perfect
match in the reference genome.
\end{definition}

When a seed matches a particular location of the genome, we will say that
it is a ``seed for that location''. For instance, we will often use
expressions such as ``a seed for the target'' or ``a seed for a
duplicate''.

This definition presents a computational challenge: to know if a given
subsequence is a seed we need to know if it exists somewhere in the
genome. This is a non trivial problem in itself, but fortunately there
exists practical methods to solve it, even when the reference genome is
very large.

These algorithms are crucial for the present theory, but describing them
in depth is outside the scope of this document. Let us just mention that
all the methods belong to a family known as exact offline string matching
algorithms, where ``offline'' means that sequences are looked up in an
index instead of the genome itself. Online methods can be used when the
reference genome is not indexed~\cite{faro2013exact}, but this case is of
little relevance in the present context.

The index is usually a hash table or a variant of the so-called
FM-index~\cite{pmid20460430}. Hash tables are typically used to index
$k$-mers, which makes them useful to search seeds of set size $k$,
(see~\cite{pmid30346548} for a recent benchmark of $k$-mer hashing
algorithms). In contrast, the FM-index has no set size so it can be used
to search seeds of different lengths~\cite{ferragina2000opportunistic,
ferragina2005indexing}. It is a compact data structure based on the suffix
array~\cite{manber1993suffix} and on the Burrows-Wheeler
transform~\cite{burrows1994block}, emulating a suffix trie with a much
smaller memory footprint. Queries in the FM-index are performed with an
algorithm known as the \emph{backward
search}~\cite{ferragina2000opportunistic}.

Other methods can be efficient (\textit{e.g.}, running a bisection on the
suffix array~\cite{dobin2013star}) but the FM-index is currently the most
popular choice for seeding methods. For MEM seeds described below, it is
even the only practical option~\cite{pmid24336412, pmid25399029,
pmid23349213, pmid19389736}, so we implicitly assume that seeding is
performed by running the backward search. This will only have practical
consequences in section~\ref{sec:est}, where we propose a method to
estimate the parameters that are required to compute the probabilities of
interest. For the rest, we simply assume that seeds are known at all times
without ambiguity because this problem has several practical solutions.

\subsection{Exact seeds}

Exact seeds have a fixed size $\gamma$. When using exact seeds, all the
perfect matches of size $\gamma$ are collected to constitute the candidate
set. This seeding heuristic was used in the first version of
BLASTN~\cite{pmid2231712}, but is no longer used because it produces many
overlapping matches in the same regions, as we are about to illustrate.

Fig.~\ref{fig:exact_seed_example} shows the example of a read produced
by an imperfect sequencer. The sequenced DNA fragment comes from a region
of the genome that has three duplicates (there are four homologous
sequences in total), and the read has three miscalled nucleotides
(indicated by stars).

\begin{figure}[h]
\centering
\includegraphics[scale=1]{exact_seed_example.pdf}
\caption{\textbf{Example of exact seeds.}
A sequenced DNA fragment (Read) is shown above the actual molecule
(Target), and above homologous sequences also present in the genome
(Duplicates). Sequencing errors are indicated by a star, each marked
\texttt{T} was actually an \texttt{A} that was misread. The mismatches
between the genomic sequences and the read are indicated in bold. The
exact seeds of size $\gamma = 7$ are indicated as horizontal grey lines
above the read. The matching regions in the genomic sequences are
shadowed in grey. Several overlapping seeds accumulate at the center of
the read, which is typical for this seeding method.}
\label{fig:exact_seed_example}
\end{figure}

The exact seeds with $\gamma=7$ are shown above the sequence of the read.
Observe that erroneous nucleotides can belong to exact seeds because they
sometimes match one of the duplicates. For instance, the first sequencing
error matches all the duplicates and belongs to a seed for the first one.
However, erroneous nucleotides that are mismatches for \emph{all} the
duplicates cannot belong to a seed. This is the case of the second
sequencing error in this example, where there is a local deficit of
seeds.

Thie middle of the read shows some clutter where consecutive seeds
match consecutive sequences at the same location. This is typical for
exact seeds and constitutes a nuisance for the implementation. Indeed, it
is a waste of computer resources to discover matches in sequences that are
already in the candidate set. In addition, this seeding method is not
particularly sensitive~\cite{pmid11934743} so it is used only for specific
applications. Nevertheless, it will be extremely useful for the
development of the present theory.

\subsection{Skip seeds}

Skip seeds have a fixed size $\gamma$, but unlike exact seeds they cannot
start at every nucleotide. Instead, a fixed amount of nucleotides is
skipped between every seed. This is a way to reduce the amount of
overlapping matches at the same location, at the cost of sensitivity. This
seeding heuristic is the core of Bowtie2~\cite{pmid22388286}, where seeds
have size $\gamma=16$ and are separated by 10 nucleotides (9 positions are
skipped). We will refer to seeds where $n$ nucleotides are skipped as
``skip-$n$ seeds''. For instance, Bowtie2 uses skip-9 seeds.

\begin{figure}[h]
\centering
\includegraphics[scale=1]{skip_seed_example.pdf}
\caption{\textbf{Example of skip seeds.}
The sequences and the annotations are the same as in
Fig.~\ref{fig:exact_seed_example}, but here we use skip-1 seeds. In other
words, seeds can never start at nucleotides 2, 4, 6 \textit{etc}. To
highlight the difference with Fig.~\ref{fig:exact_seed_example}, the
missing seeds are represented by dotted lines. Skipping nucleotides
reduces the clutter at the center of the read, but it also removes the
only seed for the second duplicate.}
\label{fig:skip_seed_example}
\end{figure}

Fig.~\ref{fig:skip_seed_example} shows what happens when exact seeds are
replaced by skip-2 seeds on the read of Fig.~\ref{fig:exact_seed_example}.
Here the size is still $\gamma=7$ but 1 nucleotide is skipped between
seeds. This amounts to removing every second seed. The consequence is that
there are fewer overlapping matches at the center of the read, but the
only seed for the second duplicate is lost. This is a rather positive
outcome because there is one off-target location fewer in the candidate
set, but the same might happen to the target.

It is clear that skipping nucleotides reduces the sensitivity of the
seeding step, but to what extent? One could test this empirically, but the
answer depends on the seed length, the number of nucleotides that are
skipped, the error rate of the sequencer and the size of the reads. The
present theory will allow us to make general statements about the
performance of skip seeds against exact seeds in different contexts.

\subsection{MEM seeds}

MEM seeds (where MEM stands for Maximal Exact Match) are somewhat harder
to define. Unlike exact seeds and skip seeds, their size is variable as we
are about to see. They are used in BWA-MEM~\cite{li2013aligning} where
they give good empirical results (but note that the actually seeding
scheme uses more complex Super Maximal Exact Matches). To describe MEM
seeds, let us first focus on MEMs.

\begin{definition}
A Maximal Exact Match (MEM) is a subsequence of the read that is present
in the reference genome and that cannot be extended --- either because the
read ends or because the extended subsequence is not in the genome.
\end{definition}

A \emph{MEM seed} is simply a MEM of minimum size $\gamma$.
Fig.~\ref{fig:MEM_example} shows what happens when exact seeds or skip
seeds are replaced by MEM seeds on the read of
Fig.~\ref{fig:exact_seed_example}. Observe that the clutter at the center
of the read has disappeared because consecutive matches are fused into a
single MEM seed.

\begin{figure}[h]
\centering
\includegraphics[scale=1]{MEM_example.pdf}
\caption{\textbf{Example of MEM seeds.}
The sequences and the annotations are the same as in
Fig.~\ref{fig:exact_seed_example}, but here we use MEM seeds of minimum
size $\gamma=7$. The clutter at the center of the read has disappeared and
there is at least one seed for each sequence.}
\label{fig:MEM_example}
\end{figure}

Two consecutive MEM seeds can overlap, in which case they always match
distinct sequences of the genome (otherwise neither of them would be a MEM
seed). There does not have to be any overlap though, because a nucleotide
can be a mismatch against \emph{all} the sequences, like the second read
error for instance.

Note that a MEM does not always match a single sequence of the genome. For
instance, the rightmost MEM seed matches two distinct genomic sequences.
This case motivates the following definition, which will play an important
role later.

\begin{definition}
A \emph{strict} MEM seed has a single match in the genome.
A \emph{shared} MEM seed has several matches in the genome.
\end{definition}

Compared to seeds of fixed size, MEM seeds have two counter-intuitive (and
undesirable) properties. The first is that there are cases where there
cannot be any good seed, even when changing the sensitivity parameter
$\gamma$. Fig.~\ref{fig:full_masking_example} shows such an example. Even
though there is a single sequencing error, the read has no MEM seed for
the target. Lowering $\gamma$ does not change this, so there is no way to
discover the true location in this case (even though it is the best
location).

\begin{figure}[h]
\centering
\includegraphics[scale=1]{full_masking_example.pdf}
\caption{\textbf{Issues with MEM seeds: inaccessible targets.}
The read, the MEM seeds and the sequences are represented as in
Fig.~\ref{fig:MEM_example}. The MEM seeds matching the two incorrect
locations at the bottom effectively hide the target, so it cannot
be discovered. This can occur even when the true location is the
best and when there is a single error on the read.}
\label{fig:full_masking_example}
\end{figure}

The second counter-intuitive property is that shortening a read can
sometimes generate a MEM seed for the target. Fig.~\ref{fig:short_vs_long}
shows an example of this case. There is no MEM seed for the target, but
there would be one if the read were two nucleotides shorter on the right
side. Indeed, in this case there would be a shared MEM seed matching the
first and the second locations (provided $\gamma \leq 12$).

\begin{figure}[h]
\centering
\includegraphics[scale=1]{short_vs_long_example.pdf}
\caption{\textbf{Issues with MEM seeds: too long reads.}
The read, the MEM seeds and the sequences are represented as in
Fig.~\ref{fig:MEM_example}. There would be a good seed (shared) if the
read were two nucleotides shorter. The true location is hidden by
the last two nucleotides.}
\label{fig:short_vs_long}
\end{figure}

These examples show that MEM seeds can perform worse than seeds of fixed
size. MEM seeds yield fewer candidates and therefore speed up the mapping
process, but the question is at what cost for accuracy? The theory
developed here will allow us to compute the probability that a read has
no MEM seed for the target and thus that the true location is missed at
the seeding stage.

\section{Model and strategy}
\label{sec:symbolic}

\subsection{Sequencing errors and divergence of the duplicates}
\label{sec:error}

We now need to model the sequencing and duplication processes so that
we can compute the probabilities of the events of interest. We assume that
the sequencing instrument has a constant substitution rate $p$, and that
insertions and deletions never occur. When a substitution occurs, we
assume that the instrument is equally likely to output any one of the
remaining three nucleotides. This corresponds more or less to the error
spectrum of the Illumina sequencing technology~\cite{pmid21576222}.

Next, we assume that the target sequence has $N \geq 0$ duplicates, so
that there are $N$ off-target sequences. We further assume that
duplication happened instantaenously at some point in the past and that
all $N+1$ sequences diverge independently of each other at a constant
rate. In other words, we ignore the complications due to the genealogy of
the duplication events and we simply assume that at each nucleotide
position, any given duplicate is identical to the target with probability
$1-\mu$. If it is not, we assume that the duplicate sequence can be any of
one the remaining three nucleotides (\textit{i.e.} each is found with
probability $\mu/3$).

Note that read errors are always mismatches against the target (because we
assume that the target is the true sequence), and they match each
off-target sequence with probability $\mu/3$. In contrast, correct
nucleotides are always matches for the target, and they match each
off-target sequence with probability $1-\mu$.

Before going further, we also need to move a practical consideration out
of the way. Seeds can match any sequence of the genome, not just the
target or the duplicates. However, we will ignore matches in the rest of
the genome because such random matches are unlikely to cause a mapping
error when seeding is off target, contrary to matches in duplicates.
Neglecting those will greatly simplify the exposition of the theory
without loss of generality. We will explain in
section~\ref{sec:short_MEMs} how to deal with this practical case and and
how to identify those matches as off target. Until then, we will consider
that the target and the duplicates are the only sequences in the reference
genome.

\subsection{Weighted generating functions}

The central object of analytic combinatorics is the generating function,
and for our purpose we will use a special kind known as \emph{weighted
generating function}.

\begin{definition}
\label{def:GF}
Let $\mathcal{A}$ be a set of combinatorial objects such that $a \in
\mathcal{A}$ has a size $|a| \in \mathbb{N}$ and a weight $w(a) \in
\mathbb{R}^+$. The weighted generating function of $\mathcal{A}$ is
defined as

\begin{equation}
\label{eq:GF1}
A(z) = \sum_{a \in \mathcal{A}} w(a) z^{|a|},
\end{equation}
Expression (\ref{eq:GF1}) also defines a sequence $(a_k)_{k \geq 0}$ such
that

\begin{equation*}
A(z) = \sum_{k=0}^\infty a_k z^k.
\end{equation*}

By definition $a_k = \sum_{a \in A_k}w(a)$, where $A_k$ is the class of
objects of size $k$ in $\mathcal{A}$. The number $a_k$ is the
total weight of objects of size $k$.
\end{definition}

To give an example, assume that a particular symbol, say $\Downarrow$, has
a probability of occurence equal to $p$. The weighted generating function
of the word containing only this symbol is $pz$. The weight of the word is
its probability (here equal to $p$) and the size is its length in number
of symbols (here 1).

In this document we will focus on the weighted generating function $A(z)$
of the set $\mathcal{A}$ of reads that do not have on-target seeds
(\textit{i.e.} reads for which seeding is either off target or null). The
weight of a read is its probability of occurrence and the size $k$ is its
number of nucleotides. The coefficient $a_k$ is thus the proportion of
reads of size $k$ that do not have an on-target seed, which is the
quantity of interest.

The motivation for introducing weighted generating functions is that
operations on combinatorial objects translate into operations on their
weighted generating functions. If $A(z)$ and $B(z)$ are the weighted
generating functions of two mutually exclusive sets $\mathcal{A}$ and
$\mathcal{B}$, the weighted generating function of $\mathcal{A} \cup
\mathcal{B}$ is $A(z) + B(z)$, as evident from expression (\ref{eq:GF1}).
Size and weight can be defined for pairs of objects in $\mathcal{A} \times
\mathcal{B}$ as $|(a,b)| = |a| + |b|$ and $w(a,b) = w(a)w(b)$. In other
words the sizes are added and the weights are multiplied.  With this
convention, the weighted generating function of the Cartesian product
$\mathcal{A} \times \mathcal{B}$ is $A(z)B(z)$. This simply follows from
expression (\ref{eq:GF1}) and from

\begin{equation*}
A(z)B(z) =
\sum_{a\in \mathcal{A}}w(a)z^{|a|} \sum_{b\in \mathcal{B}}w(b)z^{|b|}
= \sum_{(a,b) \in \mathcal{A} \times \mathcal{B}} w(a)w(b)z^{|a|+|b|}.
\end{equation*}

These two operations are all we need in order to compute the weighted
generating functions of the reads of interest. Addition corresponds to
creating a new family by merging reads from families $\mathcal{A}$ and
$\mathcal{B}$. Multiplication corresponds to creating a new family by
concatenating reads from families $\mathcal{A}$ and $\mathcal{B}$.

\subsection{Analytic representation}

The analytic combinatorics framework relies on a strategy referred to as
the \emph{symbolice method}~\cite{sedgewick2013introduction}. The idea is
to combine simple objects into more complex objects. Each combinatorial
operation on the objects corresponds to a mathematical operation on their
weighted generating functions. One can thus obtain the weighted generating
function of complex objects, whose coefficients $a_k$ are the
probabilities of interest.

As explained in~\cite{filion2017analytic} and~\cite{filion2018analytic},
we recode the reads in alphabets of custom symbols and we specify a
construction plan of the reads using a so-called \emph{transfer matrix}
$M(z)$. The transfer matrix specifies which types of segments can follow
each other in the reads of interest: the entry at coordinates $(i,j)$ is
the weighted generating function of segments of type $j$ that can be
appended to segments of type $i$.

The entry of $M(z) + M(z)^2 + M(z)^3 + \ldots = M(z) \cdot (I-M(z))^{-1}$
at coordinates $(i,j)$ is the weighted generating function of all the
reads that end with a segment of type $j$ and that can be appended to a
segment of type $i$. If segments of types $i$ and $j$ are exactly those
that can be placed at the beginning and at the end of the reads, then this
entry is the weighted generating function of all the reads without
on-target seed. The examples below will clarify the key steps of this
strategy.

A complete description of the strategy to compute seeding probabilities
with the symbolic method can be found in~\cite{filion2017analytic,
filion2018analytic}. The interested readers can also read more about
analytic combinatorics in the popular
textbooks~\cite{flajolet2009analytic, sedgewick2013introduction}.

\subsection{Example 1: on-target exact seeds}
\label{sec:example_exact}

We highlight the strategy above with an example that will turn out to be
central for the development of the theory. In addition, it is simple
enough to provide a gentle introduction to the general methodology. This
example was described in detail in~\cite{filion2017analytic} and
in~\cite{filion2018analytic}, but we repeat it here with a different
formalism to fit the present manuscript.

The first step is to note that the nucleotide sequence of the read is
irrelevant. Indeed, the read has an on-target seed if and only if it
contains a stretch of $\gamma$ nucleotides without error. For this reason,
we recode reads as sequences of correct or erroneous nucleotides.

We define the mismatch alphabet $\mathcal{A}_0 = \{\square, |,
\Downarrow\}$, where $\square$ represents a correct nucleotide,
$\Downarrow$ represents an erroneous nucleotide and $|$ is special symbol
appended after the last nucleotide to mark the end of the read.

This recoding allows us to partition the read in \emph{segments} defined
as a sequence of 0 or more symbols $\square$ followed by a symbol
$\Downarrow$. The last part of the read is called the tail and it is
terminated by the symbol $|$ instead of $\Downarrow$. The symbol $|$ does
not correspond to a nucleotide so it has size 0. Since the decomposition
is unique, we can view a read as a sequence of segments with a tail,
instead of a sequence of nucleotides.

Fig.~\ref{fig:simple} shows an example of decomposition in segments.
On-target seeds cannot contain sequencing errors, therefore they must be
completely embedded inside a segment or inside the tail. So the size of
the segments and of the tail indicates whether the read contains an
on-target seed or not.

\begin{figure}[h]
\centering
\includegraphics[scale=0.85]{sketch_simple.pdf}
\caption{\textbf{The mismatch encoding.}
An example read is represented in the mismatch alphabet. The symbol
$\Downarrow$ represents a mismatch against the target (it is an erroneous
nucleotide) and the symbol $\square$ represents a match (it is a correct
nucleotide). The symbolic sequence of the target is represented below,
where an open square stands for a match and a closed square stands for a
mismatch.}
\label{fig:simple}
\end{figure}

The probability of occurrence of the symbol $\Downarrow$ is $p$, the error
rate of the sequencer, and the probability of occurrence of the symbol
$\square$ is thus $1-p = q$. Both symbols have size 1, so their respective
weighted generating function are $pz$ and $qz$. Using the rule for
concatenation, we see that the weighted generating function of a segment
with $i$ symbols $\square$ is $(qz)^ipz$. The symbol $|$ has size 0, so
the weighted generating function of a tail segment with $i$ symbols
$\square$ is $(qz)^i$.

The key insight is that reads without on-target seed are exactly the reads
that are made of segments and tail with fewer than $\gamma$ symbols
$\square$. The weighted generating function of such segments is $pz \cdot
\big( 1+qz+\ldots+(qz)^{\gamma-1} \big)$, and that of the tail is
$1+qz+\ldots+(qz)^{\gamma-1}$. This gives a construction plan that can be
encoded in a transfer matrix.

Since there are two kinds of objects, the segments and the tails, the
dimension of the transfer matrix is $2 \times 2$. A segment can be
followed by another segment or by the tail. The tail cannot be followed by
anything. The expression of the transfer matrix $M_0(z)$ is thus
\begin{equation*}
\begin{blockarray}{ccc}
   & \scriptstyle{\Downarrow} & \scriptstyle{|} \\
\begin{block}{c[cc]}
\scriptstyle{\Downarrow} & pz \cdot \big( 1+qz+\ldots+(qz)^{\gamma-1}
\big)  & 1+qz+\ldots+(qz)^{\gamma-1} \\
\scriptstyle{|} & 0 & 0 \\
\end{block}
\end{blockarray}.
\end{equation*}

The types of objects are identified by their terminator, \textit{i.e.} by
their last symbol. Segments are identified by the terminator $\Downarrow$
and tails by the terminator $|$. The terminators are written in the
margins of the transfer matrix for readability.

We can compute the matrix $M_0(z) + M_0(z)^2 + \ldots = M_0(z) \cdot
(I-M_0(z))^{-1}$ and extract the entry of interest, which is the top right
term --- associated with terminators $\Downarrow$ and $|$. To see why,
observe that every read can be prepended by a segment and only by a
segment (not by a tail). Thus, reads are precisely the sequences of
segments that can follow the symbol $\Downarrow$ and that are terminated
by a tail, whose weighted generating function is the top right entry of
the matrix. It is easy to check that this term is equal to
\begin{equation}
\label{eq:F}
\frac{1+qz+\ldots+(qz)^{\gamma-1}}
  {1-pz \big(1+qz+\ldots+(qz)^{\gamma-1} \big)} = a_0 + a_1z + a_2z^2 +
\ldots
\end{equation}

The coefficients of the Taylor series $a_0 + a_1z + a_2z^2 + \ldots$ are
unknown. But by construction, $a_k$ is the quantity of interest,
\textit{i.e.} it is the probability that a read of size $k$ does not
contain an on-target seed, so we now need to extract the coefficients
from expression (\ref{eq:F}). There are several methods to do so; the one
we choose here is to build a recurrence equation to compute $a_k$ from the
previous terms of the series $a_0, a_1, \ldots, a_{k-1}$. For this, we
observe that for $z \neq 1/q$, equation (\ref{eq:F}) is equivalent to
\begin{equation*}
(1-qz) \left(1-pz \big(1+qz+\ldots+(qz)^{\gamma-1} \big)\right)
(a_0 + a_1z +a_2z^2 + \ldots) = 1-(qz)^\gamma.
\end{equation*}
Balancing the terms of same degree on both sides of the equality leads
to the following relationship between the coefficients of the series:
\begin{equation}
\label{eq:recur}
a_k = 
\begin{cases}
1            &\quad\text{if } k < \gamma, \\
1 -pq^\gamma &\quad\text{if } k = \gamma, \\
a_{k-1} -pq^\gamma \cdot a_{k-\gamma-1} &\quad\text{otherwise.}
\end{cases}
\end{equation}

The terms of interest can be computed recursively using expression
(\ref{eq:recur}). This approach is very efficient because every iteration
involves at most one multiplication and one subtraction. Also, the default
floating-point arithmetic on modern computers gives sufficient precision
to not worry about numeric instability for the problems considered here
(we rarely need to compute those probabilities for reads above 500
nucleotides).

This example shows how the symbolic approach yields a non-trivial and yet
simple algorithm to compute the probability that a read of size $k$ does
not contain an on-target exact seed.

\subsection{Example 2: on-target skip seeds}
\label{sec:example_skip}

Here we go through another essential example, namely we devise a method to
compute the probability that a read contains no on-target skip seed. Using
the same strategy as in the previous example, we start by recoding the
reads using a specialized alphabet to solve this problem.

Once again, we need to know whether a nucleotide is a sequencing error,
but this time we also need to know its phase in the repeated cycles of
skipped positions. For this, we define the skip-$n$ alphabet
$\mathcal{A}_n = \{\square, |, \Downarrow_0, \Downarrow_1, \ldots,
\Downarrow_n \}$. As before, the symbol $\square$ represents a correct
nucleotide and the symbol $|$ is a terminator appended to the end of the
read. The symbols $\Downarrow_i$ ($i \in \mathbb{N}$) represent sequencing
errors and $i$ indicates the number of nucleotides till the next
non-skipped position (\textit{i.e.} $i=0$ for nucleotides immediately
before a non-skipped position and $i=n$ for nucleotides at a non-skipped
position).

The segments in this alphabet are defined as sequences of 0 or more
symbols $\square$ followed by any of the symbols $\Downarrow_i$. The tail
is again defined as a sequence of 0 or more symbols $\square$ followed by
the symbol $|$. Given that this decomposition is unique, we can again view
a read as a sequence of segments with a tail. The example of
Fig.~\ref{fig:simple} is shown again in Fig.~\ref{fig:skip}, where
segments in the mismatch alphabet have been replaced by segments in the
skip-3 alphabet.

\begin{figure}[h]
\centering
\includegraphics[scale=0.85]{sketch_skip.pdf}
\caption{\textbf{The skip encoding.}
The read of Fig.~\ref{fig:simple} is represented in the skip-3 alphabet.
The symbols $\Downarrow_i (i = 1,2,3)$ represent mismatches against the
target (they are erroneous nucleotides) and the symbol $\square$
represents a match (it is a correct nucleotide). The positions of the
reads with a vertical bar indicate non-skipped positions (the potential
start of a seed). The number $i$ in $\Downarrow_i$ indicates the distance
to the next non-skipped position in number of nucleotides. For $i=0$ the
next nucleotide is at a non-skipped position. Other graphical
representations are as in Fig.~\ref{fig:simple}.}
\label{fig:skip}
\end{figure}

The probability of occurrence of a sequencing error is $p$, so every
symbol $\Downarrow_i$ has the same weighted generating function $pz$ ---
implicitly assuming that the next non-skipped position is at distance $i$,
otherwise the weighted generating function is 0. The weighted generating
function of the symbol $\square$ is again $qz$, and so the weighted
generating function of a segment with $i$ symbols $\square$ is $(qz)^ipz$.
Likewise, the weighted generating function of a tail with $i$ symbols
$\square$ is $(qz)^i$.

Reads that do not contain any on-target skip seed can contain segments
with $\gamma$ or more symbols $\square$, so this case is more complex than
the previous one. For instance, if there is a sequencing error $i$
nucleotides before the next non-skipped position, the status of the next
$i$ nucleotides does not have any influence on the presence of an
on-target seed. Indeed, there is no on-target seed upstream of the error
(by construction) and the next seed is scheduled to start $i$ nucleotides
downstream, so there can be up to $\gamma+i-1$ symbols $\square$ in a row.
In order to enforce the absence of on-target seeds, we thus have to adjust
the maximum size of the segments depending on the preceding terminator.

In the skip-$n$ alphabet, reads are composed segments terminated by the
$n+1$ symbols $\Downarrow_0, \Downarrow_1, \ldots, \Downarrow_n$, plus the
tail. The dimensions of the transfer matrix are thus $(n+2) \times (n+2)$.

As mentioned above, after the terminator $\Downarrow_i$, we can append
segments with up to $\gamma+i-1$ symbols $\square$. The terminators of
those $\gamma+i$ possible segments are distributed according to the laws
of the arithmetic modulo $n+1$. For $i > 0$ for instance, if the next
segment has no symbol $\square$ then it is terminated by
$\Downarrow_{i-1}$. But the same goes if the segment has $n+1$ symbols
$\square$ (assuming $n+1 < \gamma$). Specifying the transfer matrix thus
involves a substantial amount of bookkeeping. One can check that the
expression of the transfer matrix $M_n(z)$ is
\begin{equation*}
\begin{blockarray}{cccccc}
   & \Dn{0} & \Dn{1} & \ldots & \Dn{n} & \nd \\
\begin{block}{c[ccccc]}
\Dn{0} & H_{0,0}(z) & H_{0,1}(z) & \ldots & H_{0,n}(z) & J_0(z) \\
\Dn{1} & H_{1,0}(z) & H_{1,1}(z) & \ldots & H_{1,n}(z) & J_1(z) \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\Dn{n} & H_{n,0}(z) & H_{n,1}(z) & \ldots & H_{n,n}(z) & J_n(z) \\
\nd & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}
\begin{eqnarray}
H_{i,j}(z) &=& pz \cdot (qz)^x \cdot \big( 1 + (qz)^{n+1} +
  \ldots + (qz)^{m(n+1)} \big), \\
\notag
  &\;& \text{where } x = i-j-1 \modulo{n+1},
  \text{ and } m = \left\lfloor
  \frac{\gamma+i-1-x}{n+1} \right\rfloor, \\
J_i(z) &=& 1 + qz + (qz)^2 + \ldots + (qz)^{\gamma+i-1}.
\end{eqnarray}

The weighted generating function of interest is the top right entry of the
matrix $M_n(z) + M_n(z)^2 + \ldots = M_n(z)\cdot(I-M_n(z))^{-1}$. To see
why, observe that at the start of every read, the next nucleotide is a
non-skipped position, so every read can be prepended by segments
terminated with $\Downarrow_0$ and only by those. Thus, reads are
precisely the sequences of segments that can follow the symbol
$\Downarrow_0$ and that are terminated by a tail, whose weighted
generating function is the top right entry of the matrix.

$M_n(z)$ is too complex to compute a close expression of
$M_n(z)\cdot(I-M_n(z))^{-1}$. We return to the definition $M_n(z) +
M_n(z)^2 + \ldots$ and we observe that the terms $M_n(z)^{k+2},
M_n(z)^{k+3}, \ldots$ have no influence on the coefficients of the
weighted generating function $a_0, a_1, \ldots, a_k$.

Indeed, a read of size $k$ has at most $k$ segments (plus one obligatory
tail). Since $M(z)^{s+1}$ contains the weighted generating functions of
reads with exactly $s$ segments and one tail, $a_k$ cannot depend on
$M(z)^{k+2}, M(z)^{k+3}$. More formally, we can prove by induction that
all the entries of $M(z)^k$ are divisible by $z^{k-1}$, showing that the
contribution of $M(z)^{k+2} + M(z)^{k+3} + \ldots$ to $a_0 + a_1z + \ldots
+a_kz^k$ is strictly $0$.

So we can compute the matrix $M_n(z) + M_n(z)^2 + \ldots + M_n(z)^{k+1}$,
extract the entry of interest and then compute the terms of the Taylor
expansion up to order $k$. But we can do better. Since we are only
interested in the coefficients up to order $k$, we can perform all
algebraic operations on truncated polynomials of order $k$, \textit{i.e.},
we discard the coefficients of order $k+1$ or greater when multiplying two
polynomials.

But we can do even better. A read with $s$ segments contains $s$ errors,
so all the entries of $M_n(z)^{s+1}$ are dominated by $p^s$ and they
rapidly converge to $0$ as $s$ increases. Instead of computing the matrix
$M_n(z) + M_n(z)^2 + \ldots + M_n(z)^{k+1}$, we can interrupt the
summation after a certain power of $M_n(z)$ because the terms become
negligible.

The number of errors $X$ in a read of size $k$ has a Binomial
distribution $X \sim B(k,p)$. From~\cite{arratia1989tutorial} we can bound
the probabilities of the tail with the expression

\begin{equation}
\label{eq:bound}
Pr(X \geq s) \leq \exp \left( (s-k)\log \frac{k-s}{k(1-p)} -s\log
\frac{s}{kp} \right).
\end{equation}

Using the formula above, we can thus bound the probability that a read has
$s$ segments or more. We compute the matrices $M_n(z) + M_n(z)^2 + \ldots
+M_n(z)^{s+1}$ where the weighted generating functions have been replaced
by truncated polynomials and we extract the entry at the first column and
the last row. When the bound is lower than a set fraction $\varepsilon$ of
the smallest coefficient, we stop the computations. Typically $\varepsilon
= 0.01$ so this method ensures that the probabilities that a read of size
$k$ has no on-target skip seed are accurate to within 1\%.

\begin{remark}
Observe that when $n=0$ the matrix $M_n(z)$ is identical to the matrix
$M_0(z)$ of section~\ref{sec:example_exact}. This is consistent
with the fact that exact seeds are skip-$n$ seeds for which $n=0$.
\end{remark}

\section{Off-target exact seeds}
\label{sec:offexact}

We now turn our attention to the problem of computing the probability that
the seeding process is off target when using exact seeds --- recall from
section~\ref{sec:twotypes} that off-target seeding means that the
candidate set contains a duplicate but not the target.

If there is $N = 0$ duplicate, seeding cannot be off target, it can only
be on target or null. So from here we assume that the target has $N \geq
1$ duplicates. Let $S_0$ denote the event that there is an on-target seed
and let $S_j$ denote the event that there is a seed for the $j$-th
duplicate. We are thus interested in computing $P\big(\overline{S}_0 \cap
(S_1 \cup \ldots \cup S_N)\big)$, where $\overline{S}_j$ denotes the
complement of the event $S_j$. For this, we first observe that
\begin{equation}
\label{eq:pcomp}
P\big(\overline{S}_0 \cap (S_1 \cup \ldots \cup S_N)\big) =
P(\overline{S}_0) - P\big(\overline{S}_0 \cap \overline{S}_1 \cap \ldots
\cap \overline{S}_N\big).
\end{equation}

Since the duplicates are assumed to evolve independently of each other and
through the same mutagenesis process, the events are conditionally
independent (we do not have full independence because $S_0$ depends on
sequencing errors only whereas the events $S_j$ depend on sequencing
errors and on the divergence between duplicates). We can thus write
\begin{eqnarray}
\notag
P(\overline{S}_0 \cap \ldots \cap \overline{S}_N) &=&
  P(\overline{S}_0) \cdot P(\overline{S}_1 \cap \ldots \cap
  \overline{S}_N | \overline{S}_0) \\
\label{eq:Pnull}
  &=& P(\overline{S}_0) \cdot P(\overline{S}_1 | \overline{S}_0)^N =
  P(\overline{S}_0) \cdot \left( \frac{P(\overline{S}_0 \cap
\overline{S}_1)}{P(\overline{S}_0)} \right)^N.
\end{eqnarray}
Combinging the two equations above, we obtain
\begin{equation}
\label{eq:Poff}
P\big(\overline{S}_0 \cap (S_1 \cup \ldots \cup S_N)\big) =
P(\overline{S}_0) - P(\overline{S}_0) \cdot \left( \frac{P(\overline{S}_0
\cap \overline{S}_1)}{P(\overline{S}_0)} \right)^N.
\end{equation}

The probability that seeding is off target thus depends on only two
quantitites. We have already seen how to compute $P(\overline{S}_0)$ in
section~\ref{sec:example_exact} using recursive expression
(\ref{eq:recur}). We now need to find a way to compute $P(\overline{S}_0
\cap \overline{S}_1)$.

\subsection{The dual encoding}
\label{sec:dual}

The event $\overline{S}_0 \cap \overline{S}_1$ is that the read contains
no seed for either the target or the first duplicate --- numbering is
arbitrary here, we simply chose a duplicate and stick to it.
As illustrated in section~\ref{sec:symbolic}, we first recode the reads
using a specialized alphabet to simplify the problem.

It will be useful to consider a slightly more general problem where we
have two sequences of interest labelled $+$ and $-$. One can say that the
$+$ sequence stands for the target and that the $-$ sequence stands for
its duplicate. We then define the dual alphabet $\tilde{\mathcal{A}}_0 =
\{\square, |, \downarrow_{/1}^-, \downarrow_{/2}^-, \ldots,
\downarrow_{/1}^+, \downarrow_{/2}^+, \ldots, \Downarrow\}$. The symbols
$\downarrow_{/i}^-$ signify that the nucleotide is a mismatch against the
$-$ sequence only, the symbols $\downarrow_{/i}^+$ signify that it is a
mismatch against the $+$ only, and the symbol $\Downarrow$ signifies that
it is a mismatch against both. As before, every other nucleotide is
replaced by the symbol $\square$, and the terminator $|$ is appended to
the end of the read. We again define reads as sequences of segments,
except that now the terminators are the symbols $\downarrow_{/j}^-$,
$\downarrow_{/j}^+$ and $\Downarrow$. The tail, as usual, is terminated by
the symbol $|$.

The number $i$ in the symbol $\downarrow_{/i}^-$ indicates the match
length of the $+$ sequence. Likewise, the number $i$ in the symbol
$\downarrow_{/i}^+$ indicates the match length of the $-$ sequence. For
instance the symbol $\downarrow_{/7}^-$ indicates that the nucleotide is a
mismatch against the $-$ sequence, that it is a match for the $+$
sequence, and that the six previous nucleotides were also a match for the
$+$ sequence. The terminators thus encode the local state of the read.

Fig.~\ref{fig:dual} shows an example of read in the dual encoding. The $+$
and $-$ sequences are represented below the read, where the matches are
represented by open squares and the mismatches by closed squares. It is
visible from this example that symbols $\downarrow^-_{/i}$ and
$\downarrow^+_{/i}$ alternate whenever the mismatches are on different
sequences when going from left to right. The symbol $\Downarrow$ occurs
only when a nucleotide is a double mismatch.

\begin{figure}[h]
\centering
\includegraphics[scale=0.85]{sketch_dual.pdf}
\caption{\textbf{Example of dual encoding.}
An example of read is represented in the dual alphabet. The symbols
$\downarrow_{/i}^-$ represent a mismatch against the $-$ sequence, the
symbols $\downarrow_{/i}^+$ represent a mismatch against the $+$ sequence,
and the symbol $\Downarrow$ represent a mismatch against both. The index
$i$ is the match length of the sequence that is not mismatched. The
symbolic $+$ and $-$ sequences are represented below, where an open square
stands for a match and a closed square stands for a mismatch.}
\label{fig:dual}
\end{figure}

We assume that for each nucleotide, $a$ is the probability that the read
matches both sequences, $b$ is the probability that it matches only the
$+$ sequence, $c$ is the probability that it matches only the $-$ sequence
and $d$ is the probability that it matches none. Obviously $a+b+c+d=1$.
These definitions allow us to specify the weighted generating function of
the symbols and of the segments. To specify the transfer matrix, we only
have to make sure that we eliminate the combinations that create a match
of size $\gamma$ or more for any of the two sequences. For notational
convenience, we define
\begin{equation*}
\begin{gathered}
r_i^+(z) = cz \cdot (az)^i, \\
r_i^-(z) = bz \cdot (az)^i, \\
R_i(z) = dz \cdot \big( 1 + az + \ldots + (az)^i \big), \\
F_i(z) = 1 + az + \ldots + (az)^i,
\end{gathered}
\end{equation*}
and we can verify that the expression of the transfer matrix
$\tilde{M}_0(z)$ is
\begin{equation*}
\begin{blockarray}{ccccccccc}
   & \scriptstyle{\Downarrow} & \scriptstyle{\downarrow_{/1}^+} & 
    \ldots & \scriptstyle{\downarrow_{\gamma-1}^+} &
    \scriptstyle{\downarrow_{/1}^-} & \ldots &
    \scriptstyle{\downarrow_{/\gamma-1}^-} & \scriptstyle{|} \\
\begin{block}{c[cccccccc]}
\scriptstyle{\Downarrow} & R_{\gamma-1}(z)  & r_0^+(z) & \ldots &
    r_{\gamma-2}^+(z) & r_0^-(z) & \ldots & r_{\gamma-2}^-(z) &
    F_{\gamma-1}(z) \\
\scriptstyle{\downarrow_{/1}^+} & R_{\gamma-2}(z) & & & & & & &
    F_{\gamma-2}(z) \\
\scriptstyle{\downarrow_{/2}^+} & R_{\gamma-3}(z) & & & & & & &
    F_{\gamma-3}(z) \\
\vdots & \vdots & & \tilde{A}(z) & & & \tilde{B}_0(z) & & \vdots \\
\scriptstyle{\downarrow_{/\gamma-1}^+} & R_0(z) & & & & & & & F_0(z) \\
\scriptstyle{\downarrow_{/1}^-} & R_{\gamma-2}(z) & & & & & & &
    F_{\gamma-1}(z) \\
\scriptstyle{\downarrow_{/2}^+} & R_{\gamma-3}(z) & & & & & & &
    F_{\gamma-2}(z) \\
\vdots & \vdots & & \tilde{C}_0(z) & & & \tilde{D}(z) & & \vdots \\
\scriptstyle{\downarrow_{/\gamma-1}^+} & R_0(z) & & & & & & & F_0(z) \\
\scriptstyle{|} & 0 & 0 & \ldots & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}
where $A(z)$, $B(z)$, $C(z)$ and $D(z)$ are matrices of dimensions
$(\gamma-1) \times (\gamma-1)$ that are defined as
\begin{equation*}
\tilde{A}(z) = 
\begin{blockarray}{cccccc}
   & \scriptstyle{\downarrow_{/1}^+} & \scriptstyle{\downarrow_{/2}^+} &
    \ldots & \scriptstyle{\downarrow_{/\gamma-2}^+} &
    \scriptstyle{\downarrow_{/\gamma-1}^+} \\
\begin{block}{c[ccccc]}
\scriptstyle{\downarrow_{/1}^+} & 0 & r_0^+(z) & \ldots &
    r_{\gamma-4}^+(z) & r_{\gamma-3}^+(z) \\
\scriptstyle{\downarrow_{/2}^+} & 0 & 0 & \ldots &
    r_{\gamma-5}^+(z) & r_{\gamma-4}^+(z) \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\scriptstyle{\downarrow_{/\gamma-2}^+} & 0 & 0 & \ldots & 0 & r_0^+(z) \\
\scriptstyle{\downarrow_{/\gamma-1}^+} & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}
\begin{equation*}
\tilde{B}_0(z) = 
\begin{blockarray}{cccccc}
   & \scriptstyle{\downarrow_{/1}^-} & \scriptstyle{\downarrow_{/2}^-} &
   \ldots & \scriptstyle{\downarrow_{/\gamma-2}^-} &
   \scriptstyle{\downarrow_{/\gamma-1}^-} \\
\begin{block}{c[ccccc]}
\scriptstyle{\downarrow_{/1}^+} & r_0^-(z) & r_1^-(z) & \ldots &
    r_{\gamma-3}^-(z) & r_{\gamma-2}^-(z) \\
\scriptstyle{\downarrow_{/2}^+} & r_0^-(z) & r_1^-(z) & \ldots &
    r_{\gamma-3}^-(z) & 0 \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\scriptstyle{\downarrow_{/\gamma-2}^+} & r_0^-(z) & r_1^-(z) &
    \ldots & 0 & 0 \\
\scriptstyle{\downarrow_{/\gamma-1}^+} & r_0^-(z) & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}
\begin{equation*}
\tilde{C}_0(z) = 
\begin{blockarray}{cccccc}
   & \scriptstyle{\downarrow_{/1}^+} & \scriptstyle{\downarrow_{/2}^+} &
    \ldots & \scriptstyle{\downarrow_{/\gamma-2}^+} &
    \scriptstyle{\downarrow_{/\gamma-1}^+} \\
\begin{block}{c[ccccc]}
\scriptstyle{\downarrow_{/1}^-} & r_0^+(z) & r_1^+(z) & \ldots &
    r_{\gamma-3}^+(z) & r_{\gamma-2}^+(z) \\
\scriptstyle{\downarrow_{/2}^-} & r_0^+(z) & r_1^+(z) & \ldots &
    r_{\gamma-3}^+(z) & 0 \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\scriptstyle{\downarrow_{/\gamma-2}^-} & r_0^+(z) & r_1^+(z) & \ldots &
    0 & 0 \\
\scriptstyle{\downarrow_{/\gamma-1}^-} & r_0^+(z) & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}
\begin{equation*}
\tilde{D}(z) = 
\begin{blockarray}{cccccc}
   & \scriptstyle{\downarrow_{/1}^-} & \scriptstyle{\downarrow_{/2}^-} &
    \ldots & \scriptstyle{\downarrow_{/\gamma-2}^-} &
    \scriptstyle{\downarrow_{/\gamma-1}^-} \\
\begin{block}{c[ccccc]}
\scriptstyle{\downarrow_{/1}^-} & 0 & r_0^-(z) & \ldots &
    r_{\gamma-4}^-(z) & r_{\gamma-3}^-(z) \\
\scriptstyle{\downarrow_{/2}^-} & 0 & 0 & \ldots &
    r_{\gamma-5}^-(z) & r_{\gamma-4}^-(z) \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\scriptstyle{\downarrow_{/\gamma-2}^-} & 0 & 0 & \ldots & 0 & r_0^-(z) \\
\scriptstyle{\downarrow_{/\gamma-1}^-} & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray}.
\end{equation*}

The term of interest is the top right entry of
$\tilde{M}_0(z)\cdot(I-\tilde{M}_0(z))^{-1}$. To see why, observe that
every read can be prepended by segments terminated by $\Downarrow$ and
only by those (every other terminator would imply that one of the two
sequences has a nonzero match size at the start of the read). Thus reads
are precisely the sequences of segments that can follow the symbol
$\Downarrow$ and that are terminated by a tail, whose weighted generating
function is the top right entry of the matrix.

It is possible to express $\tilde{M}_0(z)\cdot(I-\tilde{M}_0(z))^{-1}$ in
closed form, but it involves polynomials of very large degree. It is
actually easier to proceed as in section~\ref{sec:example_skip} and to
compute the powers of $\tilde{M}_0(z)$ (using the arithmetic of truncated
polynomials) up to a given threshold value. Since each segment contains a
mismatch against at least one sequence, we define $\tilde{p}$ as the upper
bound on the probability of a mismatch $\tilde{p} = \max\{b,c,d\}$. The
updated formula (\ref{eq:bound}) now gives an upper bound of the
probability that a read of size $k$ contains $s$ or more mismatches as
\begin{equation*}
Pr(X \geq s) \leq \exp \left( (s-k)\log \frac{k-s}{k(1-\tilde{p})} -s\log
\frac{s}{k\tilde{p}} \right).
\end{equation*}

Now returning to the problem of computing $P(\overline{S}_0 \cap
\overline{S}_1)$, the $+$ sequence is interpreted as the target and the
$-$ sequence as the duplicate. Based on the assumptions of the error model
presented in section~\ref{sec:error}, this implies that $a =
(1-p)(1-\mu)$, $b = (1-p)\mu$, $c = p\mu/3$, and $d = p(1-\mu/3)$.

With these values, we can fully specify the matrix $\tilde{M}_0(z)$, and
compute its powers until the upper bound above is lower than a set
fraction $\varepsilon$ of the current value, which gives an accurate
estimate of $P(\overline{S}_0 \cap \overline{S}_1)$.

\subsection{Corollary: null seeding}

Observe that expression (\ref{eq:Pnull}) is the probability of null
seeding. The method described above thus allows us to compute this
probability at no additional cost. This probablity is less relevant than
the probabilities that the seeding process is on target or off target, but
at times, it may be useful to know the probability that a read is
not mappable, especially when reads are relatively short.

\subsection{Illustration}
\label{sec:illdual}

Using the method above, we can compute the probability that a seeding
scheme using exact seeds is off target. This depends on the exact
specifications of the problem, so for illustration purposes, we choose
reads of size $k=50$ sequenced with an instrument with error rate
$p=0.01$, and se use exact seeds of size $\gamma=17$. We then vary the
number of duplicates $N$ and their divergence rate $\mu$, which both
depend on the particular read being sequenced.

\section{Off-target skip seeds}

To compute the probability that the seeding process is off target when
using skip seeds, we observe that the logic of section~\ref{sec:offexact}
can be transposed with very few modifications. In particular, the
probability can be computed through expression (\ref{eq:Poff}), where
$S_0$ is the event that the read has a \emph{skip} seed for the target and
$S_1$ that it has a \emph{skip} seed for the first duplicate.

We have already seen how to compute $P(\overline{S}_0)$ in
section~\ref{sec:example_skip}, we now need to find a way to compute
$P(\overline{S}_0 \cap \overline{S}_1)$ when using skip seeds.

\subsection{The skip dual encoding}

Once again, we start by recoding the reads in a specialized alphabet. We
define the skip-$n$ dual alphabet $\tilde{\mathcal{A}}_n = \{\square, |,
\Downarrow_0, \Downarrow_1, \Downarrow_2, \ldots, \Downarrow_n,
\downarrow^-_{/1}, \downarrow^-_{/2}, \ldots, \downarrow^+_{/1},
\downarrow^+_{/2}, \ldots\}$. The symbols $\square$, $|$,
$\downarrow^-_{/i}$ and $\downarrow^+_{/i}$ have the same meaning as in
the dual alphabet of section~\ref{sec:dual}. The symbols $\Downarrow_i$
indicate that both sequences have match length 0 and that the next
non-skipped position is $i$ nucleotides downstream.

Fig.~\ref{fig:skip_dual} shows the same read as Fig.~\ref{fig:dual}, but
here it is represented in the skip-3 dual encoding. It is important to
note several differences with Fig.~\ref{fig:dual}. The first is that the
symbols $\Downarrow_i$ are not always associated with double mismatches.
For instance, the symbol $\Downarrow_0$ on the right side of the read
corresponds to a mismatch for the $+$ sequence only. This happens whenever
the $+$ and the $-$ sequences are mismatched in the same interval between
non-skipped positions (the mismatches do not need to be on the same
nucleotide).

Another difference is that when the symbols $\downarrow^-_{/i}$ and
$\downarrow^+_{/i}$ alternate, $i$ is no longer the distance to the
previous terminator. This is visible for the symbol $\downarrow^+_{/4}$ on
the left side of the read, where the previous terminator is 6 nucleotides
upstream. When the symbols alternate, $i$ is the distance to the previous
non-skipped position.

\begin{figure}[h]
\centering
\includegraphics[scale=0.85]{sketch_skip_dual.pdf}
\caption{\textbf{Example skip dual encoding.}
The read of Fig.~\ref{fig:dual} is represented in the dual alphabet. The
vertical bars above the read indicate non-skipped positions. The symbols
$\downarrow_{/i}^-$ and $\downarrow_{/i}^+$ have the same meaning as in
the dual alphabet. The symbols $\Downarrow_i$ indicate that both sequences
have match length 0 and that the next non-skipped position is located $i$
nucleotides downstream. Other graphical representations are as in
Fig.~\ref{fig:dual}.}
\label{fig:skip_dual}
\end{figure}

As in the case of the skip encoding, a fair amount of bookkeeping is
required to specify the weighted generating functions of interest.
We reuse the quantities $a+b+c+d=1$ and give them the same meaning as in
section~\ref{sec:dual}. For notational convenience, we define
\begin{eqnarray*}
N_i(z) &=& 1+z+\ldots+z^i, \\
W_j(z) &=& dz \cdot (az)^x \cdot \big( 1 + (az)^{n+1} + \ldots +
  (az)^{(n+1)m} \big), \\
  &\quad& \text{where } x = n-j \modulo{n+1},
  \text{ and } m = \left\lfloor
  \frac{\gamma-1-x}{n+1} \right\rfloor. \\
U_{i,j}(z) &=& dz \cdot (az)^x \cdot
  \big( 1 + (az)^{n+1} + \ldots + (az)^{(n+1)m} \big) +
\begin{cases}
  bz\cdot(az)^x \text{ if } j < y, \\
  0 \text{ otherwise}
\end{cases} \\
  &\quad& \text{where } x = -i-j-1 \modulo{n+1},
  y = n-i+1 \modulo{n+1}, \\
  &\quad& \text{and } m = \left\lfloor \frac{\gamma-1-i-x}{n+1}
    \right\rfloor, \\
V_{i,j}(z) &=& dz \cdot (az)^x \cdot
  \big( 1 + (az)^{n+1} + \ldots + (az)^{(n+1)m} \big) +
\begin{cases}
  cz\cdot(az)^x \text{ if } j \leq x, \\
  0 \text{ otherwise}
\end{cases} \\
  &\quad& \text{ where } x = -i-j-1 \modulo{n+1},
  m = \left\lfloor \frac{\gamma-1-i-x}{n+1} \right\rfloor.
\end{eqnarray*}

With these definitions, one can check that the expression of the transfer
matrix $\tilde{M}_n(z)$ is

\begin{equation*}
\rotatebox{90}{$
\begin{blockarray}{cccccccccccc}
   & \scriptstyle{\Downarrow_0} & \scriptstyle{\Downarrow_1} &
   \ldots & \scriptstyle{\Downarrow_n} &
    \scriptstyle{\downarrow_{/1}^+} & 
    \ldots & \scriptstyle{\downarrow_{\gamma-1}^+} &
    \scriptstyle{\downarrow_{/1}^-} & \ldots &
    \scriptstyle{\downarrow_{/\gamma-1}^-} & \scriptstyle{|} \\
\begin{block}{c[ccccccccccc]}
\scriptstyle{\Downarrow_0} & W_0(z) & W_1(z) & \ldots & W_n(z) &
    r_0^+(z) & \ldots & r_{\gamma-2}^+(z) & r_0^-(z) & \ldots &
    r_{\gamma-2}^-(z) & F_{\gamma-1}(z) \\
\scriptstyle{\Downarrow_1} & z  & 0 & \ldots &
  0 & 0 & \ldots & 0 & 0 & \ldots & 0 & N_0(z) \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots & \ddots &
  \vdots & \vdots & \ddots & \vdots & \vdots \\
\scriptstyle{\Downarrow_n} & z^n & 0 & \ldots &
  0 & 0 & \ldots & 0 & 0 & \ldots & 0 & N_{n-1}(z) \\
\scriptstyle{\downarrow_{/1}^+} & U_{1,0}(z) & U_{1,1}(z) & \ldots &
  U_{1,n}(z) & & & & & & & F_{\gamma-2}(z) \\
\vdots & \vdots & \vdots & \ddots & \vdots & & \tilde{A}(z) & & &
  \tilde{B}_n(z) & & \vdots \\
\scriptstyle{\downarrow_{/\gamma-1}^+} & U_{\gamma-1,0}(z) &
  U_{\gamma-1,1}(z) & \ldots & U_{\gamma-1,n}(z) & & & & & & & F_0(z) \\
\scriptstyle{\downarrow_{/1}^-} & V_{1,0}(z) & V_{1,1}(z) & \ldots &
    V_{1,n}(z) & & & & & & & F_{\gamma-2}(z) \\
\vdots & \vdots & \vdots & \ddots & \vdots & & \tilde{C}_n(z) &
  & & \tilde{D}(z) & & \vdots \\
\scriptstyle{\downarrow_{/\gamma-1}^-} & V_{\gamma-1,0}(z) &
    V_{\gamma-1,1}(z) & \ldots & V_{\gamma-1,n}(z) & & &
    & & & & F_0(z) \\
\scriptstyle{|} & 0 & 0 & \ldots & 0 & 0 & \ldots & 0 & 0 &
  \ldots & 0 & 0 \\
\end{block}
\end{blockarray}.
$}
\end{equation*}

The matrices $\tilde{A}(z)$ and $\tilde{C}(z)$ in the expression of
$\tilde{M}_n(z)$ are the same as in section~\ref{sec:dual}. The matrices
$\tilde{B}_n(z)$ and $\tilde{C}_n(z)$ are defined as
\begin{equation*}
\tilde{B}_n(z) = 
\begin{blockarray}{cccccc}
   & \scriptstyle{\downarrow_{/1}^-} & \scriptstyle{\downarrow_{/2}^-} &
   \ldots & \scriptstyle{\downarrow_{/\gamma-2}^-} &
   \scriptstyle{\downarrow_{/\gamma-1}^-} \\
\begin{block}{c[ccccc]}
\scriptstyle{\downarrow_{/1}^+} & s_{1,1}(z) & s_{1,2}(z) &
    \ldots & s_{1,\gamma-2}(z) & s_{1,\gamma-1}(z) \\
\scriptstyle{\downarrow_{/2}^+} & s_{2,1}(z) & s_{2,2}(z) &
    \ldots & s_{2,\gamma-2}(z) & 0 \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\scriptstyle{\downarrow_{/\gamma-2}^+} & s_{\gamma-2,1}(z) &
    s_{\gamma-2,2}(z) & \ldots & 0 & 0 \\
\scriptstyle{\downarrow_{/\gamma-1}^+} & s_{\gamma-1,1}(z) & 0 &
    \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}
\begin{equation*}
\tilde{C}_n(z) = 
\begin{blockarray}{cccccc}
   & \scriptstyle{\downarrow_{/1}^+} & \scriptstyle{\downarrow_{/2}^+} &
    \ldots & \scriptstyle{\downarrow_{/\gamma-2}^+} &
    \scriptstyle{\downarrow_{/\gamma-1}^+} \\
\begin{block}{c[ccccc]}
\scriptstyle{\downarrow_{/1}^-} & t_{1,1}(z) & t_{1,2}(z) &
    \ldots & t_{1,\gamma-2}(z) & t_{1,\gamma-1}(z) \\
\scriptstyle{\downarrow_{/2}^-} & t_{2,1}(z) & t_{2,2}(z) &
    \ldots & t_{2,\gamma-2}(z) & 0 \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\scriptstyle{\downarrow_{/\gamma-2}^-} & t_{\gamma-2,1}(z) &
    t_{\gamma-2,2}(z) & \ldots & 0 & 0 \\
\scriptstyle{\downarrow_{/\gamma-1}^-} & t_{\gamma-1,1}(z) &
    0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}
with
\begin{eqnarray*}
s_{i,j} &=&
  \begin{cases}
  cz \cdot (az)^{x+j-1} &\text{ if } i+j+x \leq \gamma \\
  0                     &\text{ otherwise,}
  \end{cases} \\
t_{i,j} &=&
  \begin{cases}
  bz \cdot (az)^{x+j-1} &\text{ if } i+j+x \leq \gamma \\
  0                     &\text{ otherwise,}
  \end{cases} \\
  \text{where } x &=& -i \modulo{n+1}, \text{ in both cases.} \\
\end{eqnarray*}

The computation is performed exactly as described in
section~\ref{sec:dual}. We compute the successive powers of
$\tilde{M}_n(z)$ in the arithmetic of truncated polynomials and stop the
iterations using the same criterion. So there is no change other than the
definition of the transfer matrix.

\begin{remark}
Observe that when $n=0$ the matrix $\tilde{M}_n(z)$, is identical to the
matrix $\tilde{M}_0(z)$ of section~\ref{sec:dual}, again consistent with
the fact that exact seeds are skip-$n$ seeds for which $n=0$. The same
applies to $\tilde{B}_n(z)$ and $\tilde{C}_n(z)$.
\end{remark}

% XXX The following algorithm works but it is not very useful    XXX
% XXX because it is substantially slower than the WGF method.    XXX
% XXX It has been used to check the accuracy of the calculations XXX
% XXX but there is no reason to put it in the manuscript.        XXX

%\begin{algorithm}[H]
%\label{alg:mcmcskip}
%\SetAlgoLined
%\KwResult {Sample a read at random. Return 1 if the read contains a
%good seed, otherwise return 0.}
%  $\lambda \leftarrow 0$ \Comment*[r]{Current read size.}
%  $s_+ \leftarrow 0$ \Comment*[r]{Current + match size.}
%  $s_- \leftarrow 0$ \Comment*[r]{Current - match size.}
%  \While {$\lambda < k$}{
%    $i \leftarrow geom(1-a)-1$
%                \Comment*[r]{Double-match nucleotides.}
%    \If {$i \geq k-\lambda$}{
%      $s_+ \leftarrow s_+ + k-\lambda$ \;
%      $s_- \leftarrow s_- + k-\lambda$ \;
%      break \;
%    }
%    $s_+ \leftarrow s_+ + i$ \;
%    $s_- \leftarrow s_- + i$ \;
%    \If {$s_+ \geq \gamma$ or $s_- \geq \gamma$}{
%      return 1 \;
%    }
%    $\lambda \leftarrow \lambda + i+1$ \;
%    $r \leftarrow rand(0,1-a)$
%                \Comment*[r]{Uniform random number.}
%    \uIf {$r < b$}{
%      $s_+ \leftarrow s_++1$ \;
%      $s_- \leftarrow -(-\lambda \mod n+1)$ \;
%    }\uElseIf{$r < b+c$}{
%      $s_+ \leftarrow -(-\lambda \mod n+1)$ \;
%      $s_- \leftarrow s_-+1$ \;
%    }\Else{
%      $s_+ \leftarrow -(-\lambda \mod n+1)$ \;
%      $s_- \leftarrow -(-\lambda \mod n+1)$ \;
%    }
%}
%\eIf {$s_+ \geq \gamma$ or $s_- \geq \gamma$}{
%  return 1 \;
%}{
%  return 0 \;
%}
%\end{algorithm}

\subsection{Illustration}
\label{sec:illskipdual}

Using the method above, we can compute the probability that a seeding
scheme using skip seeds is off target. Using the same settings as in
section~\ref{sec:illdual} ($k=50$, $p=0.01$ and $\gamma=17$), we set
decide to use skip-9 seeds. We then vary the number of duplicates $N$ and
their divergence rate $\mu$.

\section{Off-target MEM seeds}

MEM seeds are substantially more complex than exact seeds and skip seeds.
As we will see below, the duplicates can prevent the existence of an
on-target MEM seed, so we need to take them into account in the
combinatorial construction.

\subsection{Hard and soft masking}

We first introduce two important notions that will be the key of
understanding the behavior of MEM seeds.

\begin{definition}
At a given position of the read, a duplicate is a \emph{hard mask} if its
match length is strictly longer than the match length of the target. A
duplicate is a \emph{soft mask} if its has the same match length as the
target.
\end{definition}

Fig.~\ref{fig:hard_vs_soft_masks} gives a graphical intuition of hard and
soft masks. It is important to bear in mind that hard and soft masks
depend on the position of interest: a sequence can be a mask at the left
end of the read but not at the right end, or the opposite.

\begin{figure}[h]
\centering
\includegraphics[scale=0.85]{hard_vs_soft_masks.pdf}
\caption{\textbf{Example of hard and soft masks.}
Genomic sequencess are shown below a read where open squres represent
nucleotides. In the sequences, the open squares represent matches and the
closed squares represent mismatches. The nucleotides contributing to the
match length are represented as grey boxes. At the focus position, the
match length of the target is 7. The first duplicate is a hard mask
because its match length is $13>7$. The second duplicate is a soft mask
because its match length is 7, as the target. The third duplicate is not
a mask because its match length is $2<7$.}
\label{fig:hard_vs_soft_masks}
\end{figure}

Hard and soft masks explain the counter-intuitive of MEM seeds. For
instance, in Fig.~\ref{fig:full_masking_example} the target cannot be
discovered because every nucleotide of the read has a hard mask. In
Fig.~\ref{fig:short_vs_long}, the target could be discovered if the read
were shorter because a hard mask would turn into a soft one.

From the definition, we see that the last nucleotide of every strict
on-target MEM seed is always unmasked. Conversely, an unmasked nucleotide
always belongs to exactly one strict on-target MEM seed. Also, the last
nucleotide of every shared on-target MEM seed is always soft-masked, but a
soft-masked nucleotide does not always belong to a shared on-target MEM
seed.

Since hard and soft masks inform us about the positions of on-target
MEM seeds, we construct an alphabet that encodes the masking status of the
nucleotides.

\subsection{The MEM alphabet}

As before, we recode the reads as sequences of letters from a specialized
alphabet called the MEM alphabet $\mathring{\mathcal{A}} = \{\square, |,
\uparrow_{/1}, \uparrow_{/2}, \uparrow_{/3} \ldots, \downarrow_{/0},
\downarrow_{/1}, \downarrow_{/2}, \ldots\}$.

The symbols $\downarrow_{/m}$ indicate that the nucleotide is a sequencing
error and $m \geq 0$ is the number of duplicates that \emph{match} the
nucleotide. Since a sequencing error is always a mismatch against the
target, the symbol $\downarrow_{/0}$ indicates that the ncleotide is a
mismatch against \emph{every} sequence. The symbols $\uparrow_{/i}$
indicate a change in masking status: the nucleotide is not masked but its
left neighbor is. The index $i \geq 1$ is the number of nucleotides since
the last mismatch or since the beginning of the read. All the other
nucleotides are represented by the $\square$ symbol and the $|$ symbol is
appended to the end of the read as usual.

Note that in the symbols $\downarrow_{/m}$ and $\downarrow_{/i}$, the
numbers $m$ and $i$ have different meanings. In the symbol
$\downarrow_{/m}$, the index $m$ is a number of sequences ($0 \leq m \leq
N$ where $N$ is the number of duplicates); in the symbol $\uparrow_{/i}$,
the index $i$ is a number of nucleotides. Fig.~\ref{fig:sketch_extended}
shows the encoding of the read from Fig.~\ref{fig:exact_seed_example} in
the MEM alphabet.

\begin{figure}[h]
\centering
\includegraphics[scale=.84]{sketch_extended.pdf}
\caption{\textbf{The MEM encoding}.
The read of Fig.~\ref{fig:exact_seed_example} is represented in the MEM
alphabet. The arrows departing from the numbers help understand their
meaning. The symbol $\downarrow_{/m}$ is indexed by the number $m$ of
locations that match the nucleotide. The symbol $\uparrow_{/i}$ is
indexed by the number $i$ of nucleotides from the last error or from the
beginning of the read.}
\label{fig:sketch_extended}
\end{figure}

The MEM alphabet captures the masking status of the nucleotide: the symbol
$\downarrow_{/m}$ indicates that the nucleotide has $m$ hard masks and
$N-m$ soft masks. The symbols $\uparrow_{/i}$ indicate that the nucleotide
is unmasked and that the previous $i-1$ nucleotides were masked.

In the MEM alphabet, strict on-target MEM seeds are the longest stretches
of symbols containing some symbol $\uparrow_{/i}$ and not containing any
symbol $\downarrow_{/m}$. Indeed, such a stretch is a match for the target
because it does not contain any symbol $\downarrow_{/m}$, it is match for
the target only because it contains at least one unmasked nucleotide
(marked by $\uparrow_{/i}$), and it cannot be extended because it is
flanked by sequencing errors or by the ends of the reads. Note that there
is exactly one symbol $\uparrow_{/i}$ per strict on-target MEM seed, and
therefore two symbols $\uparrow_{/i}$ must be separated by at least one
symbol $\downarrow_{/m}$.

Shared on-target MEM seeds are the longest stretches of symbols $\square$
flanked by $\downarrow_{/0}$, or by the ends of the read. Indeed, such a
stretch is a MEM seed because it matches the target and it cannot be
extended ($\downarrow_{/0}$ is a mismatch against every sequence). Also,
it cannot be a strict on-target MEM seed because it does not contain any
$\uparrow_{/i}$ symbol, so it must be a shared on-target MEM seed.

As before, the read is converted from a sequence of symbols to a sequence
of segments that consist of 0 or more symbols $\square$ followed by a
terminator. We then specify the weighted generating functions of those
segments and fill the tansfer matrix $\mathring{M}_N(z)$ of the reads that
do not contain an on-target MEM seed. We introduce the terms of the matrix
by increasing oder of complexity.

\subsection{Segments following $\uparrow_{/i}$}

A segment terminated by $\uparrow_{/i}$ is the beginning of an on-target
MEM seed of size at least $i$. The seed reaches the next sequencing
error or the end of the read, so the number of $\square$ symbols in the
next segment must be at most $\gamma-i-1$ and it must be terminated by a
$\downarrow$ symbol or by the tail terminator $|$.

The following definition will simplify the notations.

\begin{definition}
The probability that a symbol is $\downarrow_{/m}$ given that the
nucleotide is a read error is
\begin{equation}
\label{eq:omega}
\omega_m = {N \choose m} \big(1 - \mu/3\big)^{N-m} \big(\mu/3\big)^m.
\end{equation}
\end{definition}

Indeed, if the nucleotide is an error, the symbol is $\downarrow_{/m}$ for
some $m$ between 0 and $N$. Each duplicate is a match with probability
$\mu/3$, so $m$ has a Binomial distribution with parameters $(N, \mu/3)$.

On this segment, the matches between the read and the duplicates are
irrelevant, so the weighted generating function of a symbol $\square$ is
simply $qz$ (recall that $p = 1-q$ is the probability of a sequencing
error). The weighted generating function of the $\downarrow_{/m}$
terminator is $\omega_m pz$, so the weighted generating function of the
segments terminated by $\downarrow_{/m}$ following a segment terminated by
$\uparrow_{/i}$ is
\begin{equation}
\label{eq:D}
D_{i,m}(z) = \omega_m pz \sum_{j=0}^{\gamma-i-1} (qz)^j.
\end{equation}

And the weighted generating function of the tail segments following
segments terminated by $\uparrow_{/i}$ is
\begin{equation}
\label{eq:E}
E_i(z) = \sum_{j=0}^{\gamma-i-1} (qz)^j.
\end{equation}

\subsection{Segments following $\downarrow_{/m}$}

The symbol $\downarrow_{/m}$ signifies that the nucleotide has $m$ hard
masks and $N-m$ soft masks. If all the masks disappear before the first
read error, the next terminator will be one of the $\uparrow_{/j}$
symbols, otherwise it will be one of the $\downarrow_{/m}$ symbols. We
separate the cases based on the terminator of the segment.

\subsubsection*{Case 1: the terminator is $\uparrow_{/j}$}

\begin{definition}
The probability that a given duplicate sequence contains a mismatch in a
sequence of $j$ error-free nucleotides is
\begin{equation}
\label{eq:xi}
\xi_j = 1-(1-\mu)^j.
\end{equation}
This is the probability that a hard or soft mask vanishes within $j$
correct nucleotides.
\end{definition}

Indeed, every nucleotide of the duplicate differs from the target
with probability $\mu$. In the absence of sequencing error, this is also
the probability that a nucleotide of the duplicate differs from the read.
The probability that $j$ nucleotides in a row are identical to the read is
thus $(1-\mu)^j$ and the probability that at least one of them is
different is the complement $1-(1-\mu)^j$.

With this notation, the probability that at least one mask survives a
sequence of $j$ error-free nucleotides is thus $1-\xi_j^N$, and the
probability that there remains a mask at the $j-1$-th but not at the 
$j$-th error-free nucleotide is $\xi_j^N - \xi_{j-1}^N$. From this we
conclude that the weighted generating function of the segments terminated
by $\uparrow_{/j}$ following a segment terminated by $\downarrow_{/m}$ is
\begin{equation}
\label{eq:B}
B_j(z) = \Big( \xi_j^N-\xi_{j-1}^N \Big) (qz)^j.
\end{equation}

The fact that the reads have no on-target seeds imposes $j < \gamma$. Also
note that this expression is the same for all $\downarrow_{/m}$ symbols
(it does not depend on $m$).

\subsubsection*{Case 2a: the terminator $|$ comes before the $\gamma$-th
nucleotide}

In this case there can be no on-target seed because the read finishes
too early. We must just enforce the condition that at least one of the $N$
masks survives until the end of the segment in order to exclude the
$\uparrow_{/j}$ symbols. The weighted generating function is
\begin{equation*}
\sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i.
\end{equation*}

\subsubsection*{Case 2b: the terminator $|$ comes after the $\gamma$-th
nucleotide}

In this case, the soft masks do not hide the target: even if one of them
survives until the end of the read, there will be an on-target seed
(shared in this case). To avoid on-target seeds, we must enforce the
condition that at least one hard mask survives until the end of the
segment (which is impossible if $m = 0$). The weighted generating function
is
\begin{equation*}
\sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \Big) (qz)^i.
\end{equation*}

Summing the expressions from cases 2a and 2b, we find that the
weighted generating function of the tail following segments terminated by
$\downarrow_{/m}$ is
\begin{equation}
\label{eq:C}
C_m(z) =
\sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i +
  \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \Big) (qz)^i.
\end{equation}

\subsubsection*{Case 3a: the terminator $\downarrow_{/n}$ comes before the
$\gamma$-th nucleotide}

In this case, there can be no on-target seed and we must only exclude the
$\uparrow_{/j}$ terminators. As we have seen above, this implies that at
least one of the $N$ masks survives until the terminator. For a read of
size $j+1$, this occurs with probability $1-\xi_j^N$. Including the
terminator and summing over $j+1 \leq \gamma$, we see that the weighted
generating function is thus
\begin{equation}
\omega_n pz \sum_{j=0}^{\gamma-1} \Big(1 - \xi_j^N \Big) (qz)^j.
\end{equation}

\subsubsection*{Case 3b: the terminator $\downarrow_{/n}$ comes after the
$\gamma$-th nucleotide}

This case is by far the most convoluted. Since the segment contains at
least $\gamma$ error-free nucleotides, we must enforce the condition that
it does not contain an on-target seed. This will be the case if any of the
two following conditions is validated: $i)$ at least one hard mask covers
all the error-free nucleotides, or $ii)$ all the hard masks vanish but at
least one soft mask covers the whole segment (including the terminator).

The two conditions are mutually exclusive by construction. They are
graphically represented in the diagram below. The left panel corresponds
to case $i)$ and the panel to case $ii)$. The top row represents the
target, and the bottom rows represent duplicates (using the same
symbols as in Fig.~\ref{fig:sketch_extended}).
\begin{inset}
\includegraphics{masks.pdf}
\end{inset}

Whenever a hard mask (here the first duplicate) covers the nucleotides as
shown in the left panel, there can be no on-target seed. The positions
marked with a question mark are irrelevant, they cannot change the fact
that there is no on-target MEM seed. If the hard masks vanish, as in the
right panel, then we need to look at the soft masks. If a soft mask covers
the whole segment, then there can be no on-target seed. In all other cases
there is an on-target MEM seed.

For a segment of size $j+1$, condition $i)$ has probability $\big(1 -
\xi_j^m \big)$. Summing over $j+1 > \gamma$ and including the terminator,
we see that the associated weighted generating function is thus
\begin{equation*}
\omega_n pz \sum_{j=\gamma}^\infty \Big(1 - \xi_j^m \Big) (qz)^j.
\end{equation*}

Condition $ii)$ is more convoluted, so we introduce some further
notations to solve this sub-case.
\begin{definition}
The probability that a given duplicate sequence contains a mismatch in a
sequence of $j$ error-free nucleotides followed by an error is
\begin{equation}
\label{eq:eta}
\eta_j = 1-(1-\mu)^j\mu/3.
\end{equation}
This is the probability that a hard or soft mask vanishes within $j$
correct nucleotides followed by a sequencing error.
\end{definition}

Indeed, the probability that there are $j$ matches between the duplicate
and the target is $(1-\mu)^j$. If there are no sequencing errors, this is
also the probability that there are $j$ matches between the duplicate and
the read. The probability that the duplicate matches the subsequent
sequencing error is $\mu/3$, so the probability of that there are $j+1$
matches including the sequencing error is $(1-\mu)^j\mu/3$. Finally, the
probability that there is a mismatch is the complement $1-(1-\mu)^j\mu/3$.

Let us for now consider a segment of fixed size $j+1$. From expressions
(\ref{eq:xi}) and (\ref{eq:eta}), the probability of condition $ii)$ is
\begin{equation*}
\xi_j^m \Big(1 - \eta_j^{N-m} \Big),
\end{equation*}
but we need to break up this term among all the possible terminators
$\downarrow_{/n}$ ($0 \leq n \leq N$) in order to fill the different
entries of the transfer matrix. For this, we split this term on the number
of soft masks that run until and including the terminator. From expression
(\ref{eq:eta}), the probability that there are $r \geq 1$ such soft masks
is
\begin{equation}
\label{eq:softmasks_r}
{N-m \choose r} (1 - \eta_j)^r \eta_j^{N-m-r}.
\end{equation}

For now we consider $r$ fixed; we will compute the marginal probability at
the final stage. By construction, each of those $r$ soft masks matches the
terminator, so the total number of matches is $r$ plus the number of
sequences that also match the last nucleotide, among the remaining $N-m-r$
soft masks and the $m$ hard masks.

Let us start with the $m$ hard masks. The probability that each of them
matches the terminator is simply $\mu/3$, and the probably that they do
not is $1-\mu/3$.

The case of the $N-m-r$ soft masks is more complicated because they can
vanish precisely on the terminator --- recall that in case $ii$) all the
hard masks are assumed to vanish before. If the soft mask failed within
the first $j$ nucleotides, then the $j+1$-th nucleotide can be anything
and it will match the terminator with probability $\mu/3$. But if the soft
mask survived the first $j$ nucleotides, then it \emph{must} fail on the
$j+1$-th and it cannot match the terminator. From expressions
(\ref{eq:xi}) and (\ref{eq:eta}), the probability that a given soft mask
fails withtin the first $j$ nucleotides is $\xi_j/\eta_j$ --- this is the
conditional probability that it fails within the first $j$ nucleotides
given that it fails within the segment. Finally the probability that such
a soft mask matches the terminator is $\mu/3 \cdot \xi_j / \eta_j$.

Summing the contributions of the hard masks ($m$ in total, each matching
the terminator with probability $\mu/3$) and of the soft masks ($N-m-r$ in
total, each matching the terminator with probability
$\mu/3\cdot\xi_j/\eta_j$), the probability that the total number of
matches is $n-r$ appears as the convolution product
\begin{equation*}
\frac{(\mu/3)^{n-r}(1-\mu/3)^{N-n}}{\eta_j^{N-m-r}}
\psi_{j,m,n,r}\text{, where \;}
\psi_{j,m,n,r} = \sum_{q \geq 0}{m \choose q}{N-m-r \choose n-r-q}
\xi_j^{n-r-q}.
\end{equation*}

Finally, we need to compute the marginal probability over the number $r$
of soft masks that survive until the end of the read. Multiplying by the
probability of $r$ from expression (\ref{eq:softmasks_r}) and summing over
$r \geq 1$, the probability that $n$ duplicate sequences match the
terminator appears as
\begin{eqnarray*}
&\;& \sum_{r\geq1} {N-m \choose r}
(1 - \eta_j)^r (\mu/3)^{n-r} (1-\mu/3)^{N-n} \psi_{j,m,n,r} \\
&=& (\mu/3)^n(1-\mu/3)^{N-n} \sum_{r\geq1} {N-m \choose r}
  (1 - \mu)^{rj} \psi_{j,m,n,r} \\
&=& \omega_n \cdot \zeta_{j,m,n},
\end{eqnarray*}
where
\begin{equation}
\label{eq:zeta}
\zeta_{j,m,n} = \sum_{r\geq1} {N-m \choose r}
(1-\mu)^{rj} \psi_{j,m,n,r} \bigg/ {N \choose n}.
\end{equation}

This is the probability that the terminator is the symbol
$\downarrow_{/n}$ given that the segment has size $j+1 > \gamma$, that
the first sequencing error occurs on the last nucleotide, that the
preceding terminator was $\downarrow_{/m}$ and that the $m$ hard masks
fail before the end of the segment.

Summing the terms from case 3a and from case 3b, we find that the weighted
generating function of the segments terminated by $\downarrow_{/n}$
following segments terminated by $\downarrow_{/m}$ is
\begin{equation}
\label{eq:A}
A_{m,n} =
\omega_n pz \sum_{j=0}^{\gamma-1} \Big(1 - \xi_j^N \Big) (qz)^j + \omega_n
pz \sum_{j=\gamma}^\infty \Big(1 - \xi_j^m \cdot
(1- \zeta_{j,m,n}) \Big) (qz)^j.
\end{equation}

\subsection{Expression of $\mathring{M}_N(z)$}
\label{sec:expression_of_M}

Collecting and arranging the results above, we can verify that the final
expression of the transfer matrix $\mathring{M}_N(z)$ is
\begin{equation*}
\begin{blockarray}{cccccccc}
   & \dn{0} & \ldots & \dn{N} & \up{1} & \ldots & \up{\gamma-1} & \nd \\
\begin{block}{c[ccccccc]}
\dn{0} & A_{0,0}(z) & \ldots & A_{0,N}(z) & B_1(z) & \ldots &
    B_{\gamma-1}(z) & C_0(z) \\
\vdots & \vdots & \ddots & \vdots & \vdots & \ddots &
    \vdots & \vdots \\
\dn{N} & A_{N,0}(z) & \ldots & A_{N,N}(z) & B_1(z) & \ldots &
    B_{\gamma-1}(z) & C_N(z) \\
\up{1} & D_{1,0}(z) & \ldots & D_{1,N}(z) & 0 & \ldots & 0 & E_1(z) \\
\vdots & \vdots & \ddots & \vdots & \vdots & \ddots &
    \vdots & \vdots \\
\up{\gamma-1} & D_{\gamma-1,0}(z) & \ldots & D_{\gamma-1,N}(z) & 0 &
  \ldots & 0 & E_{\gamma-1}(z) \\
\nd & 0 & \ldots & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray}
\end{equation*}
where
\begin{gather}
\tag{\ref{eq:A}}
A_{m,n} =
\omega_n pz \sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i + \omega_n
pz \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \cdot
(1- \zeta_{i,m,n}) \Big) (qz)^i \\
\tag{\ref{eq:B}}
B_i(z) = \Big( \xi_i^N-\xi_{i-1}^N \Big) (qz)^i \\
\tag{\ref{eq:C}}
C_m(z) =
\sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i +
  \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \Big) (qz)^i \\
\tag{\ref{eq:D}}
D_{j,m}(z) = \omega_m pz \sum_{i=0}^{\gamma-j-1} (qz)^i \\
\tag{\ref{eq:E}}
E_j(z) = \sum_{i=0}^{\gamma-j-1} (qz)^i
\end{gather}
and where
\begin{gather}
\tag{\ref{eq:omega}}
\omega_m = {N \choose m} \big(1 - \mu/3\big)^{N-m} \big(\mu/3\big)^m \\
\tag{\ref{eq:xi}}
\xi_j = 1-(1-\mu)^j \\
\tag{\ref{eq:eta}}
\eta_j = 1-(1-\mu)^j\mu/3 \\
\tag{\ref{eq:zeta}}
\zeta_{j,m,n} = \sum_{r\geq1} {N-m \choose r}
(1-\mu)^{rj} \psi_{j,m,n,r} \bigg/ {N \choose n} \\
\notag
\psi_{j,m,n,r} = \sum_{q \geq 0}{m \choose q}{N-m-r \choose n-r-q}
\xi_j^{n-r-q}.
\end{gather}

\begin{remark}
In the special case $N=0$, the transfer matrix simplifies to the extent
that we can compute the weighted generating function of the reads without
on-target MEM seed in closed form. The result is
\begin{equation}
\tag{\ref{eq:F}}
\frac{1+qz+\ldots+(qz)^{\gamma-1}}
  {1-pz \big(1+qz+\ldots+(qz)^{\gamma-1} \big)}.
\end{equation}

Expression (\ref{eq:F}) was shown in section~\ref{sec:example_exact} to be
the weighted generating function of reads without on-target exact seed.
This shows that when there are no duplicates, MEM seeds have exactly the
same properties as exact seeds.
\end{remark}

\subsection{Computing MEM seeding probabilities}

The matrix $\mathring{M}_N(z)\cdot(I-\mathring{M}_N(z))^{-1}$ contains the
weighted generating functions of all the reads without on-target MEM
seeds. The term of interest, as usual, is the top right entry. Indeed,
every read can be prepended by segments terminated by $\downarrow_{/0}$
and only by those, otherwise the read would start with fewer than $N$ soft
masks. Thus, reads without an on-target MEM seed are precisely the
sequences of segments that can be appended to the symbol $\downarrow_{/0}$
and that are terimated by a tail.

To compute this term, we proceed as in section~\ref{sec:example_skip},
\textit{i.e.}, we compute the powers of $\mathring{M}_N(z)$ in the
arithmetic of truncated polynomials and we stop the iterations when the
terms are negligible. We bound the probability that the read contains more
than $e$ sequencing errors using the expression
\begin{equation*}
\tag{\ref{eq:bound}}
Pr(X \geq e) \leq \exp \left( (e-k)\log \frac{k-e}{k(1-p)} -e\log
\frac{e}{kp} \right),
\end{equation*}
but here not every segment contains an error. However, there cannot be two
symbols $\uparrow_{/j}$ in a row, so a read with $s$ segments so a tail
must contain a minimum number of sequencing errors which is
$e=\lfloor{s/2\rfloor}$. So we compute the powers of $\mathring{M}_N(z)$
until the upper bound is less than a set fraction $\varepsilon$ of the
current value.

If we call $M_0$ the event that the read contains an on-target MEM seed,
the method above gives us $P(\overline{M}_0)$. Calling $M_j$ the event
that the read contains a MEM seed for the $j$-th duplicate, we are
interested in the probability
\begin{equation*}
\tag{\ref{eq:pcomp}}
P\big(\overline{M}_0 \cap (M_1 \cup \ldots \cup M_N)\big) =
P(\overline{M}_0) - P\big(\overline{M}_0 \cap \overline{M}_1 \cap \ldots
\cap \overline{M}_N\big).
\end{equation*}

The key insight to compute $P\big(\overline{M}_0 \cap \overline{M}_1 \cap
\ldots \cap \overline{M}_N\big)$ is to realize that there is some MEM
seed, on-target or not, if and only if the read contains a match of size
$\gamma$ or more for any of the $N+1$ sequences. Therefore, this
probability is the same as the term $P\big(\overline{S}_0 \cap
\overline{S}_1 \cap \ldots \cap \overline{S}_N\big)$ computed in
section~\ref{sec:offexact}.

In conclusion, the probability that the MEM seeding process is off target
is
\begin{equation*}
P(\overline{M}_0) - P(\overline{S}_0) \cdot \left( \frac{P(\overline{S}_0
\cap \overline{S}_1)}{P(\overline{S}_0)} \right)^N,
\end{equation*}
where $P(\overline{M}_0)$ is computed using $\mathring{M}_N(z)$ as
explained in this section, $P(\overline{S}_0)$ is computed using a
recursive equation as explained in section~\ref{sec:example_exact}, and
$P(\overline{S}_0\cap\overline{S}_1)$ is computed using $\tilde{M}_0(z)$
as explained in section~\ref{sec:dual}.

\subsection{Monte Carlo sampling}
\label{sec:montecarlo}

One potential difficulty in computing $P(\overline{M}_0)$ is that the
matrix $\mathring{M}_N(z)$ has dimension $(N+\gamma+1) \times
(N+\gamma+1)$. So for MEM seeds the problem can become computationally
intractable because $N$ can be very large. For instance, the sequences
called \emph{Alu} have more than one million duplicates in the human
genome. There is no hope to compute the powers of $\mathring{M}_N(z)$ in
these conditions and we need an alternative method.

The symbolic representation as MEM segments can be used for an efficient
method to sample reads. Instead of generating the nucleotides of the $N+1$
sequences one by one, one can generate a single sequence of segments.
Since the number of segment does not depend on $N$, we can obtain a fast
Monte Carlo method to sample millions of reads and count the proportion
that contain an on-target MEM seed.

The principle is to proceed in cycles of two steps. We first sample the
position of the next sequencing error, which gives the position of the
next symbol $\downarrow_{/m}$, where $m$ will be determined at a later
stage. The second step is to determine whether there is a symbol
$\uparrow_{/j}$ before that. For this we sample the number of masks that
vanish before the symbol $\downarrow_{/m}$. If they all vanish, the read
contains an on-target MEM seed, provided the next read error is at a
distance greater than $\gamma$. Otherwise, we sample the number $m$ of
hard masks at the sequencing error, and the process is repeated until we
generate an on-target MEM seed, or until the read has size $k$ or greater
(in which case it has no on-target MEM seed).

The method is summarized in algorithm~\ref{alg:mcmc} below. It requires
efficient methods to sample from the geometric and from the binomial
distributions. Sampling from a geometric distribution can be done by
computing the logarithm of a uniform $(0,1)$ random variable. Sampling
from a binomial distribution can be done by the method of
Kachitvichyanukul and Schmeiser~\cite{kachitvichyanukul1988binomial}.

The running speed depends on the error rate of the instrument $p$, but not
on the number of duplicates $N$. For values of $p$ around 0.01,
the running time for 10,000,000 iterations is $1$-$10$ seconds on current
hardware.

\begin{algorithm}[H]
\label{alg:mcmc}
\SetAlgoLined
\KwResult {Sample a read at random. Return 1 if the read contains a
good MEM seed, otherwise return 0.}
  $\lambda \leftarrow 0$ \Comment*[r]{Current read size.}
  $m \leftarrow 0$ \Comment*[r]{Current number of hard masks.}
  \While {$\lambda < k$}{
    $i \leftarrow geom(p)-1$ \Comment*[r]{Error-free nucleotides.}
    \eIf {$i \geq k-\lambda$}{
      \eIf{$k-\lambda < \gamma$} {
        return 0\; }{
        $h \leftarrow binom(m,(1-\mu)^{k-\lambda})$
            \Comment*[r]{Surviving hard masks.}
        return 1 if $h = 0$, otherwise return 0\; }
    }
    {
      $h \leftarrow binom(m,(1-\mu)^i)$
          \Comment*[r]{Surviving hard masks.}
      $s \leftarrow binom(N-m,(1-\mu)^i \mu/3)$
          \Comment*[r]{Surviving soft masks.}
      \eIf {$ i \geq \gamma$ and $h = 0$ and $s = 0$}{
        return 1\;}{
        $m \leftarrow s + binom(N-s, \mu/3)$\;
        $\lambda \leftarrow \lambda + i+1$\;}
    }
 }
\end{algorithm}

Algorithm~\ref{alg:mcmc} is an important result. It gives a compact
solution to the problem of estimating the probability that a read can be
mapped when using MEM seeds. The algorithm is also much faster than the
naive approach of sampling every nucleotide of every sequence, because it
is equivalent to sampling the nucleotide sequence of \emph{all} the
duplicates.

\subsection{Illustration}

Here goes.

\section{Random seeds}
\label{sec:short_MEMs}

We have assumed throughout that seeds can match only the target or one of
its duplicates. In practice, seeds that overlap a sequencing error often
match random locations of the genome instead of the duplicates.

It is important for the validity of the theory that all the sequences that
have a seed are considered candidate locations (as long as the seed is
above the threshold $\gamma$ of course). If not, the estimates of the
mismapping rate are off. The hope is that spurious matches are shorter
than $\gamma$ so that they are invisible, but this is not always the case.
And if they qualify as a seed we need to verify the candidate location.

Such random hits can be problematic for two reasons: First, the time spent
verifying them is wasted. Second, they can cause false positives.
Fortunately, both issues can be addressed.

To save time during the alignment phase, we can prioritize the seeds so
that the best hit is likely to be discovered first, allowing us to bail
out from other alignments as early as possible. We can even bound
the maximum quality of some candidates so that the alignment can be
skipped altogether. For instance, if we already have a candidate at
distance 1 from the read, there is no need to verify candidates from MEM
seeds than span from neither end of the read. Indeed, the hit must be at
distance at least 2 because there is a mismatch before the MEM seed and
another one after it. These considerations depend on the implementation of
the mapper so we will not delve deeper here. What matters is that random
hits do not have to impose a significant burden on the time needed to
verify the candidates.

The second issue is that random hits can generate false positives. Such
cases occur when seeding is null (meaning that there is no seed for the
target nor for its duplicates). Since the hits are not homologous to the
read, the alignment score is essentially random, which makes it is easy to
detect.

Indeed, if the candidate location is random, mismatches occur with
probability 3/4. If it is the true location, mismatches occur with
probability $p$. We discard the seed because it is an automatic match of
size $\gamma$ (and possibly larger in the case of MEM seeds). Say that
there remains $L$ nucleotides and that $m$ of them are mismatches for the
candidate location. From Bayes formula, the probability that the location
is random given the number of mismatches is
\begin{equation}
\label{eq:bayes}
\frac{1}{1 + 4^L(p/3)^mq^{L-m}(1-\beta)/\beta},
\end{equation}
where $q=1-p$ and $\beta$ is the prior probability that the hit is
random.

The value of $\beta$ has little importance if $p$ is small. Say that
$p=0.01$ and that a read generates a hit in the genome such that 33
nucleotides need to be aligned after seeding. If the hit is random there
is a 99.99\% chance that at least 15 are mismatches. The denominator of
expression (\ref{eq:bayes}) is approximately $1 +
4.3\cdot10^{-18}(1-\beta)/\beta$. So unless $\beta < 10^{-17}$, the
support for the hypothesis that the hit is random is overwhelming.
Conversely, if the hit is not random, there is a 99.99\% chance that 4 or
fewer nucleotides are mismathes. In this case, the denominator is
approximately $1+6.8\cdot10^9(1-\beta)/\beta$, so unless $1-\beta <
10^{-9}$, the support for the hypothesis that the hit is not random is
overwhelming. If $p$ is small, we do not need to worry about the value of
$\beta$, one can choose for instance $\beta=1/2$ so that the term
$(1-/\beta)/\beta$ disappears in expression (\ref{eq:bayes}).

Since the probability that the best candidate is a random sequence is
either very small or very large, it either has no influence or dominates
the probability of a false positive. For every read, we can use as
confidence score the maximum of this probability of expression
(\ref{eq:bayes}) and of the estimated probability that seeding is off
target.

In summary, seeds in random sequences happen relatively frequently, but it
is possible to minimize their computational burden. Also, they are no
cause for concern regarding false positives because they can easily be
detected using Bayes formula, as shown in expression (\ref{eq:bayes}).

\section{Discussion}

We have devised a set of methods to compute the probability that seeding
heuristics fail and sets the mapping process for a false positive. This
fills a knowledge gap to understand and calibrate the performance of the
seeding heuristics. The pillars of our strategy are borrowed from analytic
combinatorics~\cite{flajolet2009analytic, sedgewick2013introduction}, even
though we do not follow the complete programme. Constructing generating
functions usually serves the purpose of finding their singularities in
order to approximate the solution. In our case, however, the weighted
generating functions cannot be computed so this strategy is not
applicable.

To find the probabilities of interest in the absence of a fully specified
weighted generating function, we compute only the first terms using
iterative methods as in explained in section~\ref{sec:example_skip}, 
or using Monte Carlo sampling as explained in
section~\ref{sec:montecarlo}. In this regard, the breakthrough is the
encoding of reads as MEM segments, which is a form of implicit Markov
imbedding~\cite{fu1994distribution}.

The method is general, but it is important to clearly state the
assumptions it depends on. First and most importantly, we have ignored
insertions and deletions. We assume that the sequencing errors are
substitutions only, which makes the method adapted to the Illumina
technology, but not to deletion-prone instruments such as the Oxford
Nanopore technology. We also assume that insertions and deletions never
occur among duplicated sequences. This is obviously incorrect, but our
initial tests with real data suggest that this is a minor impediment. On
real genomes, the estimations are only approximately correct and this is
but one of several simplifying assumptions. Incorporating insertions and
deletions would make the theory intractable, so it is presently unclear
how to deal with this type of error.

The second assumption is that all the candidate locations with a seed are
tested with an exact sequence alignment algorithm, so that the best
location is always discovered if it is in the candidate set. This is not
completely unrealistic, but it is important to realize that for plant and
animal genomes, a single seed may have tens of thousands of hits.
Therefore, most mappers impose a limit on the number of alignments so the
best hit may not be found even if it is in the candidate set. This is
again a minor impediment, since we can correct the probability that the
location is a false positive when only some candidates have been tested.

The third assumption is that all the duplicate sequences evolve
independently of each other and at the same rate. This is again incorrect,
because duplication events can happen continuously, creating complex
ancestry relationships. It is possible to infer the ancestry using tree
reconstruction techniques, but it would be challenging to incorporate this
information in the present theory. The symbols of the alphabets developed
above implicitly assume that the sequences are exchangeable and the
complexity of the calculations explodes if it is not the case.

The last assumption is that seeds can match only the target or one of its
duplicates. This does not hold in general but we have shown how to deal
with this possibility \textit{a posteriori} in
section~\ref{sec:short_MEMs}.

The Monte Carlo approach of algorithm~\ref{alg:mcmc} is relatively
straightforward, so one may wonder why the approach with weighted
generating functions is necessary for MEM seeds. The only reason is
precision. To estimate the frequency of an event by Monte Carlo sampling,
it must occur at least a few times in the simulation. For instance, with
10 million rounds of sampling, frequencies around $1/1,000,000$ or lower
cannot be measured accurately. When not interested in rare events, it is
thus a perfectly reasonable strategy.

For $N < 20$, the probability that MEM seeding is null or off-target is
relatively small, so we need a method that is accurate in this range.
Fortunately, the transfer matrix method is fast because the dimensions of
the matrix $\mathring{M}_N(z)$ are not prohibitive.

The proposed methods meet the demand for speed. One needs to compute the
probabilities only once for a given value of $N$ and $\mu$ (the error rate
$p$ is known and constant). For $N > 20$, the iterative method is usually
prohibitive and we need to use Monte Carlo sampling instead. The running
time depends on $p$, on the size of the reads $k$, and on the desired
number of iterations. Since those are constant throughout the sequencing
run, the method always takes the same time (around 1-10 seconds for reads
of size $100$-$200$ nucleotides on modern hardware). The values of $N$ and
$\mu$ can be binned in intervals of interest so that there are only around
$100$ pairs for a total cost of a few minutes per run. Considering that
mappers can process at most $10,000$ reads per second per core, the time
of mapping a sequencing run of 250 million reads is over 7 hours, two
orders of magnitude larger than the time required to estimate the
probabilities of error.

Finally, one may wonder if our approach has any advantage over methods
based on machine learning. This class of algorithms has already proved
useful~\cite{lee2014mosaik} and the rapid progress in the field of deep
learning suggests that it is possible to train algorithms to accurately
estimate mapping quality. In time, such algorithms may prove faster and/or
more robust because they could learn intrinsic biases of the mapping
algorithms. Yet, the main benefit of our approach will remain: the
combinatorial constructions are a direct access to the nature of the
problem. For instance, viewing MEM seeds through the lens of hard and soft
masks turns a seemingly intractable process into a relatively simple one
(see algorithm~\ref{alg:mcmc}). The combinatorial stance is that there is
value in clarity.

In conclusion, we presented a practical solution to the problem of
estimating the probability of false positives when using seeding
heuristics. This solution is adapted for mapping short reads sequenced
with the Illumina technology. Being able to calibrate the seeding
heuristic not only allows the user to choose how to balance speed versus
accuracy, but also opens new applications. For instance, one can map reads
from contaminated samples in pools of closely related genomes
(\textit{e.g.}, modern human and Neanderthal) in order to assign the reads
to the organism they belong to. The probabilities of false positives give
the right level of confidence in the assignment.

More generally, the analytic combinatorics programme is a very powerful
tool to address problems in bioinformatics. Here we have seen how this
strategy can be useful even when the generating functions cannot be
computed. Using the same ideas, one could calibrate heuristics used in
other alignment methods, especially the expanding field of long-read
technologies.

\section*{Acknowledgements}

We thank Ruggero Cortini for suggesting Monte Carlo methods as a way to
tackle intractable cases. We acknowledge the financial support of the
Spanish Ministry of Economy and Competitiveness (‘Centro de Excelencia
Severo Ochoa 2013-2017’, Plan Nacional BFU2012-37168), of the CERCA
Programme~/~Generalitat de Catalunya, and of the European Research Council
(Synergy Grant 609989).

%---------------------------------------------------------------
%---------------------------------------------------------------

\bibliography{references,pubmed}
\bibliographystyle{unsrt}

%----------------------------------------------------------------

\end{document}

%gs -dNoOutputFonts -sDEVICE=pdfwrite -o out.pdf latex.pdf